### 2012-08-29 sections/design.tex
* moved sections to separate files in sections/ * added nfql pipeline figures * added flow-tools, nfdump, tcpdump, wireshare, libpcap blurb in related work * added design blurb

The pipeline consists of a number of independent processing elements that are
connected to one another using UNIX-based pipes. Each element receives the
content from the previous pipe, performs an operation and pushes it to the
next element in the pipeline. Fig. \ref{fig:nfql-pipeline} shows an
overview of the processing pipeline. A complete description on the semantics
of each element in the pipeline can be found in \cite{vmarinov:2009}

The splitter takes the flow-records data as input in the \texttt{flow-tools}
compatible format. It is responsible to duplicate the input data out to
several branches without any processing whatsoever. This
allows each of the branches to have an identical copy of the flow data to
process it independently.

The filter performs \emph{absolute} filtering on the input flow-records data.
The flow-records that pass the filtering criterion are forwarded to the
grouper, the rest of the flow-records are dropped. The
filter compares separate fields of a flow-record against either a constant
value or a value on a different field of the \emph{same} flow-record. The
filter cannot \emph{relatively} compare two different incoming flow-records

The grouper performs aggregation of the input flow-records data. It consists
of a number of rule modules that correspond to a specific subgroup. A
flow-record in order to be a part of the group should be a part of at-least
one subgroup. A flow-record can be a part of multiple subgroups within a
group. A flow-record cannot be part of multiple groups.
The grouping rules can be either absolute or relative. The newly formed groups
which are passed on to the group filter can also contain meta-information
about the flow-records contained within the group using the aggregate clause
defined as part of the grouper query.

The group-filter performs \emph{absolute} filtering on the input group-records
data. The group-records that pass the filtering criterion are forwarded to the
merger, the rest of the group-records are dropped. The group-filter compares
separate fields (or aggregated fields) of a
flow-record against either a constant value or a value on a different field of
the \emph{same} flow-record. The group-filter cannot \emph{relatively} compare
two different incoming group-records

The merger performs relative filtering on the N-tuples of groups formed from
the N stream of groups passed on from the group-filter as input. The merger
rule module consists of a number of a submodules, where the output of the
merger is the set difference of the output of the first
submodule with the union of the output of the rest of the submodules. The
relative filtering on the groups are applied to express timing and concurrency
constraints using Allen interval algebra \cite{fallen:1983}

The ungrouper unwraps the tuples of group-records into individual
flow-records, ordered by their timestamps. The duplicate
flow-records appearing from several group-records are eliminated and are sent
as output only once.

### 2012-08-29 sections/implementation.tex
* moved sections to separate files in sections/ * added nfql pipeline figures * added flow-tools, nfdump, tcpdump, wireshare, libpcap blurb in related work * added design blurb

\subsection{Splitter}
\subsection{Filter}
\subsection{Grouper}
\subsubsection{Group Aggregations}
\subsection{Group Filter}
\subsection{Merger}
\subsection{Ungrouper}

### 2012-08-29 sections/introduction.tex
* moved sections to separate files in sections/ * added nfql pipeline figures * added flow-tools, nfdump, tcpdump, wireshare, libpcap blurb in related work * added design blurb

\todo{...}

\cleardoublepage

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\linewidth]{nfql-pipeline}
\caption{NFQL Pipeline \cite{vmarinov:2009}}
\label{fig:nfql-pipeline}
\end{figure*}

### 2012-08-29 sections/relatedwork.tex
* moved sections to separate files in sections/ * added nfql pipeline figures * added flow-tools, nfdump, tcpdump, wireshare, libpcap blurb in related work * added design blurb

\texttt{flow-tools} \cite{sromig:2000} is a suite of programs for capturing
and processing NetFlow v5 flow records. It consists of 24 separate tools that
work together by connecting them via UNIX pipes. \todo{...}

\texttt{nfdump} \cite{phaag:2006} works similar to flow-tools but uses a
different storage format. Flow records are captured using \texttt{nfcapd} and
then processed by \texttt{nfdump} which can filter as well as display the
sorted and filtered result. The power of its filtering rules is similar to
that of flow-tools and as such is mostly limited to absolute comparisons of
flow attributes.

\texttt{tcpdump} and \texttt{wireshark} are the most popular tools used for
packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage} is a
premier command-line utility that uses the \texttt{libpcap}
\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
comes from the richness of its expressions, the ability to combine them using
logical connectives and extract specific portions of a packet using filters.
\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
both journeymen and packet analysis experts. It supports a large number of
protocols, has a straightforward layout, excellent documentation, and can run
on all major operating systems.

### 2012-08-31 sections/implementation/adaptable-compression-levels.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

The flow-records echoed to the standard output can also be written to a
Netflow $v5$ flow-tools file. The \texttt{--dirpath} switch allows one to
provide a directory path where the results can be stored.  Each stream is
stored as its own file with an ID to disambiguate it. Results from each stage
of the pipeline can also be written to separate files with the increase in the
verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose} work well
together to adjust the level at which the writes are to be made.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{zlevel}
    \caption{F$(v2)$: \texttt{z-level} Effect on Performance}
    \label{fig:engine-zlevel}
  \end{center}
\end{figure}

The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
the results written to the flow-tools files. \texttt{zlib} supports $9$
compression levels with $9$ being the highest. The \ac{NFQL} engine supports
\texttt{--zlevel} switch to allow the user to supply its desired choice of the
of the compression level. A default level of $5$ is used for writes if the
switch is not supplied during runtime.  Fig. \ref{fig:engine-zlevel} shows
the time taken to write a sample of records passing the filter stage for each
\texttt{z-level}. It goes to show that each level adds its own performance
overhead and must be used with discretion.

### 2012-08-31 sections/implementation/execution-workflow.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

\subsubsection{Splitter} A custom C library was written to directly read/write
data in the \texttt{flow-tools} format. The library sequentially reads the
complete flow-records into memory to support random access required for
relative filtering. Each flow-record is stored in a \texttt{char} array and
the offsets to each field are stored in separate \texttt{structs}.  The array
of such records are indexed allowing fast retrieval in $O(1)$ time. This
allowed the flow-records to be easily referenced by an identifier, thereby
giving away the need to every time copy all the flow-records when moving ahead
in the processing pipeline.

In the default method, there is a comparison function defined for every
possible field length $(33)$ and comparison operations $(19)$. These functions
are generated using a Python script. The rule definitions are now able to make
calls using a function name derived from the combination of field length,
delta type and operation.  This subverts the need to define complex branching
statements and reduces complexity.

\subsubsection{Filter}
\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{filter-fv1-fv2}
    \caption{Filter Stage: F$(v1)$ vs F$(v2)$}
    \label{fig:fv1-fv2-filter}
  \end{center}
\end{figure}

The flow query language suggests to read all the flow records of a supplied
trace into memory before starting the processing pipeline.  Since, the filter
stage uses the supplied set of absolute rules to make a decision on whether or
not to keep a flow record; it has to pass through the whole in-memory
recordset \emph{again} to fill in the filter results. This technique involves
multiple linear runs on the trace and therefore slows down when the ratio of
number of filtered records to the total number of flow-records is high. The
\ac{NFQL} implementation merges the filter stage with in-memory read of the
trace. This means, a decision on whether or not to make room for a record in
memory and eventually hold a pointer for it in filter results is done upfront
as soon as the record is read from the trace. In addition, if a request to
write the filter stage results to a flow-tools file has been made, the writes
are also made as soon the filter stage decision is available, thereby allowing
reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
records in the trace. It is important to note that the filtered records are
saved in common location from where they are referenced by each branch. This
helps keep the memory costs at a minimum when multiple branches are involved.
A publically available Netflow $v5$ trace
\footnote{\url{http://traces.simpleweb.org/traces/netflow/netflow1/netflow000.tar.bz2}}
was used to compare the performance of the new filter in F$(v2)$ with that
from F$(v1)$ as shown in Fig. \ref{fig:fv1-fv2-filter}. The trace has around
$20M$ flow-records.

\subsubsection{Grouper}

In order to be able to make comparisons on field offsets, a simple approach is
to linearly walk through each filtered record against the filtered recordset
leading to a complexity of $O(n^2)$, where $n$ is the number of filtered
records. A smarter approach is to put the copy in a hash table and then try to
map each pointer while walking down the filtered recordset once, leading to a
complexity of $O(n)$. The hash table approach, although will work on this
specific example, will fail badly on other relative comparisons. A better
approach, as implemented in F$(v2)$ is to sort the filtered recordset on all
the requested grouping rules. This helps the execution engine perform a
nested \texttt{bsearch} to reduce the linear pass to a fairly small filtered
recordset. However, the number of elements upon which the search has to be
performed needs to be known at each level of the binary lookup. In order to
avoid a linear run to count the number of elements, the parent level
\texttt{bsearch} invocation returns a boundary with the first and last element
to enable such a calculation. This helps the grouper perform faster search
lookups to find records that must group together in $O(n*lg(k))$ time with a
preprocessing step taking $O(n*lg(n)) + O(n)$ in the average case, where $n$
is the number of filtered records and $k$ is the number of unique filtered
records.

\begin{figure}[h!]
  \begin{center}
    \includegraphics*[width=1.0\linewidth]{grouper-fv2-bsearch-fv2}
    \caption{Grouper Stage: F$(v2)$ (Generic) vs F$(v2)$ (Specific)}
    \label{fig:fv1-fv2-grouper}
  \end{center}
\end{figure}

The grouping approach has further been optimized when the filtered records are
grouped for equality. In such a scenario, the need to search for unique
records and a  subsequent binary search goes away.  The groups can now be
formed in $O(n)$ time with a more involved preprocessing step taking
$O(p*n*lg(n)$ where is $n$ is the number of filtered records, and $p$ is the
number of grouping rules. The performance evaluation of the grouper handling
this special case against its behavior when handling generic cases is shown in
Fig. \ref{fig:fv1-fv2-grouper}. Profiling the performance of the grouper on
queries that produce higher ratios reveal that a significant amount of time is
taken in \texttt{bsearch}. The special case of equality comparisons eliminate
these calls and further reduce the time in the long tail as shown in Fig.
\ref{fig:fv1-fv2-grouper}.

\subsubsection{Group Aggregations} Group records are a conglomeration of
several flow records with some common characteristics defined by the flow
query. Some of the non-common characteristics can also be aggregated into a
single value using group aggregations. It is useful if a single group record
can be again mapped into a NetFlow $v5$ record template, so that it can be
written to file as a representative of all its members.

\subsubsection{Group Filter} The groupfilter is used to filter the groups
produced by the grouper based on some absolute rules defined in a \ac{DNF}
expression. The \texttt{struct} term holds information about the flow record
offset, the value being compared to and the operator which maps to a unique
\texttt{enum} value.  This enum value is used to map the operation to a
specific group-filter function. The group-filter functions are auto-generated
using a python script.

\subsubsection{Merger}

The merger is used to relate groups from different branches according to a
merging criterean. The implementation is not trivial since the number of
branches that need to be spawned is read from the query and is not known until
compile time. As a result, an iterator that can provide all possible
permutations of $m-$tuple (where $m$ is the number of branches) group record
$ID$s was needed.  The result of the iterator can then be used to make a
match. The merger stage, begins by initializing this iterator passing it the
number of branches, and information about each branch. Then, it loops over to
get a new $m-$tuple of group record $ID$s on each iteration until the iterator
returns \texttt{false}.

The merger as formulated in the flow query language needs to match each group
record from one branch with every other record of each branch. This leads to a
complexity of $O(n^m)$ where $n$ is the number of filtered group records and
$m$ is the number of branches. The possible number of tries when matching
group records however can be reduced by sorting the group records on the field
offsets used for a match. The \ac{NFQL} implementation optimizes the merger to
skip over iterator permutations when a state of a current field offset value
may not allow any further match beyond the index in the current branch.  For
such an optimization to work, the filtered group records must be sorted in the
order of field offsets specified in the merger clause. Specifying the filtered
group records in any other order may lead to undefined behavior. This means,
that if the same field offsets were used in the grouper stage, the terms in
the group clause can be rearranged by the query designer to align with the
order of terms in the merger clause.

The flow query language also bases the merger matches on the notion of matched
tuples. This means that a filtered group record can be written to a file
multiple times if it is part of multiple matched tuples. This situation is
very common and it worsens when different branches have similar filtered
groups records. Since, the function of the merger is to find a match of groups
records across branches based on a predefined condition, all the group records
across branches that satisfy the condition can be clubbed into one collection
instead of separate tuples. All the group records within a collection can then
be written to the file. This eliminates the inherent redundancy and
significantly improves the merger performance.

The performance comparison of this approach against the one suggested by the
specification is tricky. The merger implementation of the original
specification is slow. It is so slow that it keeps churning the CPU for days
without results. The newer approach takes less than an hour. The performance
evaluation of this newer approach is discussed in more details in the next
section

\subsubsection{Ungrouper} The approach of clubbing the merged group records into
a collection incurs a reimplementation of the ungrouper. The ungrouper, as a
result accepts a collection of matched filtered group records as input. It
then iterates over each collection to unfold it groups and write their flow
record members to files.

### 2012-08-31 sections/implementation/flow-query-intermediate-format.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

A \ac{DNF} is a disjunction of conjunctive clauses. The elements of the
conjuctive clauses are terms. Each stage of the pipeline is represented in the
flowquery as a \ac{DNF} expression. The clauses in the \ac{DNF} are OR'd
together, while the terms in each clauses themselves are AND'd. The abstract
objects that store the JSON query and the results that incubate from each
stage are designed to be self-descriptive and hierarchically chainable. The
complete JSON query information for instance, is held within the
\texttt{flowquery} struct. Each individual branch of the flowquery itself is
described in a \texttt{branch} struct. A collection of these \texttt{branch}
structs are referenced in the parent \texttt{flowquery} struct.

### 2012-08-31 sections/implementation/flow-query-intermediate-format.tex
* flow query intermediate format   * parsing using json-c   * generating json queries using python scripts   * runtime query internals

@@
-A \ac{DNF} is a disjunction of conjunctive clauses. The elements of the
-conjuctive clauses are terms. Each stage of the pipeline is represented in the
-flowquery as a \ac{DNF} expression. The clauses in the \ac{DNF} are OR'd
-together, while the terms in each clauses themselves are AND'd. The abstract
-objects that store the JSON query and the results that incubate from each
-stage are designed to be self-descriptive and hierarchically chainable. The
-complete JSON query information for instance, is held within the
-\texttt{flowquery} struct. Each individual branch of the flowquery itself is
-described in a \texttt{branch} struct. A collection of these \texttt{branch}
-structs are referenced in the parent \texttt{flowquery} struct.
+The \ac{NFQL} execution engine reads the flow query at runtime in a
+\texttt{JSON} intermediate format. The entire flow query is a collection of
+\ac{DNF} expression. A \ac{DNF} expression is a disjunction of conjunctive
+clauses. The elements of the conjuctive clauses are terms. The clauses in the
+\ac{DNF} are OR'd together, while the terms in each clauses themselves are
+AND'd. The branchsets and each \ac{DNF} expression of the pipeline stage is a
+\texttt{JSON array}. \texttt{json-c}\footnote{\url{http://oss.metaparadigm.com/json-c/}} is used to
+parse the flow query file, and python scripts have been written to quickly
+serialize the python pipeline objects to a \texttt{JSON} query.
+
+The mapping of the \texttt{JSON} query to the structs defined in the execution
+engine is not trivial. When reading the \texttt{JSON} query at runtime, the
+field offsets of the NetFlow $v5$ record struct are read in as strings.
+Utility functions are defined that maps the read names to struct offsets and
+read type of each offset and the operations to unique \texttt{enum} members.
+
+The abstract objects that store the JSON query and the results that incubate
+from each stage are designed to be self-descriptive and hierarchically
+chainable. The complete JSON query information for instance, is held within
+the \texttt{flowquery} struct. Each individual branch of the flowquery itself
+is described in a \texttt{branch} struct. A collection of these
+\texttt{branch} structs are referenced in the parent \texttt{flowquery}
+struct.

### 2012-09-01 sections/conclusion.tex
* introduction   * traffic measurement approach blurb   * flow export protocol blurb * related work   * silk; and how it differs from nfql * performance evaluations   * stressing the filter   * stressing the grouper   * stressing the group filter   * stressing the merger   * stressing the ungrouper * conclusion

We presented \ac{NFQL}, an efficient C implementation of the stream-based flow
query language. The language allows applying absolute (or relative) filters,
aggregating flows into groups, evaluating intricate timing relationships among
them, and merging them into one collection. \ac{NFQL} can execute such complex
query tasks in matter of minutes, thereby expanding the scope of current flow
record processing tools. The conducted performance evaluations reveal that
\ac{NFQL} is on par with tools that support only absolute filters. SiLK, the
only package that provides tools that are similar to the rest of \ac{NFQL}'s
processing pipeline appears faster. This is because it can optimize its
operations in favor of the limited set of comparisons that are only based on
equality, and its usage of a proprietory file storage format.
\todo{...}

### 2012-09-01 sections/evaluation.tex
* introduction   * traffic measurement approach blurb   * flow export protocol blurb * related work   * silk; and how it differs from nfql * performance evaluations   * stressing the filter   * stressing the grouper   * stressing the group filter   * stressing the merger   * stressing the ungrouper * conclusion

We used a public flow trace containing $20M$ records to run a number of flow
queries using our benchmarking suite. We used trace $7$
\footnote{\url{http://traces.simpleweb.org/traces/netflow/netflow1/netflow000.tar.bz2}},
from Simpleweb \footnote{Simpleweb is a data repository of traffic traces from
University of Twente}.  The input trace was compressed at \texttt{ZLIB\_LEVEL}
$5$ using the \texttt{zlib} suite. It was also converted to \texttt{nfdump}
and SiLK compatible formats and compressed with \texttt{zlib}
keeping the same compression level.  The suite was run on a high-end
machine\footnote{\url{crystal.eecs.jacobs-university.de}} with $24$ cores of
$2.5$ GHz clock speed and $18$ GiB of memory.

The first set of queries attempt to stress the filter stage.  We use varying
values on the packet field offset to determine the amount of flow records that
are passed by the filter. The resultant filtered records are written to
\texttt{flow-tools} compatible file format and compressed at
\texttt{ZLIB\_LEVEL} $5$ using \texttt{zlib} suite. The ratio of the number of
filtered records in the output trace to the number of the flow records in the
input trace is plotted against time. The evaluation results are shown in
Fig. \ref{fig:benchmarks-filter}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{filter}
    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
    \label{fig:benchmarks-filter}
  \end{center}
\end{figure}

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then again SilK and \texttt{nfdump} also use their own
proprietary format for trace files. As a result, the amount of data that needs
to read (or written) may be different to what it is for \ac{NFQL} and
\texttt{flowtools}. On the other hand, \texttt{nfdump} appears to be
significantly faster than the rest. This is because \texttt{nfdump} lacks
\texttt{zlib} support, and as such the files that are read and written used
\texttt{lzo} compression scheme which trades space for achieving faster
compression and decompression. It is important to note, that all the tools
were single-threaded in this evaluation, and did not completely utilize the
$24$ cores that were at their disposal. It comes as a realization, that
filtering the input using multiple threads by memory mapping the trace and
adding \texttt{lzo} compression will drastically improve \ac{NFQL}'s filter
performance.

The second set of queries attempt to stress the grouper stage. We reuse the
filter query that produces a $1.0$ ratio to allow the grouper to receive the
entire trace as a filtered recordset. The grouper part of the query then
gradually increases the number of grouping terms in the \ac{DNF} expression to
increase the output/input ratio. The resultant groups are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of groups formed to the number of the input filtered
records is plotted against time. \texttt{nfdump} and \texttt{flowtools} do not
support grouping, and therefore are not considered in this evaluation. The
evaluation results are shown in Fig. \ref{fig:benchmarks-grouper}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{grouper}
    \caption{Grouper Stage: NFQL vs SiLK}
    \label{fig:benchmarks-grouper}
  \end{center}
\end{figure}

The evaluation graph reveals that the performance of the \ac{NFQL} grouper
stage is close. The time taken by the tools are comparable on lower ratios,
but on higher ratios, \ac{NFQL} starts to drift apart. Since most of the time
is taken in writing the records to files, it is unclear whether SiLK's usage
of a proprietary format which may reduce reads/writes is responsible for the
drift on higher ratios. SiLK's \texttt{rwgroup} tool is also supplied a
\texttt{--summarize} flag in all the queries. This gives SiLK the leverage to
not store information about which members are part of the group.  \ac{NFQL} on
the other hand needs to allocate resources (which may take time) to keep this
information in its data structures, since the ungrouper later may request to
write the members of a group while unfolding the tuples. It is also important
to note that both the tools again remained single-threaded throughout the
evaluation. SiLK took an advantage of an inherent concurency arising from how
the query is structured as one single \texttt{bash} script using pipes. The
pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two process run
concurrently, the effect of which gets more pronounced on higher ratios and
can be a drift determining factor.  The profiling results from GNU
\texttt{gprof} \cite{graham:1982} indicate that $60\%$ of the time is taken in
\texttt{qsort} comparator calls.  As a result, it comes as no surprise, that
bifurcating \texttt{qsort} invocation to multiple threads and later merging
the results back using merge sort will help parallelize the grouper stage and
maybe reduce the drift on higher ratios. In addition, since all of the
evaluation queries had grouping terms using an equality comparator, \ac{NFQL}
can introspect such a grouping rule to dynamically optimize processing
searches using a hashtable and turn to \texttt{qsort} based grouping only as a
fallback.

The third set of queries attempt to stress the group filter stage. We reuse
the filter and grouper queries that produce a $1.0$ ratio to allow the group
filter to receive the entire trace as input. This means, each flow record of
the original trace now becomes a group record for the group filter.  The group
filter then reuses the same varying values of the packet offset to determine
the amount of groups that are filtered ahead. The filtered groups are again
written as \texttt{flowtools} files using the same \texttt{zlib} compression
level. The ratio of the number of filtered groups formed to the number of the
input group records is plotted against time. The evaluation results are shown
in Fig. \ref{fig:benchmarks-groupfilter}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{groupfilter}
    \caption{Group Filter Stage: NFQL vs SiLK}
    \label{fig:benchmarks-groupfilter}
  \end{center}
\end{figure}

It can be seen that the timings of \ac{NFQL} are far apart from that of SiLK.
It is due to the drift already created by the grouper at $1.0$ ratio in the
previous stage. As a result, the group filter comes into play only after $300$
seconds, whereas SiLK's group filtering already starts just below $150$
seconds. Even if we normalize the graph, it can be observed that the group
filter has a significantly higher slope. This is because it is only executed
once the grouper returns, and therefore has to reiterate the groups to make a
filtering decision.

The fourth set of queries attempt to stress the merger stage. We reuse the
filter, grouper and group filter queries that produce a $1.0$ ratio. These
queries are then run in two separate branches to produce identical filtered
group records.  The merger then applies match rules to produce different
output to input ratios. The groups that are merged are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of merged groups to twice\footnote{Each branch pushes the
entire trace as an input to the merger.} the number of flow records in the
original trace is plotted against time. A data point for SiLK for the $0.2$
ratio is not available since the \ac{NFQL} query executed at that data point
uses \texttt{OR} expressions which are not supported by SiLK. As a result, an
equivalent SiLK query is not formulated.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{merger}
    \caption{Merger Stage}
    \label{fig:benchmarks-merger}
  \end{center}
\end{figure}

It can be seen that the merger is the most performance hit and time consuming
stage of the \ac{NFQL} pipeline thus far. It is due to the fact that the
merger is working on twice the number of flow records than any other previous
stage. In addition, each branch is writing the results of the filter, grouper
and group filter stage to \texttt{flowtools} files. As a result, the amount of
disk I/O involved is twice as much as well. Even though each branch is
delegated to a separate core using affinity masks, most of time is taken in
writing these results to the file. These results although look less promising,
they are way better than the previous merger implementation. The newer merger
takes advantage of sorted nature of filtered groups and therefore can
signifantly reduce the number of merger matches. It also writes a merged group
record to file only once despite the number of times it has matched. Without
these optimizations, running such queries on the merger would keep the CPU
churning for days without results.

The last set of queries attempt to stress the ungrouper stage. They reuse the
entire merger queries as is, but enable the ungrouper now as well. This means,
that the ungrouper now attempts to unfold the merged groups returned by the
merger to write their member flow records to \texttt{flowtools} files.
However, since the merger receives each flow record as its own filtered group,
each merged group has only one member. As a result, the ungrouper ends up
executing its logic and rewriting the merged groups as \texttt{flowtools}
files using the using the same \texttt{zlib} compression level. The ratio of
the number of result flow records to twice the number of the flow records in
the input trace is plotted against time. The evaluation results are shown in
Fig. \ref{fig:benchmarks-ungrouper}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{ungrouper}
    \caption{Ungrouper Stage}
    \label{fig:benchmarks-ungrouper}
  \end{center}
\end{figure}

It can be seen that the evaluation behavior is similar to that shown by the
merger, but is just more involved. This is because the ungrouper also has to
write the entire merged groupset to file, as is done by merger. This is the
reason for the execution engine in taking twice the amount of time. In fact,
this evaluation is performed only to stress the functioning of the ungrouper
and running queries that merge each filtered group filtered individually is
less useful in practise. Ideally, the execution engine must shutdown the
ungrouper if the merger merges every filtered group record in its output. This
is because, such a behavior implicitly eliminates the need of the ungrouper.

### 2012-09-01 sections/introduction.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

@@
+Researchers, service providers and security analysts have long been interested
+in network and user behavioral patterns of the traffic crossing the internet
+backbone. They want to use this information for the purpose of billing and
+mediation, bandwidth provisioning, detecting malicious attacks, network
+performance evaluation and overall improvement. Traffic measurement techniques
+that have been rapidly evolving in the last decade, have matured enough today
+to provide such an insight.
+
+Flow capture today, has emerged out to be one of the favored network
+measurement techniques. This has largely been due to the reduction in the
+monitoring traffic at the flow-level and the fine-grained control which was
+not previously possible using SNMP interface-level queries. As a result,
+each networking vendor has tried to come up with a standard protocol that
+defines the semantics of this flow export. In this pursuit, Cisco eventually
+managed to make their proprietary protocol so ubiquitously available, that the
+next-generation universal standard is based on it.
+
 \todo{...}
-
-\cleardoublepage
 

### 2012-09-01 sections/relatedwork.tex
* introduction   * traffic measurement approach blurb   * flow export protocol blurb * related work   * silk; and how it differs from nfql * performance evaluations   * stressing the filter   * stressing the grouper   * stressing the group filter   * stressing the merger   * stressing the ungrouper * conclusion

@@
+\texttt{tcpdump} and \texttt{wireshark} are the most popular tools used for
+packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage} is a
+premier command-line utility that uses the \texttt{libpcap}
+\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
+comes from the richness of its expressions, the ability to combine them using
+logical connectives and extract specific portions of a packet using filters.
+\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
+both journeymen and packet analysis experts. It supports a large number of
+protocols, has a straightforward layout, excellent documentation, and can run
+on all major operating systems.
+
 \texttt{flow-tools} \cite{sromig:2000} is a suite of programs for capturing
@@
 
-\texttt{tcpdump} and \texttt{wireshark} are the most popular tools used for
-packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage} is a
-premier command-line utility that uses the \texttt{libpcap}
-\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
-comes from the richness of its expressions, the ability to combine them using
-logical connectives and extract specific portions of a packet using filters.
-\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
-both journeymen and packet analysis experts. It supports a large number of
-protocols, has a straightforward layout, excellent documentation, and can run
-on all major operating systems.
+SiLK\footnote{\url{http://tools.netsa.cert.org/silk/}} is a network traffic
+collection and analysis tool developed and maintained by the CERT Network
+Situational Awareness Team (CERT NetSA) at Carnegie Mellon University. SiLK is
+the tool that comes quite close to providing similar capabilities as provided
+by \ac{NFQL} and is therefore used as a reference point to compare the
+performance of the \ac{NFQL} execution engine in this paper. The design and
+implementation of SiLK, however, differs a lot from that of \ac{NFQL}.For
+instance, in SiLK there are separate tools to perform the task of each stage
+of the \ac{NFQL} processing pipeline. The stage functionality is not
+full-fledged though. The grouping and merging operations can only be performed
+using an equality operator. This is assumed in the tool, thereby allowing it
+to perform optimization such as using hash tables to perform lookups. There
+are also stringent requirements to how flow-data needs to be organized before
+it can be piped into a tool. The grouping tool, for instance, assumes that the
+to-be supplied input flow data is already sorted on the field column.  These
+requirements make it a little cumbersome to design a full-fledged flow query.
+For instance, trying to mimic a \ac{NFQL} query in SiLK ends up as a bash
+script with over a dozen of SiLK tools piped together.

### 2012-09-01 sections/implementation/flow-query-intermediate-format.tex
* implementation   * flow query intermediate format     * json query disables the stages   * performance optimizations     * ignoring redundant aggregation requests     * lazy comparator evaluations     * early filtering     * early thread exits     * context-aware pipeline stages     * inline writes

@@
 serialize the python pipeline objects to a \texttt{JSON} query.
+
+The \texttt{JSON} query can also trigger and disable the stages at runtime.
+This means that one only has to supply the constructs that one wishes to use.
+The constructs that are not defined in the \texttt{JSON} query are inferred by
+the execution engine as a disable request. The execution engine uses disable
+flags that are turned on when the \texttt{JSON} query is parsed. These flags
+are used throughout the engine to only enable the requested functionality.
 

### 2012-09-01 sections/implementation/performance-optimizations.tex
* implementation   * flow query intermediate format     * json query disables the stages   * performance optimizations     * ignoring redundant aggregation requests     * lazy comparator evaluations     * early filtering     * early thread exits     * context-aware pipeline stages     * inline writes

There can be a situtation where the query designer may incorrectly ask for
aggregation on a field already specified in a grouper (or filter) clause. If
the relative operator is an equality comparison, the aggregation on such a
field becomes less useful, since the members of the grouped record will always
have the same value for that field. \ac{NFQL} realizes this redundant request
and ignores such aggregations.

\ac{NFQL} has dedicated comparator functions for each type of operation and
the type of the field offset it operates upon. It is not guaranteed that given
the type of the query and trace, the program will eventually go through each
stage of the pipeline. It is also possible that the program exits before,
because there is nothing more for the next stage to compute. The function
pointers are therefore set as late as possible and are are called from their
respective stages just before the comparison is needed.  As a result, we save
the computation time wasted in setting the function pointer for stage if it is
is never executed.

\ac{NFQL} pushes the filter stage out of the branch and closer to where the
trace is originally read. This enables the memory allocation of only records
that passed the filter stage thereby reducing the runtime memory footprint of
the execution engine.

Each branch runs in its own thread. If any of the stages of the branch return
a \texttt{NULL} when returning from their public interface function, there is
no reason to continue the thread. The subsequent stages of the branch cannot
do much with a \texttt{NULL} result.  Therefore, the branch thread returns
with a \texttt{EXIT\_FAILURE} if either stage returns \texttt{NULL}, and with
\texttt{EXIT\_SUCCESS} on normal exit.

Each stage of the processing pipeline is dependent on the result of the
previous one. As a result, the stages should only proceed and process, when
the previous returned results. Implementing such a response was
straightforward for the grouper and group filter, the merger although was a
little trickier.  The merger stage proceeds only when every branch has
non-zero filtered groups.  The iterator initializer
\texttt{iter\_init(\ldots)} deallocates and returns \texttt{NULL} if any one
branch has $0$ filtered groups.  Consequently a check is performed in the
merger to make sure \texttt{iter} is \emph{not} \texttt{NULL}.

The results from each stage of the pipeline are echoed to the standard output
just before the execution engine exits. This leads to an additional loop to
echo the results, however echoes to the standard output are only used for
debugging purposes. The writes to a file can however be legitimately requested
in a real network analysis task. As a result, it is essential to avoid
additional loops when performing writes to a file. The execution engine,
therefore writes each result record to a file as soon as it is seen by the
pipeline stage.

### 2012-09-02 sections/design.tex
* introduction   * capturing flows     * elevator pitch blurb     * netflow and ipfix     * wide applicability of flow capture techniques     * flow query language     * NFQL implementation     * paper organization * related work   * tcpdump, wireshark, flow-tools, nfdump, silk   * silk citation, simpleweb citation * flow query lanauage

@@
-The pipeline consists of a number of independent processing elements that are
-connected to one another using UNIX-based pipes. Each element receives the
-content from the previous pipe, performs an operation and pushes it to the
-next element in the pipeline. Fig. \ref{fig:nfql-pipeline} shows an
-overview of the processing pipeline. A complete description on the semantics
-of each element in the pipeline can be found in \cite{vmarinov:2009}
+The flow query language pipeline consists of a number of independent elements
+that are connected to one another using UNIX-based pipes.  Each element
+receives the content from the previous pipe, performs an operation and pushes
+it to the next element in the pipeline as shown in Fig.
+\ref{fig:nfql-pipeline}. A complete description on the semantics of each
+element in the pipeline can be found in \cite{vmarinov:2009}
 
-The splitter takes the flow-records data as input in the \texttt{flow-tools}
-compatible format. It is responsible to duplicate the input data out to
-several branches without any processing whatsoever. This
-allows each of the branches to have an identical copy of the flow data to
-process it independently.
+The pipeline starts with the splitter. It takes the flow-records data as input
+in the \texttt{flow-tools} compatible format. The splitter is responsible to
+duplicate the input data out to several branches without any processing
+whatsoever. This allows each of the branch to receive an identical copy of the
+flow data to process it independently.
 
-The filter performs \emph{absolute} filtering on the input flow-records data.
-The flow-records that pass the filtering criterion are forwarded to the
-grouper, the rest of the flow-records are dropped. The
-filter compares separate fields of a flow-record against either a constant
-value or a value on a different field of the \emph{same} flow-record. The
-filter cannot \emph{relatively} compare two different incoming flow-records
+The filter is the first processing element of a branch in the pipeline.  It
+performs \emph{absolute} filtering on the input flow-records data.  The
+flow-records that pass the filtering criterion are forwarded to the grouper,
+the rest of the flow-records are dropped. The filter compares separate fields
+of a flow-record against either a constant value or a value on a different
+field of the \emph{same} flow-record. The filter cannot \emph{relatively}
+compare two different incoming flow-records
 
 The grouper performs aggregation of the input flow-records data. It consists
-of a number of rule modules that correspond to a specific subgroup. A
-flow-record in order to be a part of the group should be a part of at-least
-one subgroup. A flow-record can be a part of multiple subgroups within a
-group. A flow-record cannot be part of multiple groups.
-The grouping rules can be either absolute or relative. The newly formed groups
-which are passed on to the group filter can also contain meta-information
-about the flow-records contained within the group using the aggregate clause
-defined as part of the grouper query.
+of a number of terms that correspond to a specific subgroup. A flow-record in
+order to be a part of the group should satisfy at-least one subgroup. A
+flow-record can be a part of multiple subgroups within a group. A flow-record
+cannot be part of multiple groups.  The grouping rules can be either absolute
+or relative. The newly formed groups which are passed on to the group filter
+can also contain meta-information about the flow-records contained within the
+group using the aggregate clause defined as part of the grouper query.
 
-The group-filter performs \emph{absolute} filtering on the input group-records
-data. The group-records that pass the filtering criterion are forwarded to the
-merger, the rest of the group-records are dropped. The group-filter compares
-separate fields (or aggregated fields) of a
-flow-record against either a constant value or a value on a different field of
-the \emph{same} flow-record. The group-filter cannot \emph{relatively} compare
-two different incoming group-records
+The group-filter is the last processing element of the branch. It performs
+\emph{absolute} filtering on the input group-records data. The group-records
+that pass the filtering criterion are forwarded to the merger, the rest of the
+group-records are dropped. The group-filter compares separate fields (or
+aggregated fields) of a flow-record against either a constant value or a value
+on a different field of the \emph{same} flow-record. The group-filter cannot
+\emph{relatively} compare two different incoming group-records.
 
 The merger performs relative filtering on the N-tuples of groups formed from
-the N stream of groups passed on from the group-filter as input. The merger
-rule module consists of a number of a submodules, where the output of the
-merger is the set difference of the output of the first
-submodule with the union of the output of the rest of the submodules. The
-relative filtering on the groups are applied to express timing and concurrency
-constraints using Allen interval algebra \cite{fallen:1983}
-
-The ungrouper unwraps the tuples of group-records into individual
-flow-records, ordered by their timestamps. The duplicate
-flow-records appearing from several group-records are eliminated and are sent
-as output only once.
+the N stream of groups passed on from the group-filter as input. The relative
+filtering on the groups are applied to express timing and concurrency
+constraints using Allen interval algebra \cite{fallen:1983}. The ungrouper is
+the final processing element of the pipeline. It unwraps the tuples of
+group-records into individual flow-records, ordered by their timestamps. The
+duplicate flow-records appearing from several group-records are eliminated and
+are sent as output only once.

### 2012-09-02 sections/evaluation.tex
* introduction   * capturing flows     * elevator pitch blurb     * netflow and ipfix     * wide applicability of flow capture techniques     * flow query language     * NFQL implementation     * paper organization * related work   * tcpdump, wireshark, flow-tools, nfdump, silk   * silk citation, simpleweb citation * flow query lanauage

@@
 We used a public flow trace containing $20M$ records to run a number of flow
-queries using our benchmarking suite. We used trace $7$
-\footnote{\url{http://traces.simpleweb.org/traces/netflow/netflow1/netflow000.tar.bz2}},
-from Simpleweb \footnote{Simpleweb is a data repository of traffic traces from
-University of Twente}.  The input trace was compressed at \texttt{ZLIB\_LEVEL}
-$5$ using the \texttt{zlib} suite. It was also converted to \texttt{nfdump}
-and SiLK compatible formats and compressed with \texttt{zlib}
-keeping the same compression level.  The suite was run on a high-end
+queries using our benchmarking suite. We used the publically available Trace 7
+from SimpleWeb \cite{simpleweb} repository for all the performance
+evaluations. The input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using
+the \texttt{zlib} suite. It was also converted to \texttt{nfdump} and SiLK
+compatible formats and compressed with \texttt{zlib} keeping the same
+compression level. The suite was run on a high-end
 machine\footnote{\url{crystal.eecs.jacobs-university.de}} with $24$ cores of

### 2012-09-02 sections/implementation/execution-workflow.tex
* introduction   * capturing flows     * elevator pitch blurb     * netflow and ipfix     * wide applicability of flow capture techniques     * flow query language     * NFQL implementation     * paper organization * related work   * tcpdump, wireshark, flow-tools, nfdump, silk   * silk citation, simpleweb citation * flow query lanauage

@@
 helps keep the memory costs at a minimum when multiple branches are involved.
-A publically available Netflow $v5$ trace
-\footnote{\url{http://traces.simpleweb.org/traces/netflow/netflow1/netflow000.tar.bz2}}
-was used to compare the performance of the new filter in F$(v2)$ with that
-from F$(v1)$ as shown in Fig. \ref{fig:fv1-fv2-filter}. The trace has around
-$20M$ flow-records.
+We used the publically available Trace 7 from the SimpleWeb \cite{simpleweb}
+to compare the performance of the new filter in F$(v2)$ with that from F$(v1)$
+as shown in Fig. \ref{fig:fv1-fv2-filter}. The trace has around $20M$
+flow-records.
 

### 2012-09-02 sections/introduction.tex
* introduction   * capturing flows     * elevator pitch blurb     * netflow and ipfix     * wide applicability of flow capture techniques     * flow query language     * NFQL implementation     * paper organization * related work   * tcpdump, wireshark, flow-tools, nfdump, silk   * silk citation, simpleweb citation * flow query lanauage

@@
 Flow capture today, has emerged out to be one of the favored network
-measurement techniques. This has largely been due to the reduction in the
-monitoring traffic at the flow-level and the fine-grained control which was
-not previously possible using SNMP interface-level queries. As a result,
-each networking vendor has tried to come up with a standard protocol that
-defines the semantics of this flow export. In this pursuit, Cisco eventually
-managed to make their proprietary protocol so ubiquitously available, that the
-next-generation universal standard is based on it.
+measurement techniques. In this technique, packets traversing a monitoring
+point are not captured raw, instead they are aggregated together based on some
+common characteristics. The common characteristics are learnt by inspecting
+the packet headers as they cross the monitoring point. Flow-records resulting
+from such an aggregation are then exported to a collector for further
+analysis. This not only reduces the amount of traffic at the monitoring point,
+but also provides fine-grained control over the network data which was not
+previously possible using SNMP interface-level queries.
 
-\todo{...}
+NetFlow and \ac{IPFIX} are the two popular standards of IP flow information
+export. NetFlow \cite{rfc3954} is a proprietary network protocol designed by
+Cisco Systems. It allow routers to generate and export flow records to a
+designated collector. NetFlow v$9$ provides flexibility of user-tailored
+export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
+\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF deemed
+to be the logical successor of NetFlow v$9$ on which it is based. The novelty
+of the standard lies in its ability to describe record formats at runtime
+using templates based on an extensible and well-defined information model. The
+data transfer mechanism is also simplistic and extensible by being
+unidirectional and transport protocol agnostic.
+
+The wide applicability of this approach is easily seen from the pervasive use
+of flow records for a vibrant set of network analysis applications. For
+instance, the authors in \cite{mkim:2004} use the flow characteristics in the
+traffic pattern to formalize a detection function that maps traffic patterns
+to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
+use the flow-record data to exploit timing characteristics of webmail clients
+to classify features that could identify webmail traffic from any other
+traffic running over HTTPS.
+
+It goes to show that understanding these intricate traffic patterns require
+sophisticated flow analysis tools that can mine flow records for such a usage.
+Unfortunately current tools fail to deliver owing to their poor language
+design and simplistic filtering methods.  We recently proposed a flow query
+language design \cite{vmarinov:thesis:2009} that aims to cater to such needs.
+It can process flow records, aggregate them into groups, apply absolute (or
+relative) filters and invoke Allen interval algebra rules \cite{fallen:1983}.
+
+In this paper, we introduce \ac{NFQL}, an efficient C implementation of the
+flow query language. It is the next iteration of the first prototype
+implementation, Flowy \cite{kkanev:2010}, which was written in Python.
+\ac{NFQL} however is not just a reimplementation in a new language, but
+the inner workings have been reimagined to allow the execution engine to
+scale to real-world sized traces. This has been possible by replacing the
+performance hit stages of the pipeline with crispier algorithms that lower
+down the execution times to factor making them comparable to contemporary
+flow analysis tools.
+
+The rest of the paper is organized as follows. In section \ref{sec:design} we
+describe the flow query language by discussing each stage of the processing
+pipeline. In section \ref{sec:implementation} we introduce the inner workings
+of \ac{NFQL}. It begins by providing an overview of the \ac{NFQL} components
+and the structure of the intermediate format used to exchange messages. The
+workflow of the execution engine is described next and is supplemented by a
+number of performance optimizations made to make the implementation scale. It
+is followed by performance evaluation results by comparing them against
+contemporary flow processing tools in section \ref{sec:evaluation} and
+concludes with a discussion on its current limitation and future outlook in
+section \ref{sec:conclusion}.
 
@@
 \includegraphics[width=1.0\linewidth]{nfql-pipeline}
-\caption{NFQL Pipeline \cite{vmarinov:2009}}
+\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
 \label{fig:nfql-pipeline}

### 2012-09-02 sections/relatedwork.tex
* introduction   * capturing flows     * elevator pitch blurb     * netflow and ipfix     * wide applicability of flow capture techniques     * flow query language     * NFQL implementation     * paper organization * related work   * tcpdump, wireshark, flow-tools, nfdump, silk   * silk citation, simpleweb citation * flow query lanauage

@@
-\texttt{tcpdump} and \texttt{wireshark} are the most popular tools used for
-packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage} is a
-premier command-line utility that uses the \texttt{libpcap}
+The study of the network traffic patterns began with tools that utilized raw
+packet captures at the monitoring point for traffic measurement.
+\texttt{tcpdump} and \texttt{wireshark} have since then become the de-facto
+tools for packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage}
+is a premier command-line utility that uses the \texttt{libpcap}
 \cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
@@
 
-\texttt{flow-tools} \cite{sromig:2000} is a suite of programs for capturing
-and processing NetFlow v5 flow records. It consists of 24 separate tools that
-work together by connecting them via UNIX pipes. \todo{...}
-
-\texttt{nfdump} \cite{phaag:2006} works similar to flow-tools but uses a
+In recent years, a number of tools have been developed that can capture the
+traffic as flow records and use them for network analysis. \texttt{flow-tools}
+and \texttt{nfdump} are one of the most popular tools used for analyzing
+NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite of programs
+for capturing and processing NetFlow v5 flow records. It consists of 24
+separate tools that work together by connecting them via UNIX pipes. It can
+capture, read, filter, and print flow records internally saved in a fixed-size
+format. \texttt{nfdump} \cite{phaag:2006} is a very similar tool that uses a
 different storage format. Flow records are captured using \texttt{nfcapd} and
-then processed by \texttt{nfdump} which can filter as well as display the
-sorted and filtered result. The power of its filtering rules is similar to
-that of flow-tools and as such is mostly limited to absolute comparisons of
-flow attributes.
-
-SiLK\footnote{\url{http://tools.netsa.cert.org/silk/}} is a network traffic
-collection and analysis tool developed and maintained by the CERT Network
-Situational Awareness Team (CERT NetSA) at Carnegie Mellon University. SiLK is
-the tool that comes quite close to providing similar capabilities as provided
-by \ac{NFQL} and is therefore used as a reference point to compare the
-performance of the \ac{NFQL} execution engine in this paper. The design and
-implementation of SiLK, however, differs a lot from that of \ac{NFQL}.For
+then processed by \texttt{nfdump}. The power of filtering rules in both the
+tools is however mostly limited to absolute comparisons of flow attributes. As
+a result, relative comparison amongst different flows or querying a timing
+relationship among them is not possible.  SiLK \cite{SiLK} is a network
+traffic collection and analysis tool developed and maintained by the CERT
+Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon University.
+SiLK is the tool that comes quite close to providing similar capabilities as
+provided by \ac{NFQL} and is therefore used as a reference point to compare
+the performance of the \ac{NFQL} execution engine in this paper. The design
+and implementation of SiLK, however, differs a lot from that of \ac{NFQL}. For
 instance, in SiLK there are separate tools to perform the task of each stage

### 2012-09-02 sections/conclusion.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

@@
 \ac{NFQL} is on par with tools that support only absolute filters. SiLK, the
-only package that provides tools that are similar to the rest of \ac{NFQL}'s
-processing pipeline appears faster. This is because it can optimize its
-operations in favor of the limited set of comparisons that are only based on
-equality, and its usage of a proprietory file storage format.
+only openly available package that provides tools that are similar to the rest
+of \ac{NFQL}'s processing pipeline appears faster. This is because it can
+optimize its operations in favor of the limited set of comparisons that are
+only based on equality, and its usage of a different file storage format.
 \todo{...}

### 2012-09-02 sections/evaluation.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

@@
 We used a public flow trace containing $20M$ records to run a number of flow
-queries using our benchmarking suite. We used the publically available Trace 7
-from SimpleWeb \cite{simpleweb} repository for all the performance
+queries using our benchmarking suite. We used a publically available Trace 7
+from the SimpleWeb \cite{simpleweb} repository for all the performance
 evaluations. The input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using
@@
 compatible formats and compressed with \texttt{zlib} keeping the same
-compression level. The suite was run on a high-end
-machine\footnote{\url{crystal.eecs.jacobs-university.de}} with $24$ cores of
+compression level. The suite was run on a high-end machine with $24$ cores of
 $2.5$ GHz clock speed and $18$ GiB of memory.
@@
 The first set of queries attempt to stress the filter stage.  We use varying
-values on the packet field offset to determine the amount of flow records that
-are passed by the filter. The resultant filtered records are written to
-\texttt{flow-tools} compatible file format and compressed at
-\texttt{ZLIB\_LEVEL} $5$ using \texttt{zlib} suite. The ratio of the number of
-filtered records in the output trace to the number of the flow records in the
-input trace is plotted against time. The evaluation results are shown in
-Fig. \ref{fig:benchmarks-filter}.
+values on the \texttt{packet} field offset to determine the amount of flow
+records that are passed by the filter. The resultant filtered records are
+written to \texttt{flow-tools} compatible file format and compressed at
+\texttt{ZLIB\_LEVEL} $5$. The ratio of the number of filtered records in the
+output trace to the number of the flow records in the input trace is plotted
+against time. The evaluation results are shown in Fig.
+\ref{fig:benchmarks-filter}.
 
@@
 comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
-lower ratios, but then again SilK and \texttt{nfdump} also use their own
-proprietary format for trace files. As a result, the amount of data that needs
-to read (or written) may be different to what it is for \ac{NFQL} and
-\texttt{flowtools}. On the other hand, \texttt{nfdump} appears to be
-significantly faster than the rest. This is because \texttt{nfdump} lacks
-\texttt{zlib} support, and as such the files that are read and written used
-\texttt{lzo} compression scheme which trades space for achieving faster
-compression and decompression. It is important to note, that all the tools
-were single-threaded in this evaluation, and did not completely utilize the
-$24$ cores that were at their disposal. It comes as a realization, that
-filtering the input using multiple threads by memory mapping the trace and
+lower ratios, but then again SilK and \texttt{nfdump} also use their own file
+format. As a result, the amount of data that needs to read (or written) may be
+different to what it is for \ac{NFQL} and \texttt{flowtools}.  On the other
+hand, \texttt{nfdump} appears to be significantly faster than the rest. This
+is because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
+that are read and written used \texttt{lzo} compression scheme which trades
+space for achieving faster compression and decompression. It is important to
+note, that all the tools were single-threaded in this evaluation, and did not
+completely utilize the $24$ available cores. It comes as a realization, that
 adding \texttt{lzo} compression will drastically improve \ac{NFQL}'s filter
@@
 support grouping, and therefore are not considered in this evaluation. The
-evaluation results are shown in Fig. \ref{fig:benchmarks-grouper}.
+results are shown in Fig. \ref{fig:benchmarks-grouper}.
 
@@
 The evaluation graph reveals that the performance of the \ac{NFQL} grouper
-stage is close. The time taken by the tools are comparable on lower ratios,
-but on higher ratios, \ac{NFQL} starts to drift apart. Since most of the time
-is taken in writing the records to files, it is unclear whether SiLK's usage
-of a proprietary format which may reduce reads/writes is responsible for the
-drift on higher ratios. SiLK's \texttt{rwgroup} tool is also supplied a
-\texttt{--summarize} flag in all the queries. This gives SiLK the leverage to
-not store information about which members are part of the group.  \ac{NFQL} on
-the other hand needs to allocate resources (which may take time) to keep this
+stage is close to the performance of SiLK. The time taken by the tools are
+comparable on lower ratios, but on higher ratios, \ac{NFQL} starts to drift
+apart. Since most of the time is taken in writing the records to files, it is
+unclear whether SiLK's usage of a proprietary format, which may reduce
+reads/writes is responsible for the drift on higher ratios. SiLK's
+\texttt{rwgroup} tool is also supplied a \texttt{--summarize} flag to force it
+to write only the first record of each group, to make both tools write the
+same number of records to files. This gives SiLK the leverage to not store
+information about which members are part of the group. \ac{NFQL} on the other
+hand needs to allocate resources (which may take time) to keep this
 information in its data structures, since the ungrouper later may request to
@@
 the query is structured as one single \texttt{bash} script using pipes. The
-pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two process run
+pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two processes run
 concurrently, the effect of which gets more pronounced on higher ratios and
@@
 the original trace now becomes a group record for the group filter.  The group
-filter then reuses the same varying values of the packet offset to determine
-the amount of groups that are filtered ahead. The filtered groups are again
-written as \texttt{flowtools} files using the same \texttt{zlib} compression
-level. The ratio of the number of filtered groups formed to the number of the
-input group records is plotted against time. The evaluation results are shown
-in Fig. \ref{fig:benchmarks-groupfilter}.
+filter then reuses the same varying values of the \texttt{packet} field offset
+to determine the amount of groups that are filtered ahead. The filtered groups
+are again written as \texttt{flowtools} files using the same \texttt{zlib}
+compression level. The ratio of the number of output filtered groups to the
+number of the input group is plotted against time. The results are shown in
+Fig. \ref{fig:benchmarks-groupfilter}.
 
@@
 It can be seen that the timings of \ac{NFQL} are far apart from that of SiLK.
-It is due to the drift already created by the grouper at $1.0$ ratio in the
-previous stage. As a result, the group filter comes into play only after $300$
-seconds, whereas SiLK's group filtering already starts just below $150$
-seconds. Even if we normalize the graph, it can be observed that the group
-filter has a significantly higher slope. This is because it is only executed
-once the grouper returns, and therefore has to reiterate the groups to make a
-filtering decision.
+It is due to the drift already created by the grouper at the $1.0$ ratio in
+the previous stage. As a result, the group filter comes into play only after
+$300$ seconds, whereas SiLK's group filtering already starts just below $150$
+seconds. Even if we normalize the graph, it can be observed that the \ac{NFQL}
+group filter has a higher slope. This is because it is only executed once the
+grouper returns, and therefore has to reiterate the groups to make a filtering
+decision.
 
@@
 
-It can be seen that the merger is the most performance hit and time consuming
-stage of the \ac{NFQL} pipeline thus far. It is due to the fact that the
-merger is working on twice the number of flow records than any other previous
-stage. In addition, each branch is writing the results of the filter, grouper
-and group filter stage to \texttt{flowtools} files. As a result, the amount of
-disk I/O involved is twice as much as well. Even though each branch is
-delegated to a separate core using affinity masks, most of time is taken in
-writing these results to the file. These results although look less promising,
-they are way better than the previous merger implementation. The newer merger
-takes advantage of sorted nature of filtered groups and therefore can
-signifantly reduce the number of merger matches. It also writes a merged group
-record to file only once despite the number of times it has matched. Without
-these optimizations, running such queries on the merger would keep the CPU
-churning for days without results.
+It can be seen that the merger is the most performance critical stage of the
+\ac{NFQL} pipeline thus far. It is due to the fact that the merger is working
+on twice the number of flow records than any other previous stage. In
+addition, each branch is writing the results of the filter, grouper and group
+filter stage to \texttt{flowtools} files. As a result, the amount of disk I/O
+involved is twice as much as well. Even though each branch is delegated to a
+separate core using affinity masks, most of time is taken in writing these
+results to the file. These results although look less promising, they are way
+better than the original \ac{NFQL} merger implementation. The optimized merger
+implementation takes advantage of sorted nature of filtered groups and
+therefore can signifantly reduce the number of merger matches. It also writes
+a merged group record to file only once despite the number of times it has
+matched. Without these optimizations, running such queries on the merger would
+keep the CPU churning for days without results.
 
 The last set of queries attempt to stress the ungrouper stage. They reuse the
-entire merger queries as is, but enable the ungrouper now as well. This means,
+entire merger queries as is, but enable the ungrouper as well. This means,
 that the ungrouper now attempts to unfold the merged groups returned by the
@@
 each merged group has only one member. As a result, the ungrouper ends up
-executing its logic and rewriting the merged groups as \texttt{flowtools}
-files using the using the same \texttt{zlib} compression level. The ratio of
-the number of result flow records to twice the number of the flow records in
-the input trace is plotted against time. The evaluation results are shown in
-Fig. \ref{fig:benchmarks-ungrouper}.
+rewriting the merged groups as \texttt{flowtools} files using the same
+\texttt{zlib} compression level. The ratio of the number of result flow
+records to twice the number of the flow records in the input trace is plotted
+against time. The evaluation results are shown in Fig.
+\ref{fig:benchmarks-ungrouper}.
 
@@
 
-It can be seen that the evaluation behavior is similar to that shown by the
-merger, but is just more involved. This is because the ungrouper also has to
-write the entire merged groupset to file, as is done by merger. This is the
-reason for the execution engine in taking twice the amount of time. In fact,
-this evaluation is performed only to stress the functioning of the ungrouper
-and running queries that merge each filtered group filtered individually is
-less useful in practise. Ideally, the execution engine must shutdown the
-ungrouper if the merger merges every filtered group record in its output. This
-is because, such a behavior implicitly eliminates the need of the ungrouper.
+It can be seen that the evaluation behavior of the ungrouper is similar to
+that shown by the merger. Since, the ungrouper also has to write the entire
+merged groupset to file, as is previously done by merger, the execution engine
+takes twice the amount of time. It is important to note, that SiLK does not
+have such an equivalent ungrouping tool is therefore not considered in
+this final evaluation.

### 2012-09-02 sections/implementation.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

@@
-\subsection{Splitter}
-\subsection{Filter}
-\subsection{Grouper}
-\subsubsection{Group Aggregations}
-\subsection{Group Filter}
-\subsection{Merger}
-\subsection{Ungrouper}
+The \ac{NFQL} architecture primarily consists of a front-end parser backed up
+by an execution engine as shown in Fig. \ref{fig:architecture}. The execution
+engine is the brain of \ac{NFQL} where the complete pipeline is processed. It
+receives the flow query at runtime using a JSON intermediate format. The
+execution engine is designed to receive this input from either a machine
+running the front-end parser or a webservice that pushes the JSON from the
+cloud. The execution engine is written in C, however the front-end parser can
+be written in any desired language.
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics* [width=1.0\linewidth]{nfql-architecture}
+    \caption{NFQL Architecture}
+    \label{fig:architecture}
+  \end{center}
+\end{figure}
+
+\subsection{Flow Query Intermediate Format}
+\input{sections/implementation/flow-query-intermediate-format}
+\label{subsec:intermediate-format}
+
+\subsection{Execution Workflow}
+\input{sections/implementation/execution-workflow}
+\label{subsec:execution-workflow}
+
+\subsection{Performance Optimizations}
+\input{sections/implementation/performance-optimizations}
+\label{subsec:performance-optimizations}
+
+\subsection{Adaptable Compression Levels}
+\input{sections/implementation/adaptable-compression-levels}
+\label{subsec:adaptable-compression-levels}

### 2012-09-02 sections/implementation/adaptable-compression-levels.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

@@
 The flow-records echoed to the standard output can also be written to a
-Netflow $v5$ flow-tools file. The \texttt{--dirpath} switch allows one to
-provide a directory path where the results can be stored.  Each stream is
-stored as its own file with an ID to disambiguate it. Results from each stage
-of the pipeline can also be written to separate files with the increase in the
-verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose} work well
-together to adjust the level at which the writes are to be made.
+Netflow $v5$ \texttt{flow-tools} file. The \texttt{--dirpath} switch allows
+one to provide a directory path where the results can be stored.  Each stream
+is stored as its own file with an ID to disambiguate it. Results from each
+stage of the pipeline can also be written to separate files with the increase
+in the verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose}
+work well together to adjust the level at which the writes are to be made.
 
@@
 The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
-the results written to the flow-tools files. \texttt{zlib} supports $9$
-compression levels with $9$ being the highest. The \ac{NFQL} engine supports
-\texttt{--zlevel} switch to allow the user to supply its desired choice of the
-of the compression level. A default level of $5$ is used for writes if the
-switch is not supplied during runtime.  Fig. \ref{fig:engine-zlevel} shows
-the time taken to write a sample of records passing the filter stage for each
-\texttt{z-level}. It goes to show that each level adds its own performance
-overhead and must be used with discretion.
+the results written to the \texttt{flow-tools} files. \texttt{zlib} supports
+$9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
+supports \texttt{--zlevel} switch to allow the user to supply its desired
+choice of the of the compression level. A default level of $5$ is used for
+writes if the switch is not supplied during runtime.  Fig.
+\ref{fig:engine-zlevel} shows the time taken to write a sample of records
+passing the filter stage for each \texttt{z-level}. It goes to show that each
+level adds its own performance overhead and must be used with discretion.

### 2012-09-02 sections/implementation/execution-workflow.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

@@
-\subsubsection{Splitter} A custom C library was written to directly read/write
-data in the \texttt{flow-tools} format. The library sequentially reads the
-complete flow-records into memory to support random access required for
-relative filtering. Each flow-record is stored in a \texttt{char} array and
-the offsets to each field are stored in separate \texttt{structs}.  The array
-of such records are indexed allowing fast retrieval in $O(1)$ time. This
-allowed the flow-records to be easily referenced by an identifier, thereby
-giving away the need to every time copy all the flow-records when moving ahead
-in the processing pipeline.
+A custom C library has been written to directly read/write data in the
+\texttt{flow-tools} format. The library sequentially reads the complete
+flow-records into memory to support random access required for relative
+filtering. Each flow-record is stored in a \texttt{char} array and the offsets
+to each field are stored in separate \texttt{structs} as shown in listing
+\ref{lst:c-lib-struct}. The array of such records are indexed allowing fast
+retrieval in $O(1)$ time.
 
-In the default method, there is a comparison function defined for every
-possible field length $(33)$ and comparison operations $(19)$. These functions
-are generated using a Python script. The rule definitions are now able to make
-calls using a function name derived from the combination of field length,
-delta type and operation.  This subverts the need to define complex branching
-statements and reduces complexity.
+\lstset{caption=Trace Data Struct,
+				tabsize=2, language=C, numbers=left,stepnumber=1,
+        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
+        keywordstyle=\color{blue}, xleftmargin=15pt,
+        rulesepcolor=\color{black}, label=lst:c-lib-struct,
+        aboveskip=10pt, captionpos=b}
+\begin{lstlisting}
+struct ft_data {
+  int                                         fd;
+  struct ftio                                 io;
+  struct fts3rec_offsets                      offsets;
+  struct ftver                                version;
+  u_int64_t                                   xfield;
+  int                                         rec_size;
+  char**                                      recordset;
+  size_t                                      num_records;
+};
+\end{lstlisting}
+
+A sample JSON query \ac{DNF} term definition is shown in listing
+\ref{lst:filter-struct}. In order to be able to make comparisons on the field
+offsets of such a term, the comparator needs to know the type of the
+comparison and the length of the field offset. This information is parsed by
+execution engine once the query is read and is therefore not available at
+compile time.  In order to subvert the need to define complex branching
+statements, a dedicated comparator is defined for every possible field length
+and comparison operation. These comparators are generated using a Python
+script. This allows the term definitions to make runtime calls using a
+function name derived from the combination of operation type and field length.
+
+\lstset{caption=Filter Term Struct,
+				tabsize=2, language=C, numbers=left,stepnumber=1,
+        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
+        keywordstyle=\color{blue}, xleftmargin=15pt,
+        rulesepcolor=\color{black}, label=lst:filter-struct,
+        aboveskip=10pt, captionpos=b}
+\begin{lstlisting}
+struct filter_term {
+  size_t                                      field_offset;
+  uint64_t                                    value;
+  uint64_t                                    delta;
+  struct filter_op*                           op;
+  bool (*func)(
+               const char* const              record,
+               size_t                         field_offset,
+               uint64_t                       value,
+               uint64_t                       delta
+              );
+};
+\end{lstlisting}
+
+\subsubsection{Splitter} \ac{NFQL} uses identifiers to reference a flow record
+in the \texttt{char} array. This gives away the need to every time copy all
+the flow-records when moving ahead in the processing pipeline. As a result,
+there is no dedicated splitter stage in the execution engine. Each branch
+references the flow records from a common memory location. This helps keep the
+memory costs at a minimum when multiple branches are involved.
 
@@
     \includegraphics* [width=1.0\linewidth]{filter-fv1-fv2}
-    \caption{Filter Stage: F$(v1)$ vs F$(v2)$}
+    \caption{Filter Stage: NFQL vs NFQL (optimized)}
     \label{fig:fv1-fv2-filter}
@@
 
-The flow query language suggests to read all the flow records of a supplied
-trace into memory before starting the processing pipeline.  Since, the filter
-stage uses the supplied set of absolute rules to make a decision on whether or
-not to keep a flow record; it has to pass through the whole in-memory
-recordset \emph{again} to fill in the filter results. This technique involves
-multiple linear runs on the trace and therefore slows down when the ratio of
-number of filtered records to the total number of flow-records is high. The
-\ac{NFQL} implementation merges the filter stage with in-memory read of the
-trace. This means, a decision on whether or not to make room for a record in
-memory and eventually hold a pointer for it in filter results is done upfront
-as soon as the record is read from the trace. In addition, if a request to
-write the filter stage results to a flow-tools file has been made, the writes
-are also made as soon the filter stage decision is available, thereby allowing
+The execution engine, as defined by the flow query language must read all the
+flow records of a supplied trace into memory before starting the processing
+pipeline.  Since, the filter stage uses the supplied set of absolute rules to
+make a decision on whether or not to keep a flow record; it has to pass
+through the whole in-memory recordset \emph{again} to fill in the filter
+results. This technique involves multiple linear runs on the trace and
+therefore slows down when the ratio of number of filtered records to the total
+number of flow-records is high. We optimized this behavior in \ac{NFQL} by
+merging the filter stage with in-memory read of the trace. This means, a
+decision on whether or not to make room for a record in memory and eventually
+hold a pointer for it in filter results is done upfront as soon as the record
+is read from the trace. In addition, if a request to write the filter stage
+results to a flow-tools file has been made, the writes are also made as soon
+the filter stage decision is available, thereby allowing
 reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
-records in the trace. It is important to note that the filtered records are
-saved in common location from where they are referenced by each branch. This
-helps keep the memory costs at a minimum when multiple branches are involved.
-We used the publically available Trace 7 from the SimpleWeb \cite{simpleweb}
-to compare the performance of the new filter in F$(v2)$ with that from F$(v1)$
-as shown in Fig. \ref{fig:fv1-fv2-filter}. The trace has around $20M$
-flow-records.
+records in the trace. We used the publically available Trace 7 from the
+SimpleWeb \cite{simpleweb} to compare the performance of the optimized
+\ac{NFQL} implementation against the one defined by the flow query language as
+shown in Fig. \ref{fig:fv1-fv2-filter}. The filter stage implementation with
+these optimizations runs 10 times faster and is more pronounced on higher
+ratios.
 
@@
 
-In order to be able to make comparisons on field offsets, a simple approach is
-to linearly walk through each filtered record against the filtered recordset
-leading to a complexity of $O(n^2)$, where $n$ is the number of filtered
-records. A smarter approach is to put the copy in a hash table and then try to
-map each pointer while walking down the filtered recordset once, leading to a
-complexity of $O(n)$. The hash table approach, although will work on this
-specific example, will fail badly on other relative comparisons. A better
-approach, as implemented in F$(v2)$ is to sort the filtered recordset on all
-the requested grouping rules. This helps the execution engine perform a
-nested \texttt{bsearch} to reduce the linear pass to a fairly small filtered
-recordset. However, the number of elements upon which the search has to be
-performed needs to be known at each level of the binary lookup. In order to
-avoid a linear run to count the number of elements, the parent level
-\texttt{bsearch} invocation returns a boundary with the first and last element
-to enable such a calculation. This helps the grouper perform faster search
-lookups to find records that must group together in $O(n*lg(k))$ time with a
-preprocessing step taking $O(n*lg(n)) + O(n)$ in the average case, where $n$
-is the number of filtered records and $k$ is the number of unique filtered
-records.
+In order to be able to make relative comparisons on field offsets, a simple
+approach is to linearly walk through each filtered record against the filtered
+recordset leading to a complexity of $O(n^2)$, where $n$ is the number of
+filtered records. A smarter approach is to put the copy in a hash table and
+then try to map each pointer while walking down the filtered recordset once,
+leading to a complexity of $O(n)$. The hash table approach however, will only
+work on equality comparisons. A better approach, as implemented in \ac{NFQL}
+is to sort the filtered recordset on all the requested grouping terms of a
+clause. This helps the execution engine perform a nested binary search to
+reduce the linear pass to a fairly small filtered recordset. However, the
+number of elements upon which the search has to be performed needs to be known
+at each level of the binary lookup. In order to avoid a linear run to count
+the number of elements, the parent level binary search invocation returns a
+boundary with the first and last element to enable such a calculation. This
+helps the grouper perform faster search lookups to find records that must
+group together in $O(n*lg(k))$ time with a preprocessing step taking
+$O(p*n*lg(n))$ in the average case, where $n$ is the number of filtered
+records, $p$ is the number of grouping terms in a clause and $k$ is the number
+of unique filtered records.
 
@@
     \includegraphics*[width=1.0\linewidth]{grouper-fv2-bsearch-fv2}
-    \caption{Grouper Stage: F$(v2)$ (Generic) vs F$(v2)$ (Specific)}
+    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
     \label{fig:fv1-fv2-grouper}
@@
 records and a  subsequent binary search goes away.  The groups can now be
-formed in $O(n)$ time with a more involved preprocessing step taking
-$O(p*n*lg(n)$ where is $n$ is the number of filtered records, and $p$ is the
-number of grouping rules. The performance evaluation of the grouper handling
-this special case against its behavior when handling generic cases is shown in
-Fig. \ref{fig:fv1-fv2-grouper}. Profiling the performance of the grouper on
-queries that produce higher ratios reveal that a significant amount of time is
-taken in \texttt{bsearch}. The special case of equality comparisons eliminate
-these calls and further reduce the time in the long tail as shown in Fig.
-\ref{fig:fv1-fv2-grouper}.
+formed in $O(n)$ time with the same preprocessing step taking $O(p*n*lg(n)$.
+The performance evaluation of the grouper handling this special case against
+its behavior when handling generic cases is shown in Fig.
+\ref{fig:fv1-fv2-grouper}. Profiling the performance of the grouper on queries
+that produce higher ratios reveal that a significant amount of time is taken
+in binary search which is eliminated in the special case of equality
+comparisons.
 
-\subsubsection{Group Aggregations} Group records are a conglomeration of
-several flow records with some common characteristics defined by the flow
-query. Some of the non-common characteristics can also be aggregated into a
-single value using group aggregations. It is useful if a single group record
-can be again mapped into a NetFlow $v5$ record template, so that it can be
-written to file as a representative of all its members.
+The resultant group records are a conglomeration of multiple flow records with
+some common characteristics. Some of the non-common characteristics can also
+be aggregated into a single value using group aggregations as defined in the
+query. Such an aggregated group record is again mapped to a NetFlow $v5$
+record template. This allows the aggregated group records to be written to
+file as a representative of all its members.
 
-\subsubsection{Group Filter} The groupfilter is used to filter the groups
-produced by the grouper based on some absolute rules defined in a \ac{DNF}
-expression. The \texttt{struct} term holds information about the flow record
-offset, the value being compared to and the operator which maps to a unique
-\texttt{enum} value.  This enum value is used to map the operation to a
+\subsubsection{Group Filter} The groupfilter stage is used to filter the
+groups produced by the grouper based on some absolute rules defined in a
+\ac{DNF} expression. The \texttt{struct} term holds information about the flow
+record offset, the value being compared to and the operator which maps to a
+unique \texttt{enum} value.  This enum value is used to map the operation to a
 specific group-filter function. The group-filter functions are auto-generated
-using a python script.
+using a Python script.
 
@@
 branches that need to be spawned is read from the query and is not known until
-compile time. As a result, an iterator that can provide all possible
-permutations of $m-$tuple (where $m$ is the number of branches) group record
-$ID$s was needed.  The result of the iterator can then be used to make a
-match. The merger stage, begins by initializing this iterator passing it the
-number of branches, and information about each branch. Then, it loops over to
-get a new $m-$tuple of group record $ID$s on each iteration until the iterator
-returns \texttt{false}.
+runtime. As a result, an iterator that can provide all possible permutations
+of $m-$tuple (where $m$ is the number of branches) group record $ID$s was
+needed.  The result of the iterator can then be used to make a match. The
+merger stage, begins by initializing this iterator passing it the number of
+branches, and information about each branch. Then, it loops over to get a new
+$m-$tuple of group record $ID$s on each iteration until the iterator returns
+\texttt{false}.
 
@@
 record from one branch with every other record of each branch. This leads to a
-complexity of $O(n^m)$ where $n$ is the number of filtered group records and
+complexity of $O(g^m)$ where $g$ is the number of filtered group records and
 $m$ is the number of branches. The possible number of tries when matching

### 2012-09-02 sections/implementation/flow-query-intermediate-format.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

--- 2012-08-31 sections/implementation/flow-query-intermediate-format.tex
@@
-A \ac{DNF} is a disjunction of conjunctive clauses. The elements of the
-conjuctive clauses are terms. Each stage of the pipeline is represented in the
-flowquery as a \ac{DNF} expression. The clauses in the \ac{DNF} are OR'd
-together, while the terms in each clauses themselves are AND'd. The abstract
-objects that store the JSON query and the results that incubate from each
-stage are designed to be self-descriptive and hierarchically chainable. The
-complete JSON query information for instance, is held within the
-\texttt{flowquery} struct. Each individual branch of the flowquery itself is
-described in a \texttt{branch} struct. A collection of these \texttt{branch}
-structs are referenced in the parent \texttt{flowquery} struct.
+The entire JSON flow query is a collection of \ac{DNF} expressions. A \ac{DNF}
+expression is a disjunction of conjunctive clauses. The elements of the
+conjuctive clauses are terms. The clauses in the \ac{DNF} are OR'd together,
+while the terms in each clause themselves are AND'd. The branchsets and each
+\ac{DNF} expression of the pipeline stage is a \texttt{JSON array}.
+\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
+pipeline stages of the flow query language are also encapsulated in Python
+classes and scripts have been written to facilitate the front-end parser to
+serialize these pipeline objects to a \texttt{JSON} flow query.  The mapping
+of the query to the structs defined in the execution engine is
+not trivial. When reading the \texttt{JSON} query at runtime, the field
+offsets of the NetFlow $v5$ record struct are read in as strings.  Utility
+functions are defined that map the read field names to struct offsets and
+field types and the operations to unique enum members.
+
+The abstract objects that store the JSON query and the results that incubate
+from each stage are designed to be self-descriptive and hierarchically
+chainable.  The complete JSON query information for instance, is held within
+the \texttt{flowquery} struct as shown in listing \ref{lst:struct}. Each
+individual branch of the flowquery itself is described in a \texttt{branch}
+struct.
+
+\lstset{caption=Flow Query and Branch Structs,
+				tabsize=2, language=C, numbers=left,stepnumber=1,
+        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
+        keywordstyle=\color{blue}, xleftmargin=15pt,
+        rulesepcolor=\color{black}, label=lst:struct,
+        aboveskip=10pt, captionpos=b}
+\begin{lstlisting}
+struct flowquery {
+  size_t                                      num_branches;
+  size_t                                      num_merger_clauses;
+
+  struct branch**                             branchset;
+  struct merger_clause**                      merger_clauseset;
+  struct merger_result*                       merger_result;
+  struct ungrouper_result*                    ungrouper_result;
+};
+
+struct branch {
+  int                                         branch_id;
+  struct ftio*                                ftio_out;
+  struct ft_data*                             data;
+
+  size_t                                      num_filter_clauses;
+  size_t                                      num_grouper_clauses;
+  size_t                                      num_aggr_clause_terms;
+  size_t                                      num_groupfilter_clauses;
+
+  struct filter_clause**                      filter_clauseset;
+  struct grouper_clause**                     grouper_clauseset;
+  struct aggr_term**                          aggr_clause_termset;
+  struct groupfilter_clause**                 groupfilter_clauseset;
+
+  struct filter_result*                       filter_result;
+  struct grouper_result*                      grouper_result;
+  struct groupfilter_result*                  gfilter_result;
+};
+\end{lstlisting}
+
+The \texttt{JSON} query can also trigger and disable the stages at runtime.
+This means that one only has to supply the constructs that one wishes to use.
+The constructs that are not defined in the \texttt{JSON} query are inferred by
+the execution engine as a disable request. The execution engine uses disable
+flags that are turned on when the query is parsed. These flags are used
+throughout the engine to only enable the requested functionality.

### 2012-09-02 sections/implementation/performance-optimizations.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

@@
-There can be a situtation where the query designer may incorrectly ask for
+There can be a situation where the query designer may incorrectly ask for
 aggregation on a field already specified in a grouper (or filter) clause. If
@@
 \ac{NFQL} pushes the filter stage out of the branch and closer to where the
-trace is originally read. This enables the memory allocation of only records
-that passed the filter stage thereby reducing the runtime memory footprint of
-the execution engine.
-
-Each branch runs in its own thread. If any of the stages of the branch return
-a \texttt{NULL} when returning from their public interface function, there is
-no reason to continue the thread. The subsequent stages of the branch cannot
-do much with a \texttt{NULL} result.  Therefore, the branch thread returns
-with a \texttt{EXIT\_FAILURE} if either stage returns \texttt{NULL}, and with
-\texttt{EXIT\_SUCCESS} on normal exit.
+trace is originally read. This reduces the memory allocation to only records
+that passed the filter stage thereby reducing the memory footprint of the
+execution engine.
 
 Each stage of the processing pipeline is dependent on the result of the
-previous one. As a result, the stages should only proceed and process, when
-the previous returned results. Implementing such a response was
-straightforward for the grouper and group filter, the merger although was a
-little trickier.  The merger stage proceeds only when every branch has
-non-zero filtered groups.  The iterator initializer
-\texttt{iter\_init(\ldots)} deallocates and returns \texttt{NULL} if any one
-branch has $0$ filtered groups.  Consequently a check is performed in the
-merger to make sure \texttt{iter} is \emph{not} \texttt{NULL}.
+previous one. As a result, the stages should only proceed, when the previous
+returned results. Implementing such a response was straightforward for the
+grouper and group filter, the merger although was a little trickier.  The
+merger stage proceeds only when every branch has non-zero filtered groups.
+The iterator initializer \texttt{iter\_init(\ldots)} deallocates and returns
+\texttt{NULL} if any one branch has $0$ filtered groups.  Consequently a check
+is performed in the merger to make sure \texttt{iter} is \emph{not}
+\texttt{NULL}.
 

### 2012-09-03 sections/relatedwork.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations

@@
-The study of the network traffic patterns began with tools that utilized raw
-packet captures at the monitoring point for traffic measurement.
-\texttt{tcpdump} and \texttt{wireshark} have since then become the de-facto
-tools for packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage}
-is a premier command-line utility that uses the \texttt{libpcap}
-\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
-comes from the richness of its expressions, the ability to combine them using
-logical connectives and extract specific portions of a packet using filters.
-\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
-both journeymen and packet analysis experts. It supports a large number of
-protocols, has a straightforward layout, excellent documentation, and can run
-on all major operating systems.
+%The study of the network traffic patterns began with tools that utilized raw
+%packet captures at the monitoring point for traffic measurement.
+%\texttt{tcpdump} and \texttt{wireshark} have since then become the de-facto
+%tools for packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage}
+%is a premier command-line utility that uses the \texttt{libpcap}
+%\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
+%comes from the richness of its expressions, the ability to combine them using
+%logical connectives and extract specific portions of a packet using filters.
+%\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
+%both journeymen and packet analysis experts. It supports a large number of
+%protocols, has a straightforward layout, excellent documentation, and can run
+%on all major operating systems.
 

### 2012-09-04 sections/implementation/execution-workflow.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)

@@
-A custom C library has been written to directly read/write data in the
-\texttt{flow-tools} format. The library sequentially reads the complete
-flow-records into memory to support random access required for relative
-filtering. Each flow-record is stored in a \texttt{char} array and the offsets
-to each field are stored in separate \texttt{structs} as shown in listing
+A custom C library has been written to directly read/write data stored in
+\texttt{flow-tools} format. The library sequentially reads the flow-records
+into memory to support random access required for relative filtering. Each
+flow-record is stored in a \texttt{char} array and the offsets to each field
+are stored in separate \texttt{structs} as shown in listing
 \ref{lst:c-lib-struct}. The array of such records are indexed allowing fast
@@
 \subsubsection{Splitter} \ac{NFQL} uses identifiers to reference a flow record
-in the \texttt{char} array. This gives away the need to every time copy all
-the flow-records when moving ahead in the processing pipeline. As a result,
-there is no dedicated splitter stage in the execution engine. Each branch
-references the flow records from a common memory location. This helps keep the
-memory costs at a minimum when multiple branches are involved.
+in the \texttt{char} array. This eliminated the need to copy all the
+flow-records when moving ahead in the processing pipeline. As a result, there
+is no dedicated splitter stage in the execution engine. Each branch references
+the flow records from a common memory location. This helps keep the memory
+costs at a minimum when multiple branches are involved.
 
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{filter-fv1-fv2}
+    \includegraphics* [width=0.9\linewidth]{filter-fv1-fv2}
     \caption{Filter Stage: NFQL vs NFQL (optimized)}
@@
 make a decision on whether or not to keep a flow record; it has to pass
-through the whole in-memory recordset \emph{again} to fill in the filter
+through the whole in-memory recordset \emph{again} to produce the filter
 results. This technique involves multiple linear runs on the trace and
@@
 reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
-records in the trace. We used the publically available Trace 7 from the
-SimpleWeb \cite{simpleweb} to compare the performance of the optimized
-\ac{NFQL} implementation against the one defined by the flow query language as
-shown in Fig. \ref{fig:fv1-fv2-filter}. The filter stage implementation with
-these optimizations runs 10 times faster and is more pronounced on higher
-ratios.
+records in the trace. The filtered records are saved in a common location from
+where they are referenced by each branch. This helps keep the memory costs at
+a minimum when multiple branches are involved.  We used the publically
+available Trace 7 from the SimpleWeb \cite{simpleweb} to compare the
+performance of the optimized \ac{NFQL} implementation against the one defined
+by the flow query language as shown in Fig.  \ref{fig:fv1-fv2-filter}. The
+filter stage implementation with these optimizations runs 10 times faster and
+is more pronounced on higher ratios.
 
@@
 work on equality comparisons. A better approach, as implemented in \ac{NFQL}
-is to sort the filtered recordset on all the requested grouping terms of a
-clause. This helps the execution engine perform a nested binary search to
-reduce the linear pass to a fairly small filtered recordset. However, the
-number of elements upon which the search has to be performed needs to be known
-at each level of the binary lookup. In order to avoid a linear run to count
-the number of elements, the parent level binary search invocation returns a
-boundary with the first and last element to enable such a calculation. This
-helps the grouper perform faster search lookups to find records that must
-group together in $O(n*lg(k))$ time with a preprocessing step taking
-$O(p*n*lg(n))$ in the average case, where $n$ is the number of filtered
-records, $p$ is the number of grouping terms in a clause and $k$ is the number
-of unique filtered records.
+is to sort the filtered recordset on the field offsets all the requested
+grouping terms of a clause. This helps the execution engine perform a nested
+binary search to reduce the linear pass to a fairly small filtered recordset.
+However, the number of elements upon which the search has to be performed
+needs to be known at each level of the binary lookup. In order to avoid a
+linear run to count the number of elements, the parent level binary search
+invocation returns a boundary with the first and last element to enable such a
+calculation. This helps the grouper perform faster search lookups to find
+records that must group together in $O(n*lg(k))$ time with a preprocessing
+step taking $O(p*n*lg(n))$ in the average case, where $n$ is the number of
+filtered records, $p$ is the number of grouping terms in a clause and $k$ is
+the number of unique filtered records.
 
@@
   \begin{center}
-    \includegraphics*[width=1.0\linewidth]{grouper-fv2-bsearch-fv2}
+    \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
     \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
@@
 records and a  subsequent binary search goes away.  The groups can now be
-formed in $O(n)$ time with the same preprocessing step taking $O(p*n*lg(n)$.
+formed in $O(n)$ time with the same preprocessing step taking $O(p*n*lg(n))$.
 The performance evaluation of the grouper handling this special case against
@@
 query. Such an aggregated group record is again mapped to a NetFlow $v5$
-record template. This allows the aggregated group records to be written to
+record template. This allows the aggregated group records to be written to a
 file as a representative of all its members.
@@
 needed.  The result of the iterator can then be used to make a match. The
-merger stage, begins by initializing this iterator passing it the number of
+merger stage begins by initializing this iterator passing it the number of
 branches, and information about each branch. Then, it loops over to get a new
@@
 $m$ is the number of branches. The possible number of tries when matching
-group records however can be reduced by sorting the group records on the field
-offsets used for a match. The \ac{NFQL} implementation optimizes the merger to
-skip over iterator permutations when a state of a current field offset value
-may not allow any further match beyond the index in the current branch.  For
-such an optimization to work, the filtered group records must be sorted in the
-order of field offsets specified in the merger clause. Specifying the filtered
-group records in any other order may lead to undefined behavior. This means,
-that if the same field offsets were used in the grouper stage, the terms in
-the group clause can be rearranged by the query designer to align with the
-order of terms in the merger clause.
+group records, however, can be reduced by sorting the group records on the
+field offsets used for a match. The \ac{NFQL} implementation optimizes the
+merger to skip over iterator permutations when a state of a current field
+offset value may not allow any further match beyond the index in the current
+branch.  For such an optimization to work, the filtered group records must be
+sorted in the order of field offsets specified in the merger clause.
+Specifying the filtered group records in any other order may lead to undefined
+behavior. This means, that if the same field offsets were used in the grouper
+stage, the terms in the group clause can be rearranged by the query designer
+to align with the order of terms in the merger clause.
 

### 2012-09-04 sections/implementation/flow-query-intermediate-format.tex
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)

@@
-The entire JSON flow query is a collection of \ac{DNF} expressions. A \ac{DNF}
-expression is a disjunction of conjunctive clauses. The elements of the
-conjuctive clauses are terms. The clauses in the \ac{DNF} are OR'd together,
-while the terms in each clause themselves are AND'd. The branchsets and each
-\ac{DNF} expression of the pipeline stage is a \texttt{JSON array}.
-\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
-pipeline stages of the flow query language are also encapsulated in Python
-classes and scripts have been written to facilitate the front-end parser to
-serialize these pipeline objects to a \texttt{JSON} flow query.  The mapping
-of the query to the structs defined in the execution engine is
-not trivial. When reading the \texttt{JSON} query at runtime, the field
-offsets of the NetFlow $v5$ record struct are read in as strings.  Utility
-functions are defined that map the read field names to struct offsets and
-field types and the operations to unique enum members.
+Each stage of the processing pipeline is expressed in the \texttt{JSON} query
+as a \ac{DNF} expression. A \ac{DNF} expression is a disjunction of
+conjunctive clauses. The elements of the conjuctive clauses are terms. The
+clauses in the \ac{DNF} are OR'd together, while the terms in each clause
+themselves are AND'd. \texttt{json-c} \cite{jsonc} is used to parse the flow
+query file. The pipeline stages of the flow query language are also
+encapsulated in Python classes and scripts have been written to facilitate the
+front-end parser to serialize these pipeline objects to a \texttt{JSON} flow
+query.  The mapping of the query to the structs defined in the execution
+engine is not trivial. When reading the \texttt{JSON} query at runtime, the
+field names of a NetFlow $v5$ record are read in as strings.  Utility
+functions are defined that map the field names to internal struct offsets and
+the field types and the operations to internal unique enum members.
 
@@
 
-The \texttt{JSON} query can also trigger and disable the stages at runtime.
-This means that one only has to supply the constructs that one wishes to use.
-The constructs that are not defined in the \texttt{JSON} query are inferred by
-the execution engine as a disable request. The execution engine uses disable
-flags that are turned on when the query is parsed. These flags are used
-throughout the engine to only enable the requested functionality.
+The \texttt{JSON} query can also disable the stages at runtime.  This means
+that one only has to supply the constructs that one wishes to use.  The
+constructs that are not defined in the \texttt{JSON} query are inferred by the
+execution engine as a disable request. The execution engine uses disable flags
+that are turned on when the query is parsed. These flags are used throughout
+the engine to only enable the requested functionality.

### 2012-09-04 sections/introduction.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)

@@
 
-Flow capture today, has emerged out to be one of the favored network
-measurement techniques. In this technique, packets traversing a monitoring
-point are not captured raw, instead they are aggregated together based on some
-common characteristics. The common characteristics are learnt by inspecting
-the packet headers as they cross the monitoring point. Flow-records resulting
-from such an aggregation are then exported to a collector for further
-analysis. This not only reduces the amount of traffic at the monitoring point,
-but also provides fine-grained control over the network data which was not
-previously possible using SNMP interface-level queries.
+Capturing network flow records has emerged out to be one of the favored
+network measurement techniques. In this technique, packets traversing an
+obvervation point are not captured raw, instead they are aggregated together
+based on some common characteristics. The common characteristics are learnt by
+inspecting the packet headers as they cross the observation point.
+Flow-records resulting from such an aggregation are then exported to a
+collector for further analysis. This reduces the amount of traffic aggregated
+at the collector.
 
-NetFlow and \ac{IPFIX} are the two popular standards of IP flow information
+NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
 export. NetFlow \cite{rfc3954} is a proprietary network protocol designed by
-Cisco Systems. It allow routers to generate and export flow records to a
+Cisco Systems. It allows routers to generate and export flow records to a
 designated collector. NetFlow v$9$ provides flexibility of user-tailored
 export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
-\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF deemed
-to be the logical successor of NetFlow v$9$ on which it is based. The novelty
-of the standard lies in its ability to describe record formats at runtime
-using templates based on an extensible and well-defined information model. The
-data transfer mechanism is also simplistic and extensible by being
+\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF and
+deemed to be the logical successor of NetFlow v$9$ on which it is based. The
+novelty of the standard lies in its ability to describe record formats at
+runtime using templates based on an extensible and well-defined information
+model. The data transfer mechanism is also simplistic and extensible by being
 unidirectional and transport protocol agnostic.
@@
 
-It goes to show that understanding these intricate traffic patterns require
-sophisticated flow analysis tools that can mine flow records for such a usage.
-Unfortunately current tools fail to deliver owing to their poor language
-design and simplistic filtering methods.  We recently proposed a flow query
-language design \cite{vmarinov:thesis:2009} that aims to cater to such needs.
-It can process flow records, aggregate them into groups, apply absolute (or
-relative) filters and invoke Allen interval algebra rules \cite{fallen:1983}.
+Understanding intricate traffic patterns require sophisticated flow analysis
+tools that can mine flow records for such a usage.  Unfortunately current
+tools fail to deliver owing to their language design and simplistic filtering
+methods.  We recently proposed a flow query language design
+\cite{vmarinov:2009} that aims to cater to such needs.  It can process
+flow records, aggregate them into groups, apply absolute (or relative) filters
+and invoke Allen interval algebra rules \cite{fallen:1983}. The applicability
+of the language can be seen from \cite{vperelman:2011} where the authors
+formulate flow queries to identify flow signatures of popular applications.
 
@@
 implementation, Flowy \cite{kkanev:2010}, which was written in Python.
-\ac{NFQL} however is not just a reimplementation in a new language, but
-the inner workings have been reimagined to allow the execution engine to
-scale to real-world sized traces. This has been possible by replacing the
-performance hit stages of the pipeline with crispier algorithms that lower
-down the execution times to factor making them comparable to contemporary
-flow analysis tools.
+\ac{NFQL}, however, is not just a reimplementation in a new language, but the
+inner workings have been reimagined to allow the execution engine to scale to
+real-world sized traces. This has been possible by replacing the performance
+critical stages of the pipeline with crispier algorithms that lower the
+execution times, making \ac{NFQL} comparable to contemporary flow analysis
+tools.
 
-The rest of the paper is organized as follows. In section \ref{sec:design} we
-describe the flow query language by discussing each stage of the processing
-pipeline. In section \ref{sec:implementation} we introduce the inner workings
-of \ac{NFQL}. It begins by providing an overview of the \ac{NFQL} components
-and the structure of the intermediate format used to exchange messages. The
-workflow of the execution engine is described next and is supplemented by a
-number of performance optimizations made to make the implementation scale. It
-is followed by performance evaluation results by comparing them against
-contemporary flow processing tools in section \ref{sec:evaluation} and
-concludes with a discussion on its current limitation and future outlook in
-section \ref{sec:conclusion}.
+The paper is organized as follows. In section \ref{sec:relatedwork} we survey
+the current state-of-the-art flow-processing tools and reason how \ac{NFQL} is
+different from each one of them. In section \ref{sec:design} we describe the
+flow query language by discussing each stage of the processing pipeline. In
+section \ref{sec:implementation} we introduce the inner workings of \ac{NFQL}.
+It begins by providing an overview of the \ac{NFQL} architecture and the
+structure of the intermediate format used to exchange messages. The workflow
+of the execution engine is described next and is supplemented by a number of
+performance optimizations made to make the implementation scale. The section
+is followed by performance evaluations comparing \ac{NFQL} against
+contemporary flow processing tools alongwith with a discussion on its current
+limitation and future outlook in section \ref{sec:evaluation}, with conclusion
+in section \ref{sec:conclusion}.
 
@@
 \centering
-\includegraphics[width=1.0\linewidth]{nfql-pipeline}
+\includegraphics[width=0.9\linewidth]{nfql-pipeline}
 \caption{NFQL Processing Pipeline \cite{vmarinov:2009}}

### 2012-09-04 sections/evaluation.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * performance optimizations

@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{filter}
+    \includegraphics* [width=0.9\linewidth]{filter}
     \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
@@
 is because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
-that are read and written used \texttt{lzo} compression scheme which trades
-space for achieving faster compression and decompression. It is important to
-note, that all the tools were single-threaded in this evaluation, and did not
-completely utilize the $24$ available cores. It comes as a realization, that
-adding \texttt{lzo} compression will drastically improve \ac{NFQL}'s filter
-performance.
+that are read and written used \texttt{lzo} compression scheme. It is
+important to note, that all the tools were single-threaded in this evaluation,
+and did not completely utilize the $24$ available cores. It comes as a
+realization, that adding \texttt{lzo} compression will drastically improve
+\ac{NFQL}'s filter performance.
 
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{grouper}
+    \includegraphics* [width=0.9\linewidth]{grouper}
     \caption{Grouper Stage: NFQL vs SiLK}
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{groupfilter}
+    \includegraphics* [width=0.9\linewidth]{groupfilter}
     \caption{Group Filter Stage: NFQL vs SiLK}
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{merger}
+    \includegraphics* [width=0.9\linewidth]{merger}
     \caption{Merger Stage}
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{ungrouper}
+    \includegraphics* [width=0.9\linewidth]{ungrouper}
     \caption{Ungrouper Stage}

### 2012-09-04 sections/implementation/adaptable-compression-levels.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)
* revision with professor's comments   * performance optimizations

@@
-The flow-records echoed to the standard output can also be written to a
-Netflow $v5$ \texttt{flow-tools} file. The \texttt{--dirpath} switch allows
-one to provide a directory path where the results can be stored.  Each stream
-is stored as its own file with an ID to disambiguate it. Results from each
+The flow-records echoed to the standard output can also be written to Netflow
+$v5$ \texttt{flow-tools} files. The \texttt{-{}-dirpath} switch allows one to
+provide a directory path where the results can be stored. Results from each
 stage of the pipeline can also be written to separate files with the increase
-in the verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose}
+in the verbosity level. In fact, \texttt{-{}-dirpath} and \texttt{-{}-verbose}
 work well together to adjust the level at which the writes are to be made.
@@
   \begin{center}
-    \includegraphics* [width=1.0\linewidth]{zlevel}
-    \caption{F$(v2)$: \texttt{z-level} Effect on Performance}
+    \includegraphics* [width=0.9\linewidth]{zlevel}
+    \caption{\texttt{z-level} Effect on Performance}
     \label{fig:engine-zlevel}
@@
 $9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
-supports \texttt{--zlevel} switch to allow the user to supply its desired
+supports \texttt{-{}-zlevel} switch to allow the user to supply its desired
 choice of the of the compression level. A default level of $5$ is used for
@@
 passing the filter stage for each \texttt{z-level}. It goes to show that each
-level adds its own performance overhead and must be used with discretion.
+level adds its own performance overhead and must be used with discretion. It
+is also important to note that other tools may use different compression
+algorithms. \texttt{nfdump} for instance, uses \texttt{lzo} \cite{lzo}
+compression to trade space for faster compression and decompression.

### 2012-09-04 sections/implementation/performance-optimizations.tex
* revision with professor's comments   * performance optimizations

@@
 because there is nothing more for the next stage to compute. The function
-pointers are therefore set as late as possible and are are called from their
+pointers are therefore set as late as possible and are called from their
 respective stages just before the comparison is needed.  As a result, we save
@@
 
-\ac{NFQL} pushes the filter stage out of the branch and closer to where the
+\ac{NFQL} pushes the filter stage out of the branches and closer to where the
 trace is originally read. This reduces the memory allocation to only records
-that passed the filter stage thereby reducing the memory footprint of the
+that passed the filter stage, thereby reducing the memory footprint of the
 execution engine.
@@
 echo the results, however echoes to the standard output are only used for
-debugging purposes. The writes to a file can however be legitimately requested
-in a real network analysis task. As a result, it is essential to avoid
-additional loops when performing writes to a file. The execution engine,
+debugging purposes. The writes to a file can, however, be legitimately
+requested in a real network analysis task. As a result, it is essential to
+avoid additional loops when performing writes to a file. The execution engine
 therefore writes each result record to a file as soon as it is seen by the

### 2012-09-04 sections/evaluation.tex
* revision with professor's comments   * performance evaluations

@@
-We used a public flow trace containing $20M$ records to run a number of flow
-queries using our benchmarking suite. We used a publically available Trace 7
-from the SimpleWeb \cite{simpleweb} repository for all the performance
+We used the first $20M$ records from a public flow trace to run a number of
+flow queries using our benchmarking suite. We used the publically available
+Trace 7 from the SimpleWeb \cite{simpleweb} repository for all the performance
 evaluations. The input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using
@@
 compatible formats and compressed with \texttt{zlib} keeping the same
-compression level. The suite was run on a high-end machine with $24$ cores of
+compression level. The suite was run on a machine with $24$ cores of
 $2.5$ GHz clock speed and $18$ GiB of memory.
@@
 lower ratios, but then again SilK and \texttt{nfdump} also use their own file
-format. As a result, the amount of data that needs to read (or written) may be
-different to what it is for \ac{NFQL} and \texttt{flowtools}.  On the other
-hand, \texttt{nfdump} appears to be significantly faster than the rest. This
-is because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
-that are read and written used \texttt{lzo} compression scheme. It is
+format. As a result, the amount of data that needs to be read (or written) may
+be different to what it is for \ac{NFQL} and \texttt{flowtools}.
+\texttt{nfdump} appears to be significantly faster than the rest. This is
+because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
+that are read and written use the \texttt{lzo} compression scheme. It is
 important to note, that all the tools were single-threaded in this evaluation,
@@
 apart. Since most of the time is taken in writing the records to files, it is
-unclear whether SiLK's usage of a proprietary format, which may reduce
-reads/writes is responsible for the drift on higher ratios. SiLK's
-\texttt{rwgroup} tool is also supplied a \texttt{--summarize} flag to force it
+unclear whether SiLK's usage of its own file format, which may reduce
+reads/writes, is responsible for the drift on higher ratios. SiLK's
+\texttt{rwgroup} tool is also supplied a \texttt{-{}-summarize} flag to force it
 to write only the first record of each group, to make both tools write the
@@
 ratio is not available since the \ac{NFQL} query executed at that data point
-uses \texttt{OR} expressions which are not supported by SiLK. As a result, an
+uses \texttt{OR} expressions, which are not supported by SiLK. As a result, an
 equivalent SiLK query is not formulated.
@@
 implementation takes advantage of sorted nature of filtered groups and
-therefore can signifantly reduce the number of merger matches. It also writes
-a merged group record to file only once despite the number of times it has
+therefore can significantly reduce the number of merger matches. It also writes
+a merged group record to a file only once despite the number of times it has
 matched. Without these optimizations, running such queries on the merger would
@@
 
-It can be seen that the evaluation behavior of the ungrouper is similar to
-that shown by the merger. Since, the ungrouper also has to write the entire
-merged groupset to file, as is previously done by merger, the execution engine
-takes twice the amount of time. It is important to note, that SiLK does not
-have such an equivalent ungrouping tool is therefore not considered in
-this final evaluation.
+It can be seen that the behavior of the ungrouper is similar to that shown by
+the merger. Since, the ungrouper also has to write the entire merged groupset
+to files, as is previously done by merger, the execution engine takes twice the
+amount of time. It is important to note, that SiLK does not have such an
+equivalent ungrouping tool and is therefore not considered in this final
+evaluation.

### 2012-09-04 sections/conclusion.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
revision with professor's comments   * conclusion, with blurb on usefulness of benchmarking queries   * removed ungrouper graph

@@
 query language. The language allows applying absolute (or relative) filters,
-aggregating flows into groups, evaluating intricate timing relationships among
-them, and merging them into one collection. \ac{NFQL} can execute such complex
-query tasks in matter of minutes, thereby expanding the scope of current flow
-record processing tools. The conducted performance evaluations reveal that
-\ac{NFQL} is on par with tools that support only absolute filters. SiLK, the
-only openly available package that provides tools that are similar to the rest
-of \ac{NFQL}'s processing pipeline appears faster. This is because it can
+aggregating flows into groups, evaluating timing relationships among them, and
+merging them into one collection. \ac{NFQL} can execute such complex queries
+in matter of minutes, thereby expanding the scope of current flow record
+processing tools. The conducted performance evaluations reveal that \ac{NFQL}
+is on par with tools that support only absolute filters. SiLK, the only openly
+available package that provides tools that are similar to the rest of
+\ac{NFQL}'s processing pipeline appears faster. This is because it can
 optimize its operations in favor of the limited set of comparisons that are
-only based on equality, and its usage of a different file storage format.
-\todo{...}
+only based on equality, and its usage of a different file storage format. The
+evaluation queries developed as a part of this research work may also develop
+into a more general benchmark suite for flow query tools and platforms.

### 2012-09-04 sections/evaluation.tex
revision with professor's comments   * conclusion, with blurb on usefulness of benchmarking queries   * removed ungrouper graph

@@
 rewriting the merged groups as \texttt{flowtools} files using the same
-\texttt{zlib} compression level. The ratio of the number of result flow
-records to twice the number of the flow records in the input trace is plotted
-against time. The evaluation results are shown in Fig.
-\ref{fig:benchmarks-ungrouper}.
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics* [width=0.9\linewidth]{ungrouper}
-    \caption{Ungrouper Stage}
-    \label{fig:benchmarks-ungrouper}
-  \end{center}
-\end{figure}
-
-It can be seen that the behavior of the ungrouper is similar to that shown by
-the merger. Since, the ungrouper also has to write the entire merged groupset
-to files, as is previously done by merger, the execution engine takes twice the
-amount of time. It is important to note, that SiLK does not have such an
-equivalent ungrouping tool and is therefore not considered in this final
-evaluation.
+\texttt{zlib} compression level. This means that the execution engine end up
+taking twice the amount of time than the merger. It is important to note, that
+SiLK does not have such an equivalent ungrouping tool and is therefore not
+considered in this final evaluation.

### 2012-09-04 sections/conclusion.tex
* revision with professor's comments   * conclusion with a footnote to nfql source code (BSD licensed)

@@
-We presented \ac{NFQL}, an efficient C implementation of the stream-based flow
-query language. The language allows applying absolute (or relative) filters,
-aggregating flows into groups, evaluating timing relationships among them, and
-merging them into one collection. \ac{NFQL} can execute such complex queries
-in matter of minutes, thereby expanding the scope of current flow record
-processing tools. The conducted performance evaluations reveal that \ac{NFQL}
-is on par with tools that support only absolute filters. SiLK, the only openly
-available package that provides tools that are similar to the rest of
-\ac{NFQL}'s processing pipeline appears faster. This is because it can
-optimize its operations in favor of the limited set of comparisons that are
-only based on equality, and its usage of a different file storage format. The
-evaluation queries developed as a part of this research work may also develop
-into a more general benchmark suite for flow query tools and platforms.
+We presented
+\ac{NFQL}\footnote{\texttt{\url{https://github.com/vbajpai/nfql}}}, an
+efficient C implementation of the stream-based flow query language. The
+language allows applying absolute (or relative) filters, aggregating flows
+into groups, evaluating timing relationships among them, and merging them into
+one collection. \ac{NFQL} can execute such complex queries in matter of
+minutes, thereby expanding the scope of current flow record processing tools.
+The conducted performance evaluations reveal that \ac{NFQL} is on par with
+tools that support only absolute filters. SiLK, the only openly available
+package that provides tools that are similar to the rest of \ac{NFQL}'s
+processing pipeline appears faster. This is because it can optimize its
+operations in favor of the limited set of comparisons that are only based on
+equality, and its usage of a different file storage format. The evaluation
+queries developed as a part of this research work may also develop into a more
+general benchmark suite for flow query tools and platforms.

### 2012-09-04 sections/introduction.tex
Rewrote the "usage scenarios" paragraph
 * added references to surveys given by the professor
 * removed old version

@@
 
-The wide applicability of this approach is easily seen from the pervasive use
-of flow records for a vibrant set of network analysis applications. For
-instance, the authors in \cite{mkim:2004} use the flow characteristics in the
-traffic pattern to formalize a detection function that maps traffic patterns
-to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
-use the flow-record data to exploit timing characteristics of webmail clients
-to classify features that could identify webmail traffic from any other
-traffic running over HTTPS.
+%The wide applicability of this approach is easily seen from the pervasive use
+%of flow records for a vibrant set of network analysis applications. For
+%instance, the authors in \cite{mkim:2004} use the flow characteristics in the
+%traffic pattern to formalize a detection function that maps traffic patterns
+%to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
+%use the flow-record data to exploit timing characteristics of webmail clients
+%to classify features that could identify webmail traffic from any other
+%traffic running over HTTPS.
+
+A survey by Sperotto \textit{et al.} \cite{sperotto:2010} gives an overview of
+how network flow analysis can be used to detect Denial of Service attacks,
+scans probing target systems and the target discovery phase of worms. In
+addition, a survey by Callado \textit{et al.} \cite{callado:2009} also lists
+behaviour analysis of Internet backbone traffic and general anomaly detection.
 

### 2012-09-04 sections/introduction.tex
* added survery paper introduction blurb * filter comparison graph changed to flowy 2.0 vs nfql

@@
 
-A survey by Sperotto \textit{et al.} \cite{sperotto:2010} gives an overview of
-how network flow analysis can be used to detect Denial of Service attacks,
-scans probing target systems and the target discovery phase of worms. In
-addition, a survey by Callado \textit{et al.} \cite{callado:2009} also lists
-behaviour analysis of Internet backbone traffic and general anomaly detection.
+The wide applicability of this approach is easily seen from the pervasive use
+of flow records for a vibrant set of network analysis applications. For
+instance, a survey by Sperotto \textit{et al.} \cite{sperotto:2010} gives an
+overview of how network flow analysis can be used to detect Denial of Service
+attacks, scan probing target systems and the target discovery phase of worms.
+In addition, a survey by Callado \textit{et al.} \cite{callado:2009} also
+lists behaviour analysis of Internet backbone traffic and general anomaly
+detection.
 
@@
 flow records, aggregate them into groups, apply absolute (or relative) filters
-and invoke Allen interval algebra rules \cite{fallen:1983}. The applicability
+and invoke Allen interval algebra rules \cite{fallen:1983}. The expressiveness
 of the language can be seen from \cite{vperelman:2011} where the authors

### 2012-09-04 sections/introduction.tex
Add flowy/NFQL history

@@
 
-In this paper, we introduce \ac{NFQL}, an efficient C implementation of the
-flow query language. It is the next iteration of the first prototype
-implementation, Flowy \cite{kkanev:2010}, which was written in Python.
-\ac{NFQL}, however, is not just a reimplementation in a new language, but the
-inner workings have been reimagined to allow the execution engine to scale to
-real-world sized traces. This has been possible by replacing the performance
-critical stages of the pipeline with crispier algorithms that lower the
-execution times, making \ac{NFQL} comparable to contemporary flow analysis
-tools.
+Flowy \cite{kkanev:2010} was a first feature complete Python prototype of that
+flow query language. Due to performance problems, it was superseeded by a
+complete rewrite in C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this
+paper, we introduce \ac{NFQL} which extends on Flowy 2.0, making it more
+feature complete and optimizing its execution engine with crispier algorithms.
+We show that this iteration of our work is able to scale to real-world sized
+traces and has comparable execution times to contemporary flow analysis tools.
 
@@
 It begins by providing an overview of the \ac{NFQL} architecture and the
-structure of the intermediate format used to exchange messages. The workflow
-of the execution engine is described next and is supplemented by a number of
-performance optimizations made to make the implementation scale. The section
-is followed by performance evaluations comparing \ac{NFQL} against
-contemporary flow processing tools alongwith with a discussion on its current
-limitation and future outlook in section \ref{sec:evaluation}, with conclusion
-in section \ref{sec:conclusion}.
+structure of the intermediate format used to exchange messages. The workflow of
+the execution engine is described next and is supplemented by a number of
+performance optimizations made to make the implementation scale. The section is
+followed by performance evaluations comparing \ac{NFQL} against contemporary
+flow processing tools alongwith with a discussion on its current limitation and
+future outlook in section \ref{sec:evaluation}, with conclusion in section
+\ref{sec:conclusion}.
 

### 2012-09-04 sections/relatedwork.tex
* bibliographic entries   * ntop, flowscan, nfsen
* revision with professor's comments   * related work comments

@@
 traffic as flow records and use them for network analysis. \texttt{flow-tools}
-and \texttt{nfdump} are one of the most popular tools used for analyzing
+and \texttt{nfdump} are among the most popular tools used for analyzing
 NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite of programs
@@
 full-fledged though. The grouping and merging operations can only be performed
-using an equality operator. This is assumed in the tool, thereby allowing it
-to perform optimization such as using hash tables to perform lookups. There
-are also stringent requirements to how flow-data needs to be organized before
-it can be piped into a tool. The grouping tool, for instance, assumes that the
-to-be supplied input flow data is already sorted on the field column.  These
-requirements make it a little cumbersome to design a full-fledged flow query.
-For instance, trying to mimic a \ac{NFQL} query in SiLK ends up as a bash
-script with over a dozen of SiLK tools piped together.
+using an equality operator. This restriction allows the tool to perform
+optimization such as using hash tables to perform lookups. There are also
+stringent requirements to how flow-data needs to be organized before it can be
+piped into a tool. The grouping tool, for instance, assumes that the to-be
+supplied input flow data is already sorted on the field column.  These
+requirements can make it a little cumbersome to design a full-fledged flow
+query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes ends
+up as a shell script with over a dozen of SiLK tools piped together.

### 2012-09-04 sections/design.tex
* revision with professor's comments   * flow query language design

@@
-The flow query language pipeline consists of a number of independent elements
-that are connected to one another using UNIX-based pipes.  Each element
-receives the content from the previous pipe, performs an operation and pushes
-it to the next element in the pipeline as shown in Fig.
-\ref{fig:nfql-pipeline}. A complete description on the semantics of each
-element in the pipeline can be found in \cite{vmarinov:2009}
+The flow query language consists of a number of independent stages that are
+connected to one another to form a processing pipeline.  Each stage receives
+the content from the previous, performs an operation and pushes the results to
+the next stage as shown in Fig. \ref{fig:nfql-pipeline}. A complete
+description of the semantics of each pipeline stage can be found in
+\cite{vmarinov:2009}.
 
 The pipeline starts with the splitter. It takes the flow-records data as input
-in the \texttt{flow-tools} compatible format. The splitter is responsible to
+in \texttt{flow-tools} compatible format. The splitter is responsible to
 duplicate the input data out to several branches without any processing
-whatsoever. This allows each of the branch to receive an identical copy of the
-flow data to process it independently.
+whatsoever. This allows each of the branches to receive an identical copy of
+the flow data to process it independently.
 
-The filter is the first processing element of a branch in the pipeline.  It
+The filter is the first processing stage of a branch in the pipeline.  It
 performs \emph{absolute} filtering on the input flow-records data.  The
@@
 field of the \emph{same} flow-record. The filter cannot \emph{relatively}
-compare two different incoming flow-records
+compare two different incoming flow-records.
 
 The grouper performs aggregation of the input flow-records data. It consists
-of a number of terms that correspond to a specific subgroup. A flow-record in
-order to be a part of the group should satisfy at-least one subgroup. A
-flow-record can be a part of multiple subgroups within a group. A flow-record
-cannot be part of multiple groups.  The grouping rules can be either absolute
-or relative. The newly formed groups which are passed on to the group filter
-can also contain meta-information about the flow-records contained within the
-group using the aggregate clause defined as part of the grouper query.
+of a number of terms that correspond to a specific clause. A flow-record in
+order to be a part of the group should satisfy at-least one clause. A
+flow-record, although, can be a part of multiple clauses, it cannot be part of
+multiple groups. The grouping terms can be either absolute or relative. The
+newly formed groups, which are passed on to the group filter, can also contain
+meta-information about the flow-records contained within the group using the
+aggregate clause defined as part of the grouper query.
 
-The group-filter is the last processing element of the branch. It performs
-\emph{absolute} filtering on the input group-records data. The group-records
+The group-filter is the last processing stage of a branch. It performs
+\emph{absolute} filtering on group-records. The group-records
 that pass the filtering criterion are forwarded to the merger, the rest of the
 group-records are dropped. The group-filter compares separate fields (or
-aggregated fields) of a flow-record against either a constant value or a value
-on a different field of the \emph{same} flow-record. The group-filter cannot
+aggregated fields) of a group-record against either a constant value or a value
+on a different field of the \emph{same} group-record. The group-filter cannot
 \emph{relatively} compare two different incoming group-records.
@@
 the final processing element of the pipeline. It unwraps the tuples of
-group-records into individual flow-records, ordered by their timestamps. The
+group-records into individual flow-records, ordered by their timestamps. Any
 duplicate flow-records appearing from several group-records are eliminated and

### 2012-09-04 sections/conclusion.tex
* (or relative) -> or relative

@@
 efficient C implementation of the stream-based flow query language. The
-language allows applying absolute (or relative) filters, aggregating flows
+language allows applying absolute or relative filters, aggregating flows
 into groups, evaluating timing relationships among them, and merging them into

### 2012-09-04 sections/introduction.tex
* (or relative) -> or relative

@@
 \cite{vmarinov:2009} that aims to cater to such needs.  It can process
-flow records, aggregate them into groups, apply absolute (or relative) filters
+flow records, aggregate them into groups, apply absolute or relative filters
 and invoke Allen interval algebra rules \cite{fallen:1983}. The expressiveness

### 2012-09-04 sections/relatedwork.tex
added graphical analysis tools to related work

@@
 In recent years, a number of tools have been developed that can capture the
-traffic as flow records and use them for network analysis. \texttt{flow-tools}
-and \texttt{nfdump} are among the most popular tools used for analyzing
-NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite of programs
-for capturing and processing NetFlow v5 flow records. It consists of 24
-separate tools that work together by connecting them via UNIX pipes. It can
-capture, read, filter, and print flow records internally saved in a fixed-size
-format. \texttt{nfdump} \cite{phaag:2006} is a very similar tool that uses a
-different storage format. Flow records are captured using \texttt{nfcapd} and
-then processed by \texttt{nfdump}. The power of filtering rules in both the
-tools is however mostly limited to absolute comparisons of flow attributes. As
-a result, relative comparison amongst different flows or querying a timing
-relationship among them is not possible.  SiLK \cite{SiLK} is a network
-traffic collection and analysis tool developed and maintained by the CERT
-Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon University.
-SiLK is the tool that comes quite close to providing similar capabilities as
-provided by \ac{NFQL} and is therefore used as a reference point to compare
-the performance of the \ac{NFQL} execution engine in this paper. The design
-and implementation of SiLK, however, differs a lot from that of \ac{NFQL}. For
-instance, in SiLK there are separate tools to perform the task of each stage
-of the \ac{NFQL} processing pipeline. The stage functionality is not
-full-fledged though. The grouping and merging operations can only be performed
-using an equality operator. This restriction allows the tool to perform
-optimization such as using hash tables to perform lookups. There are also
-stringent requirements to how flow-data needs to be organized before it can be
-piped into a tool. The grouping tool, for instance, assumes that the to-be
-supplied input flow data is already sorted on the field column.  These
-requirements can make it a little cumbersome to design a full-fledged flow
-query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes ends
-up as a shell script with over a dozen of SiLK tools piped together.
+traffic as flow records and use them for network analysis.
+
+Simple analysis of network traffic can be done by a range of graphical
+utilities like ntop \cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen
+\cite{phaag:2006} and Stager \cite{oslebo:2006}. All these tools understand the
+NetFlow format while ntop and Stager can also process \ac{IPFIX}. They use a
+round robin database to store flow information, except for Stager which uses a
+MySQL database. Graphical presentation is done via a webinterface or in the
+case of FlowScan via RRDTool.
+
+\texttt{flow-tools} and \texttt{nfdump} are among the most popular tools used
+for analyzing NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite
+of programs for capturing and processing NetFlow v5 flow records. It consists
+of 24 separate tools that work together by connecting them via UNIX pipes. It
+can capture, read, filter, and print flow records internally saved in a
+fixed-size format.  \texttt{nfdump} \cite{phaag:2006} is a very similar tool
+that uses a different storage format. Flow records are captured using
+\texttt{nfcapd} and then processed by \texttt{nfdump}. The power of filtering
+rules in both the tools is however mostly limited to absolute comparisons of
+flow attributes. As a result, relative comparison amongst different flows or
+querying a timing relationship among them is not possible.  SiLK \cite{SiLK} is
+a network traffic collection and analysis tool developed and maintained by the
+CERT Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon
+University.  SiLK is the tool that comes quite close to providing similar
+capabilities as provided by \ac{NFQL} and is therefore used as a reference
+point to compare the performance of the \ac{NFQL} execution engine in this
+paper. The design and implementation of SiLK, however, differs a lot from that
+of \ac{NFQL}. For instance, in SiLK there are separate tools to perform the
+task of each stage of the \ac{NFQL} processing pipeline. The stage
+functionality is not full-fledged though. The grouping and merging operations
+can only be performed using an equality operator. This restriction allows the
+tool to perform optimization such as using hash tables to perform lookups.
+There are also stringent requirements to how flow-data needs to be organized
+before it can be piped into a tool. The grouping tool, for instance, assumes
+that the to-be supplied input flow data is already sorted on the field column.
+These requirements can make it a little cumbersome to design a full-fledged
+flow query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes
+ends up as a shell script with over a dozen of SiLK tools piped together.

### 2012-09-04 sections/relatedwork.tex
* penultimate PDF

@@
 In recent years, a number of tools have been developed that can capture the
-traffic as flow records and use them for network analysis.
-
-Simple analysis of network traffic can be done by a range of graphical
-utilities like ntop \cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen
-\cite{phaag:2006} and Stager \cite{oslebo:2006}. All these tools understand the
-NetFlow format while ntop and Stager can also process \ac{IPFIX}. They use a
-round robin database to store flow information, except for Stager which uses a
-MySQL database. Graphical presentation is done via a webinterface or in the
-case of FlowScan via RRDTool.
+traffic as flow records and use them for network analysis.Simple analysis of
+network traffic can be done by a range of graphical utilities like ntop
+\cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen \cite{phaag:2006} and
+Stager \cite{oslebo:2006}. All these tools understand the NetFlow format while
+ntop and Stager can also process \ac{IPFIX}. They use a round robin database
+to store flow information, except for Stager which uses a MySQL database.
+Graphical presentation is done via a webinterface or in the case of FlowScan
+via RRDTool.
 
@@
 flow attributes. As a result, relative comparison amongst different flows or
-querying a timing relationship among them is not possible.  SiLK \cite{SiLK} is
-a network traffic collection and analysis tool developed and maintained by the
-CERT Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon
-University.  SiLK is the tool that comes quite close to providing similar
-capabilities as provided by \ac{NFQL} and is therefore used as a reference
-point to compare the performance of the \ac{NFQL} execution engine in this
-paper. The design and implementation of SiLK, however, differs a lot from that
-of \ac{NFQL}. For instance, in SiLK there are separate tools to perform the
-task of each stage of the \ac{NFQL} processing pipeline. The stage
-functionality is not full-fledged though. The grouping and merging operations
-can only be performed using an equality operator. This restriction allows the
-tool to perform optimization such as using hash tables to perform lookups.
-There are also stringent requirements to how flow-data needs to be organized
-before it can be piped into a tool. The grouping tool, for instance, assumes
-that the to-be supplied input flow data is already sorted on the field column.
-These requirements can make it a little cumbersome to design a full-fledged
-flow query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes
-ends up as a shell script with over a dozen of SiLK tools piped together.
+querying a timing relationship among them is not possible.
+
+SiLK \cite{SiLK} is a network traffic collection and analysis tool developed
+and maintained by the CERT Network Situational Awareness Team (CERT NetSA) at
+Carnegie Mellon University.  SiLK is the tool that comes quite close to
+providing similar capabilities as provided by \ac{NFQL} and is therefore used
+as a reference point to compare the performance of the \ac{NFQL} execution
+engine in this paper. The design and implementation of SiLK, however, differs
+a lot from that of \ac{NFQL}. For instance, in SiLK there are separate tools
+to perform the task of each stage of the \ac{NFQL} processing pipeline. The
+stage functionality is not full-fledged though. The grouping and merging
+operations can only be performed using an equality operator. \ac{NFQL} on the
+other hand, provides a much richer set of comparisons operations, such as
+\emph{equal, not equal, greater than, less than, greater or equal, less or
+equal}. This restriction allows the tool to perform optimization such as using
+hash tables to perform lookups.  There are also stringent requirements to how
+flow-data needs to be organized before it can be piped into a tool. The
+grouping tool, for instance, assumes that the to-be supplied input flow data
+is already sorted on the field column.  These requirements can make it a
+little cumbersome to design a full-fledged flow query.  For instance, trying
+to mimic a \ac{NFQL} query in SiLK sometimes ends up as a shell script with
+over a dozen of SiLK tools piped together.

### 2012-09-04 sections/implementation/execution-workflow.tex
caption consistency with graph legends
penultimate, corrections

@@
 
-A sample JSON query \ac{DNF} term definition is shown in listing
-\ref{lst:filter-struct}. In order to be able to make comparisons on the field
-offsets of such a term, the comparator needs to know the type of the
-comparison and the length of the field offset. This information is parsed by
-execution engine once the query is read and is therefore not available at
-compile time.  In order to subvert the need to define complex branching
-statements, a dedicated comparator is defined for every possible field length
-and comparison operation. These comparators are generated using a Python
-script. This allows the term definitions to make runtime calls using a
-function name derived from the combination of operation type and field length.
+In order to be able to make comparisons on the field offsets of such a term,
+the comparator needs to know the type of the comparison and the length of the
+field offset. This information is parsed by execution engine once the query is
+read and is therefore not available at compile time.  In order to subvert the
+need to define complex branching statements, a dedicated comparator is defined
+for every possible field length and comparison operation. A Python script
+generates C source code for these comparators at compile time conforming to
+the structure shown in listing \ref{filter-struct}. This allows the term
+definitions to make runtime calls using a function name derived from the
+combination of operation type and field length.
 
@@
     \includegraphics* [width=0.9\linewidth]{filter-fv1-fv2}
-    \caption{Filter Stage: NFQL vs NFQL (optimized)}
+    \caption{Filter Stage: Flowy 2.0 vs NFQL}
     \label{fig:fv1-fv2-filter}
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \begin{center}

### 2012-09-04 sections/implementation/flow-query-intermediate-format.tex
penultimate, corrections

@@
-Each stage of the processing pipeline is expressed in the \texttt{JSON} query
-as a \ac{DNF} expression. A \ac{DNF} expression is a disjunction of
-conjunctive clauses. The elements of the conjuctive clauses are terms. The
-clauses in the \ac{DNF} are OR'd together, while the terms in each clause
-themselves are AND'd. \texttt{json-c} \cite{jsonc} is used to parse the flow
-query file. The pipeline stages of the flow query language are also
-encapsulated in Python classes and scripts have been written to facilitate the
-front-end parser to serialize these pipeline objects to a \texttt{JSON} flow
-query.  The mapping of the query to the structs defined in the execution
-engine is not trivial. When reading the \texttt{JSON} query at runtime, the
-field names of a NetFlow $v5$ record are read in as strings.  Utility
-functions are defined that map the field names to internal struct offsets and
-the field types and the operations to internal unique enum members.
+Each stage of the processing pipeline is expressed in the JSON query as a
+\ac{DNF} expression. The elements of the conjuctive clauses are terms.
+\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
+pipeline stages of the flow query language are also encapsulated in Python
+classes and scripts have been written to facilitate the front-end parser to
+serialize these pipeline objects to a JSON flow query.  The mapping
+of the query to the structs defined in the execution engine is not trivial.
+When reading the JSON query at runtime, the field names of a NetFlow
+$v5$ record are read in as strings.  Utility functions are defined that map
+the field names to internal struct offsets and the field types and the
+operations to internal unique enum members.
 
@@
 
-The \texttt{JSON} query can also disable the stages at runtime.  This means
+The JSON query can also disable the stages at runtime.  This means
 that one only has to supply the constructs that one wishes to use.  The
-constructs that are not defined in the \texttt{JSON} query are inferred by the
+constructs that are not defined in the JSON query are inferred by the
 execution engine as a disable request. The execution engine uses disable flags

### 2012-09-04 sections/introduction.tex
penultimate, corrections

@@
 performance evaluation and overall improvement. Traffic measurement techniques
-that have been rapidly evolving in the last decade, have matured enough today
+that have been rapidly evolving in the last decade have today matured enough
 to provide such an insight.
@@
 network measurement techniques. In this technique, packets traversing an
-obvervation point are not captured raw, instead they are aggregated together
-based on some common characteristics. The common characteristics are learnt by
-inspecting the packet headers as they cross the observation point.
-Flow-records resulting from such an aggregation are then exported to a
+obvervation point are not captured raw, instead common characteristics of the
+packet header are aggregated into flow records. They are then exported to a
 collector for further analysis.  This reduces the amount of traffic aggregated
@@
 NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
-export. NetFlow \cite{rfc3954} is a proprietary network protocol designed by
-Cisco Systems. It allows routers to generate and export flow records to a
-designated collector. NetFlow v$9$ provides flexibility of user-tailored
+export. NetFlow \cite{rfc3954} is a network protocol designed by Cisco
+Systems, which allows routers to generate and export flow records to a
+designated collector. NetFlow v$9$ provides flexibility through user-tailored
 export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
-\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF and
-deemed to be the logical successor of NetFlow v$9$ on which it is based. The
-novelty of the standard lies in its ability to describe record formats at
-runtime using templates based on an extensible and well-defined information
-model. The data transfer mechanism is also simplistic and extensible by being
-unidirectional and transport protocol agnostic.
-
-%The wide applicability of this approach is easily seen from the pervasive use
-%of flow records for a vibrant set of network analysis applications. For
-%instance, the authors in \cite{mkim:2004} use the flow characteristics in the
-%traffic pattern to formalize a detection function that maps traffic patterns
-%to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
-%use the flow-record data to exploit timing characteristics of webmail clients
-%to classify features that could identify webmail traffic from any other
-%traffic running over HTTPS.
+\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF based
+on NetFlow v$9$. The novelty of the standard lies in its ability to describe
+record formats at runtime using templates based on an extensible and
+well-defined information model. The data transfer mechanism is also simplistic
+and extensible by being unidirectional and transport protocol agnostic.
 

### 2012-09-04 sections/relatedwork.tex
penultimate, corrections

--- 2012-09-04 sections/relatedwork.tex
@@
-%The study of the network traffic patterns began with tools that utilized raw
-%packet captures at the monitoring point for traffic measurement.
-%\texttt{tcpdump} and \texttt{wireshark} have since then become the de-facto
-%tools for packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage}
-%is a premier command-line utility that uses the \texttt{libpcap}
-%\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
-%comes from the richness of its expressions, the ability to combine them using
-%logical connectives and extract specific portions of a packet using filters.
-%\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
-%both journeymen and packet analysis experts. It supports a large number of
-%protocols, has a straightforward layout, excellent documentation, and can run
-%on all major operating systems.
-
 In recent years, a number of tools have been developed that can capture the
-traffic as flow records and use them for network analysis.
-
-Simple analysis of network traffic can be done by a range of graphical
-utilities like ntop \cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen
-\cite{phaag:2006} and Stager \cite{oslebo:2006}. All these tools understand the
-NetFlow format while ntop and Stager can also process \ac{IPFIX}. They use a
-round robin database to store flow information, except for Stager which uses a
-MySQL database. Graphical presentation is done via a webinterface or in the
-case of FlowScan via RRDTool.
+traffic as flow records and use them for network analysis. Simple analysis of
+network traffic can be done by a range of graphical utilities like ntop
+\cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen \cite{phaag:2006} and
+Stager \cite{oslebo:2006}. All these tools understand the NetFlow format while
+ntop and Stager can also process \ac{IPFIX}. They use a round robin database
+to store flow information, except for Stager which uses a MySQL database.
+Graphical presentation is done via a webinterface or in the case of FlowScan
+via RRDTool.
 
@@
 flow attributes. As a result, relative comparison amongst different flows or
-querying a timing relationship among them is not possible.  SiLK \cite{SiLK} is
-a network traffic collection and analysis tool developed and maintained by the
-CERT Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon
-University.  SiLK is the tool that comes quite close to providing similar
-capabilities as provided by \ac{NFQL} and is therefore used as a reference
-point to compare the performance of the \ac{NFQL} execution engine in this
-paper. The design and implementation of SiLK, however, differs a lot from that
-of \ac{NFQL}. For instance, in SiLK there are separate tools to perform the
-task of each stage of the \ac{NFQL} processing pipeline. The stage
+querying a timing relationship among them is not possible.
+
+SiLK \cite{SiLK} is a network traffic collection and analysis tool developed
+and maintained by the CERT Network Situational Awareness Team (CERT NetSA) at
+Carnegie Mellon University.  SiLK is the tool that comes quite close to
+providing similar capabilities as \ac{NFQL} and is therefore used as a
+reference point to compare the performance of the \ac{NFQL} execution engine
+in this paper. The design and implementation of SiLK, however, differs a lot
+from that of \ac{NFQL}. For instance, in SiLK there are separate tools to
+perform the task of each stage of the \ac{NFQL} processing pipeline. The stage
 functionality is not full-fledged though. The grouping and merging operations
@@
 tool to perform optimization such as using hash tables to perform lookups.
-There are also stringent requirements to how flow-data needs to be organized
-before it can be piped into a tool. The grouping tool, for instance, assumes
-that the to-be supplied input flow data is already sorted on the field column.
-These requirements can make it a little cumbersome to design a full-fledged
-flow query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes
-ends up as a shell script with over a dozen of SiLK tools piped together.
+\ac{NFQL} on the other hand, provides a much richer set of comparisons
+operations, such as \emph{equal, not equal, greater than, less than, greater
+or equal, less or equal}. There are also stringent requirements to how
+flow-data needs to be organized before it can be piped into a tool. The
+grouping tool, for instance, assumes that the to-be supplied input flow data
+is already sorted on the field column.  These requirements can make it a
+little cumbersome to design a full-fledged flow query.  For instance, trying
+to mimic a \ac{NFQL} query in SiLK sometimes ends up as a shell script with
+over a dozen of SiLK tools piped together.

### 2012-09-04 sections/implementation/execution-workflow.tex
moved acronym section, fixed filter-struct reference
fix two spelling errors

@@
 generates C source code for these comparators at compile time conforming to
-the structure shown in listing \ref{filter-struct}. This allows the term
+the structure shown in listing \ref{lst:filter-struct}. This allows the term
 definitions to make runtime calls using a function name derived from the
@@
 where they are referenced by each branch. This helps keep the memory costs at
-a minimum when multiple branches are involved.  We used the publically
+a minimum when multiple branches are involved.  We used the publicly
 available Trace 7 from the SimpleWeb \cite{simpleweb} to compare the
@@
 The merger is used to relate groups from different branches according to a
-merging criterean. The implementation is not trivial since the number of
+merging criterion. The implementation is not trivial since the number of
 branches that need to be spawned is read from the query and is not known until

### 2012-09-04 sections/design.tex
* penultimate PDF
penultimate, corrections
the listings don't break into multiple pages

@@
 
-The grouper performs aggregation of the input flow-records data. It consists
-of a number of terms that correspond to a specific clause. A flow-record in
+The grouper performs aggregation of the input flow-records. A flow-record in
 order to be a part of the group should satisfy at-least one clause. A
@@
 meta-information about the flow-records contained within the group using the
-aggregate clause defined as part of the grouper query.
+aggregate clause defined as part of the grouper query. The possible
+aggregation operations are \emph{static, count, product, sum, logical
+and/or/xor, arithmetic mean, standard deviation, union, median, minimum and
+maximum}.
 
@@
 filtering on the groups are applied to express timing and concurrency
-constraints using Allen interval algebra \cite{fallen:1983}. The ungrouper is
-the final processing element of the pipeline. It unwraps the tuples of
-group-records into individual flow-records, ordered by their timestamps. Any
-duplicate flow-records appearing from several group-records are eliminated and
-are sent as output only once.
+constraints using Allen interval algebra \cite{fallen:1983}.
+
+The ungrouper is the final processing element of the pipeline. It unwraps the
+tuples of group-records into individual flow-records, ordered by their
+timestamps. Any duplicate flow-records appearing from several group-records
+are eliminated and are sent as output only once.

### 2012-10-15 rebuttal.txt
added rebuttal questions (need to be responded)

Review 1:
- wrong track
- lack of description of the language
- too many implementation details
- metrics used
- research questions and conclusions

Review 2:
- Gigascope, point to the original NFQL paper

Review 3:
- Explain that NFQL is defined elsewhere, this paper is discussing
  implementation aspects.
- Discuss storage requirement vs. state requirement (after all NFQL
  is a streaming language but it requires state to be maintained)

Review 4:
- Explain what NFQL can do that other tools can't do
- Clarify the order splitter/filter in the query language and the
  order of filter/splitter in the implementation

### 2012-10-15 rebuttal.txt
make some initial changes to rebuttal.txt

@@
 Review 1:
-- wrong track
-- lack of description of the language
-- too many implementation details
-- metrics used
-- research questions and conclusions
+1. wrong track
+2. lack of description of the language
+3. too many implementation details
+4. metrics used
+5. research questions and conclusions
 
 Review 2:
-- Gigascope, point to the original NFQL paper
+1. Gigascope, point to the original NFQL paper
 
 Review 3:
-- Explain that NFQL is defined elsewhere, this paper is discussing
+1. Explain that NFQL is defined elsewhere, this paper is discussing
    implementation aspects.
-- Discuss storage requirement vs. state requirement (after all NFQL
+2. Discuss storage requirement vs. state requirement (after all NFQL
    is a streaming language but it requires state to be maintained)
@@
 Review 4:
-- Explain what NFQL can do that other tools can't do
-- Clarify the order splitter/filter in the query language and the
+1. Explain what NFQL can do that other tools can't do
+2. Clarify the order splitter/filter in the query language and the
    order of filter/splitter in the implementation
+
+JoSch's notes/comments
+======================
+
+Review 1
+--------
+
+Review 2
+--------
+
+Review 3
+--------
+
+Review 4
+--------
+
+1. The reviewer notes that the work is not state of the art. I think it is
+mentioned a couple of times (even in the abstract) what flowy can do that
+other's cant and in "related work" it is listed how it has even more abilities
+than SiLK.

### 2012-10-16 rebuttal.txt
* rebuttal: review 4     * splitter and filter order     * filtering flows on the fly     * array indexes     * state of the art: performance and features

@@
 Review 4:
-1. Explain what NFQL can do that other tools can't do
-2. Clarify the order splitter/filter in the query language and the
-   order of filter/splitter in the implementation
 
-JoSch's notes/comments
-======================
+- can you explain why you need the splitter before the filtering blocks
+and not the other way round? IMHO filtering should happen before
+splitting (fig 1)
 
-Review 1
---------
+The the order of execution of the pipeline stages in the flow query
+language [1] and in the NFQL implementation is different.  According to
+the language specification, as briefly described in section III:
 
-Review 2
---------
+"The pipeline starts with the splitter. It takes the flow-records data
+as input in flow-tools compatible format. The splitter is responsible to
+duplicate the input data out to several branches without any processing
+whatsoever. This allows each of the branches to receive an identical
+copy of the flow data to process it independently."
 
-Review 3
---------
+However, we (and as did the reviewer) realized that in order to
+efficiently process large traces, the NFQL implementation needs to sway
+away from this design decision. As such, NFQL is implemented in a way to
+give away the need for a splitter entirely, as described in section IV-B
 
-Review 4
---------
+"NFQL uses identifiers to reference a flow record in the char array.
+This eliminated the need to copy all the flow-records when moving ahead
+in the processing pipeline. As a result, there is no dedicated splitter
+stage in the execution engine. Each branch references the flow records
+from a common memory location. This helps keep the memory costs at a
+minimum when multiple branches are involved."
 
-1. The reviewer notes that the work is not state of the art. I think it is
-mentioned a couple of times (even in the abstract) what flowy can do that
-other's cant and in "related work" it is listed how it has even more abilities
-than SiLK.
+As a result, there is no splitter in the NFQL implementation.
+
+[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
+Record Query Language,” DSOM ’09.Berlin: Springer-Verlag, 2009, pp. 15–28
+
+- can you explain why you do not filter flows on the fly?
+
+We _do_ filter flows on the fly. This is described in section IV-B
+
+"The execution engine, as defined by the flow query language must read
+all the flow records of a supplied trace into memory before starting the
+processing pipeline. Since, the filter stage uses the supplied set of
+absolute rules to make a decision on whether or not to keep a flow
+record; it has to pass through the whole in-memory recordset again to
+produce the filter results. This technique involves multiple linear runs
+on the trace and therefore slows down when the ratio of number of
+filtered records to the total number of flow-records is high. We
+optimized this behavior in NFQL by merging the filter stage with
+in-memory read of the trace. This means, a decision on whether or not to
+make room for a record in memory and eventually hold a pointer for it in
+filter results is done upfront as soon as the record is read from the
+trace. In addition, if a request to write the filter stage results to a
+flow-tools file has been made, the writes are also made as soon the
+filter stage decision is available, thereby allowing
+reading-filtering-writing to happen in O(n) time, where n is the number
+of records in the trace."
+
+The previous prototype implementations (Flowy and Flowy 2.0) had this
+issue. There was a separate filter stage and the flows were not filtered
+on the fly. NFQL however _does_ filter flows on the fly. We even
+evaluated the performance improvements of NFQL against Flowy 2.0 and saw
+that NFQL now runs 10 times faster as shown in Fig 3.
+
+- in IV.B you talk about indexes. What indexes have you used. Please cover
+this topic better
+
+Each flow record is stored as a char array. The trace is stored as an
+array of these flow records. This allows NFQL to reference any flow
+record using identifiers (indexes) in the array for O(1) retrieval.
+
+- The work described on this paper is interesting but probably not the
+state of the art in terms of performance and features (i.e. it's not a
+step ahead but more a student project). Some examples are not easy to
+understand.
+
+We have tried to set forth the novelty of NFQL from the very outset,
+The abstract for instance, clearly briefs out the functionality that
+separates it from contemporary flow-processing tools:
+
+"NFQL can process flow records, aggregate them into groups, apply
+absolute or relative filters, invoke Allen interval algebra rules, and
+merge group records."
+
+In section II (related work), we reason how NFQL does more than the
+surveyed contemporary flow-processing tools:
+
+flow-tools and nfdump:
+"The power of filtering rules in both the tools is however mostly
+limited to absolute comparisons of flow attributes. As a result,
+relative comparison amongst different flows or querying a timing
+relationship among them is not possible." - which is possible in NFQL.
+In fact, the capabilities of flow-tools and nfdump are just one of the
+five stages of the NFQL's processing pipeline. They do not support
+grouping, aggregation, merging et al.
+
+SiLK: 
+SiLK can group flows and merge group records. However, the parameters
+over which grouping/merging can be done is faily limited:
+"The grouping and merging operations can only be performed using
+an equality operator. NFQL on the other hand, provides a much richer set
+of comparisons operations, such as equal, not equal, greater than, less
+than, greater or equal, less or equal."
+
+"The newly formed groups, which are passed on to the group filter, can
+also contain meta-information about the flow-records contained within
+the group using the aggregate clause defined as part of the grouper
+query. The possible aggregation operations are static, count, product,
+sum, logical and/or/xor, arithmetic mean, standard deviation, union,
+median, minimum and maximum." - whereas SiLK only supports sum and or
+aggregation operations.
+
+"The relative filtering on the groups are applied to express timing and
+concurrency constraints using Allen interval algebra [6]." - The merger
+can also match group records by putting timing characteristics under
+considerations. No such Allen operations are possible in SiLK.
+
+"The ungrouper is the final processing element of the pipeline. It
+unwraps the tuples of group-records into individual flow-records,
+ordered by their timestamps." - NFQL maintains flow membership information of
+each group record. SiLK does not provide this feature. As a result, once
+flows are grouped together in SiLK, there is no way to unfold the groups
+to identify their member flows. As such, there is no ungrouper in SiLK.
+
+In section V, we evaluate the performance of each tool against NFQL.
+flow-tools and nfdump could only participate in one of the evaluations
+due to their limited feature sets. In rest of the evaluations only SiLK
+could participate. Even so, we had to carefully design queries to find a
+common ground since most of the NFQL operations are not possible in SiLK.
+
+We believe NFQL is major step ahead in this effort. Our first python
+prototype Flowy was feature complete, but was extremely slow in
+comparison to contemporary tools and therefore could not scale to
+real-world traces.  NFQL is an efficient feature complete implementation
+of our flow-query language. It is not just a reimplementation in C, but
+the processing pipeline has been reimagined and optimized with crispier
+data structures to allow it to scale to real-world traces. This makes
+NFQL comparable to current flow-processing tools, in addition to
+offering more feature-sets that are just not available elsewhere.
+
+It is not clear from the review which examples are not easy to
+understand, so it's hard to comment.

### 2012-10-16 rebuttal.txt
rebuttal: review 3   * design advantages discussed elsewhere.   * paper discusses implementation aspects, not derivative.   * state vs storage requirement (stream-based)

@@
 Review 3:
-1. Explain that NFQL is defined elsewhere, this paper is discussing
-   implementation aspects.
-2. Discuss storage requirement vs. state requirement (after all NFQL
-   is a streaming language but it requires state to be maintained)
+
+- design advantages of the proposed tool are not explained, if this is
+  because NFQL has already been proposed in another paper, then this
+  work is derivative to a high degree
+- again, because the proposed method (not tool) is not explained, and
+  explanation of related tools lacks functional and methodological
+  detail, it is difficult to judge WHY the proposed tool is better
+
+The flow query language was introduced in [1]. It describes the _design_
+advantages of the language. In fact, the expressiveness of the language
+can be seen from [2] where we define queries to identify flow signatures
+of popular applications. We have had a python prototype implementation
+[3] in the past, however the prototype was extremely slow and could not
+scale to real-world traces. This paper introduces NFQL, an efficient
+implementation of the language. This is not just a reimplementation in
+C, but the pipeline has been reimagined with crispier data structures to
+allow the implementation to scale to real-world traces. This is the
+first time we evaluate and show that NFQL is comparable to contemporary
+flow-processing tools. This paper discusses these implementation aspects.
+
+[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
+Record Query Language,” DSOM ’09
+[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
+Popular Applications,” IM '11.
+[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
+stream-based IP flow record query language,” AIMS’10
+
+- the method is a -QL -- a query language, which assumes fixed
+  storage, while recent research on traffic processing moves away from
+  the store-then-process approach
+
+The query language of NFQL is _indeed_ stream-based. The processing
+pipeline consists of a number of elements, which are connected with each
+other via pipes. Each element gets a input _stream_, performs some sort
+of operation on it and the output _stream_ is piped to the next element.
+This only requires that NFQL maintains state information between the
+elements. NFQL does not necessarily follow a store-then-process approach
 
@@
 of our flow-query language. It is not just a reimplementation in C, but
-the processing pipeline has been reimagined and optimized with crispier
-data structures to allow it to scale to real-world traces. This makes
-NFQL comparable to current flow-processing tools, in addition to
-offering more feature-sets that are just not available elsewhere.
+the processing pipeline has been reimagined to allow it to scale to
+real-world traces. This makes NFQL comparable to current flow-processing
+tools, in addition to offering more feature-sets that are just not
+available elsewhere.
 

### 2012-10-16 rebuttal.txt
some comments about stream-based

@@
 elements. NFQL does not necessarily follow a store-then-process approach
+
+# i do not think this is what the reviewer understands as "stream-based"
+# nfql is not able to process packages as they come. Because it allows
+# to do relative comparison in the grouper and the merger, those stages
+# need to have access to any record of the dataset
+# an extension that could make nfql stream-based for the grouper is a
+# shifting window which only evaluates a small neighbourhood of the
+# packet that is currently evaluated
+# but this idea fails for the merger
+#  - josch
 

### 2012-10-16 rebuttal.txt
rebuttal: review 2   * this paper is a leap over prior work   * reference to gigascope

--- 2012-10-16 rebuttal.txt
@@
 Review 2:
-1. Gigascope, point to the original NFQL paper
+
+- great content, but it does not appear that this paper is a giant leap
+  over prior work
+
+We have had a python prototype implementation [1] in the past, however
+the prototype was extremely slow and could not scale to real-world
+traces. This paper introduces NFQL, an efficient implementation of the
+language. This is not just a reimplementation in C, but the pipeline has
+been reimagined with crispier data structures to allow the
+implementation to scale to real-world traces. This is the first time we
+evaluate and show that NFQL is comparable to contemporary
+flow-processing tools. This paper discusses these implementation aspects.
+
+[1] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
+stream-based IP flow record query language,” AIMS’10
+
+- As such some of the older work is perhaps missing from the references.
+  Comes to mind: "Gigascope: A stream database for network applications
+  C Cranor, T Johnson, O Spataschek - SIGMOD03."
+
+The reference to Gigascope is in our flow query language paper [1],
+which is referenced in this paper.
+
+[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
+Record Query Language,” DSOM ’09
 
@@
 can be seen from [2] where we define queries to identify flow signatures
-of popular applications. We have had a python prototype implementation
-[3] in the past, however the prototype was extremely slow and could not
-scale to real-world traces. This paper introduces NFQL, an efficient
-implementation of the language. This is not just a reimplementation in
-C, but the pipeline has been reimagined with crispier data structures to
-allow the implementation to scale to real-world traces. This is the
-first time we evaluate and show that NFQL is comparable to contemporary
-flow-processing tools. This paper discusses these implementation aspects.
+of popular applications.
 
@@
 Popular Applications,” IM '11.
-[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
-stream-based IP flow record query language,” AIMS’10
 

### 2012-10-16 rebuttal.txt
* rebuttal: review 1   * repeating experiments (queries in repository)   * highlighting conclusions   * merge group-records   * reasoning for the metric used

@@
 Review 1:
-1. wrong track
-2. lack of description of the language
-3. too many implementation details
-4. metrics used
-5. research questions and conclusions
+
+- it has been submitted to the wrong track, as it focusses mostly on
+  implementation aspects.
+
+@juergen
+can you please address this concern?
+
+- " and merge group records." I guess it should be either 'merge' or 'group'.
+
+The merger merges group-records. Group-records are a conglomeration of
+multiple flows. The congolomeration takes place according to the query
+given at runtime.
+
+- What are your main conclusions?
+
+* NFQL, an efficient feature-complete implementation of the flow query
+  language that supersedes our original python prototype Flowy.
+* NFQL, can now process real-world sized traces in matter of minutes,
+  something which used to take days (or even weeks) in Flowy. This
+  expands the scope of flow-processing tools given the added feature set
+  capability offered by the language which is not available elsewhere.
+* The performance evaluation reveal that NFQL is comparable to
+  contemporary flow-processing tools. SiLK appears faster in some situations.
+* The evaluation queries developed as a part of this research can
+  incubate into a more general benchmark suite for flow query tools
+
+- Why don't you describe the flow query language? Instead, you describe
+  a processing pipeline. Neither does the paper contain any real
+  description of the query language, nor does it show any example query.
+
+The flow query language was introduced in [1]. It describes the language
+in full detail. This paper discusses the implementation aspects of the
+language. We briefly describe the processing pipeline because that
+helps create a stand ground necessary to understand the tool. Our
+previous usecase paper [2] defines a number of queries that use the language.
+
+[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
+Record Query Language,” DSOM ’09
+[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
+Popular Applications,” IM '11.
+
+- Why do you always use the 'Output Flows/Input Flows' metric? Explain that.
+
+The metric was used to blackbox each stage of the pipeline. This helped
+isolate the stages to benchmarks them separately to identify areas that
+may need attention.
+
+- Where/what are the used queries?
+  How can I reproduce/repeat your measurements?
+
+The queries are in the git repository referenced as a footnote in the
+paper. The JSON queries were omitted from the paper for brevity reasons
+
+The measurements can be repeated by using the cites public trace
+alongwith the queries in the repository. The benchmarking suite
+developed as a part of the research will help automate the task.
 

### 2012-10-16 rebuttal.txt
* reducing rebuttal content   * currently at 4393 (should be below 3K)

--- 2012-10-16 rebuttal.txt
@@
-Review 1:
-1. wrong track
-2. lack of description of the language
-3. too many implementation details
-4. metrics used
-5. research questions and conclusions
+Reviewer 1:
+- main conclusions?
 
-Review 2:
+* NFQL, an efficient feature-complete implementation of the flow query
+  language that supersedes our original python prototype Flowy [3].
+* NFQL, can now process real-world sized traces in matter of minutes,
+  something which used to take days (or even weeks) in Flowy. This
+  expands the scope of flow-processing tools given the added feature set
+  capability offered by the language which is not available elsewhere.
+* The performance evaluation reveal that NFQL is comparable to
+  contemporary flow-processing tools. SiLK appears faster in some situations.
+* The evaluation queries developed as a part of this research can
+  incubate into a more general benchmark suite for flow query tools
 
-- great content, but it does not appear that this paper is a giant leap
-  over prior work
+Reviewer 1, 3:
+- flow query language description (design advantages)?
+- example queries?
 
-We have had a python prototype implementation [1] in the past, however
-the prototype was extremely slow and could not scale to real-world
-traces. This paper introduces NFQL, an efficient implementation of the
-language. This is not just a reimplementation in C, but the pipeline has
-been reimagined with crispier data structures to allow the
-implementation to scale to real-world traces. This is the first time we
-evaluate and show that NFQL is comparable to contemporary
-flow-processing tools. This paper discusses these implementation aspects.
+The flow query language was introduced in [1]. It describes the language
+in full detail. This paper discusses the implementation aspects of the
+language. We briefly describe the processing pipeline because that
+helps create a stand ground necessary to understand the tool. Our
+previous use-case paper [2] defines a number of queries that use the language.
 
-[1] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
-stream-based IP flow record query language,” AIMS’10
+Reviewer 1:
+- 'Output Flows/Input Flows' metric?
 
-- As such some of the older work is perhaps missing from the references.
-  Comes to mind: "Gigascope: A stream database for network applications
-  C Cranor, T Johnson, O Spataschek - SIGMOD03."
+The metric was used to blackbox each stage of the pipeline. This helped
+benchmarks them separately to identify areas that may need attention.
 
-The reference to Gigascope is in our flow query language paper [1],
-which is referenced in this paper.
+Reviwer 1:
+- used queries? reproduce/repeat your measurements?
 
-[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
-Record Query Language,” DSOM ’09
+The queries are in the git repository referenced as a footnote in the
+paper. The JSON queries were omitted from the paper for brevity reasons
+The measurements can be repeated by using the cited public trace
+alongwith the queries in the repository. The benchmarking suite
+developed as a part of the research will help automate the task.
 
-Review 3:
+Reviewer 4:
+- filtering before splitting?
 
-- design advantages of the proposed tool are not explained, if this is
-  because NFQL has already been proposed in another paper, then this
-  work is derivative to a high degree
-- again, because the proposed method (not tool) is not explained, and
-  explanation of related tools lacks functional and methodological
-  detail, it is difficult to judge WHY the proposed tool is better
+The filtering _does_ happen before splitting. In fact, NFQL is
+implemented in a way to give away the need for a splitter entirely. As
+such, there is no splitter in the NFQL implementation. This is described
+in section IV-B: Splitter. This is one of the performance optimizations
+that makes NFQL scale well with real-world traces.
 
-The flow query language was introduced in [1]. It describes the _design_
-advantages of the language. In fact, the expressiveness of the language
-can be seen from [2] where we define queries to identify flow signatures
-of popular applications.
+Reviewer 4:
+- filter flows on the fly?
 
-[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
-Record Query Language,” DSOM ’09
-[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
-Popular Applications,” IM '11.
-
-- the method is a -QL -- a query language, which assumes fixed
-  storage, while recent research on traffic processing moves away from
-  the store-then-process approach
-
-The query language of NFQL is _indeed_ stream-based. The processing
-pipeline consists of a number of elements, which are connected with each
-other via pipes. Each element gets a input _stream_, performs some sort
-of operation on it and the output _stream_ is piped to the next element.
-This only requires that NFQL maintains state information between the
-elements. NFQL does not necessarily follow a store-then-process approach
-
-Review 4:
-
-- can you explain why you need the splitter before the filtering blocks
-and not the other way round? IMHO filtering should happen before
-splitting (fig 1)
-
-The the order of execution of the pipeline stages in the flow query
-language [1] and in the NFQL implementation is different.  According to
-the language specification, as briefly described in section III:
-
-"The pipeline starts with the splitter. It takes the flow-records data
-as input in flow-tools compatible format. The splitter is responsible to
-duplicate the input data out to several branches without any processing
-whatsoever. This allows each of the branches to receive an identical
-copy of the flow data to process it independently."
-
-However, we (and as did the reviewer) realized that in order to
-efficiently process large traces, the NFQL implementation needs to sway
-away from this design decision. As such, NFQL is implemented in a way to
-give away the need for a splitter entirely, as described in section IV-B
-
-"NFQL uses identifiers to reference a flow record in the char array.
-This eliminated the need to copy all the flow-records when moving ahead
-in the processing pipeline. As a result, there is no dedicated splitter
-stage in the execution engine. Each branch references the flow records
-from a common memory location. This helps keep the memory costs at a
-minimum when multiple branches are involved."
-
-As a result, there is no splitter in the NFQL implementation.
-
-[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
-Record Query Language,” DSOM ’09.Berlin: Springer-Verlag, 2009, pp. 15–28
-
-- can you explain why you do not filter flows on the fly?
-
-We _do_ filter flows on the fly. This is described in section IV-B
-
-"The execution engine, as defined by the flow query language must read
-all the flow records of a supplied trace into memory before starting the
-processing pipeline. Since, the filter stage uses the supplied set of
-absolute rules to make a decision on whether or not to keep a flow
-record; it has to pass through the whole in-memory recordset again to
-produce the filter results. This technique involves multiple linear runs
-on the trace and therefore slows down when the ratio of number of
-filtered records to the total number of flow-records is high. We
-optimized this behavior in NFQL by merging the filter stage with
-in-memory read of the trace. This means, a decision on whether or not to
-make room for a record in memory and eventually hold a pointer for it in
-filter results is done upfront as soon as the record is read from the
-trace. In addition, if a request to write the filter stage results to a
-flow-tools file has been made, the writes are also made as soon the
-filter stage decision is available, thereby allowing
-reading-filtering-writing to happen in O(n) time, where n is the number
-of records in the trace."
+We _do_ filter flows on the fly.
+This is described in section IV-B: Filter
 
@@
 
-- in IV.B you talk about indexes. What indexes have you used. Please cover
-this topic better
+Reviewer 2:
+- older references (gigascope)?
 
-Each flow record is stored as a char array. The trace is stored as an
-array of these flow records. This allows NFQL to reference any flow
-record using identifiers (indexes) in the array for O(1) retrieval.
+The reference to Gigascope is in our flow query language paper [1],
 
-- The work described on this paper is interesting but probably not the
-state of the art in terms of performance and features (i.e. it's not a
-step ahead but more a student project). Some examples are not easy to
-understand.
+Reviewer 2, 4:
+- not state of the art?
+- not a giant leap over prior work?
 
-We have tried to set forth the novelty of NFQL from the very outset,
-The abstract for instance, clearly briefs out the functionality that
-separates it from contemporary flow-processing tools:
+We have had a python prototype implementation [3] in the past, however
+the prototype was extremely slow. NFQL is not just a reimplementation in
+C, but the pipeline has been reimagined with better data structures to
+allow the implementation to scale to real-world traces. We show that
+NFQL is comparable to contemporary flow-processing tools. This paper
+discusses these implementation aspects.
 
-"NFQL can process flow records, aggregate them into groups, apply
-absolute or relative filters, invoke Allen interval algebra rules, and
-merge group records."
-
-In section II (related work), we reason how NFQL does more than the
-surveyed contemporary flow-processing tools:
+The abstract briefs out the novelty and is elaborated in further
+sections. For instance, In section II, we reason how NFQL does more:
 
 flow-tools and nfdump:
-"The power of filtering rules in both the tools is however mostly
-limited to absolute comparisons of flow attributes. As a result,
-relative comparison amongst different flows or querying a timing
-relationship among them is not possible." - which is possible in NFQL.
-In fact, the capabilities of flow-tools and nfdump are just one of the
+The capabilities of flow-tools and nfdump are just one of the
 five stages of the NFQL's processing pipeline. They do not support
@@
 SiLK: 
-SiLK can group flows and merge group records. However, the parameters
+* SiLK can group flows and merge group records. However, the parameters
 over which grouping/merging can be done is faily limited:
-"The grouping and merging operations can only be performed using
-an equality operator. NFQL on the other hand, provides a much richer set
-of comparisons operations, such as equal, not equal, greater than, less
-than, greater or equal, less or equal."
+* SiLK only supports sum and or aggregation operations compared to a huge
+set of aggregation options available in NFQL.
+* The NFQL merger can also match group records by putting timing
+characteristics under considerations. No such Allen operations are
+possible in SiLK.
+* Once flows are grouped together in SiLK, there is no way to unfold the
+groups to identify their member flows. As such, there is no ungrouper in SiLK.
 
-"The newly formed groups, which are passed on to the group filter, can
-also contain meta-information about the flow-records contained within
-the group using the aggregate clause defined as part of the grouper
-query. The possible aggregation operations are static, count, product,
-sum, logical and/or/xor, arithmetic mean, standard deviation, union,
-median, minimum and maximum." - whereas SiLK only supports sum and or
-aggregation operations.
-
-"The relative filtering on the groups are applied to express timing and
-concurrency constraints using Allen interval algebra [6]." - The merger
-can also match group records by putting timing characteristics under
-considerations. No such Allen operations are possible in SiLK.
-
-"The ungrouper is the final processing element of the pipeline. It
-unwraps the tuples of group-records into individual flow-records,
-ordered by their timestamps." - NFQL maintains flow membership information of
-each group record. SiLK does not provide this feature. As a result, once
-flows are grouped together in SiLK, there is no way to unfold the groups
-to identify their member flows. As such, there is no ungrouper in SiLK.
-
-In section V, we evaluate the performance of each tool against NFQL.
-flow-tools and nfdump could only participate in one of the evaluations
-due to their limited feature sets. In rest of the evaluations only SiLK
-could participate. Even so, we had to carefully design queries to find a
-common ground since most of the NFQL operations are not possible in SiLK.
-
-We believe NFQL is major step ahead in this effort. Our first python
-prototype Flowy was feature complete, but was extremely slow in
-comparison to contemporary tools and therefore could not scale to
-real-world traces.  NFQL is an efficient feature complete implementation
-of our flow-query language. It is not just a reimplementation in C, but
-the processing pipeline has been reimagined to allow it to scale to
-real-world traces. This makes NFQL comparable to current flow-processing
-tools, in addition to offering more feature-sets that are just not
-available elsewhere.
-
-It is not clear from the review which examples are not easy to
-understand, so it's hard to comment.
+[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
+Record Query Language,” DSOM ’09
+[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
+Popular Applications,” IM '11.
+[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
+stream-based IP flow record query language,” AIMS’10

### 2012-10-16 rebuttal.txt
rebuttal under 3000 characters

@@
 
-* NFQL, an efficient feature-complete implementation of the flow query
-  language that supersedes our original python prototype Flowy [3].
-* NFQL, can now process real-world sized traces in matter of minutes,
-  something which used to take days (or even weeks) in Flowy. This
-  expands the scope of flow-processing tools given the added feature set
-  capability offered by the language which is not available elsewhere.
-* The performance evaluation reveal that NFQL is comparable to
-  contemporary flow-processing tools. SiLK appears faster in some situations.
-* The evaluation queries developed as a part of this research can
-  incubate into a more general benchmark suite for flow query tools
+* an efficient feature-complete implementation of the flow query
+  language, supersedes our python prototype Flowy [3].
+* can now process real-world sized traces in matter of minutes,
+  something which used to take days in Flowy. This expands the scope of
+  flow-processing tools, given the added feature set capability
+  which is not available elsewhere.
+* Performance evaluation reveal that NFQL is comparable to
+  contemporary tools. SiLK appears faster in some situations.
+* The evaluation queries developed can incubate into a more general 
+  benchmark suite for flow query tools
 
@@
 
-The flow query language was introduced in [1]. It describes the language
-in full detail. This paper discusses the implementation aspects of the
-language. We briefly describe the processing pipeline because that
-helps create a stand ground necessary to understand the tool. Our
-previous use-case paper [2] defines a number of queries that use the language.
+The flow query language is described in full detail in [1]. This paper
+discusses the implementation aspects of the language. We briefly
+describe the processing pipeline because that helps create a stand
+ground necessary to understand the tool. Our previous use-case paper [2]
+defines a number of queries that use the language.
 
@@
 The queries are in the git repository referenced as a footnote in the
-paper. The JSON queries were omitted from the paper for brevity reasons
-The measurements can be repeated by using the cited public trace
+paper. The measurements can be repeated by using the cited public trace
 alongwith the queries in the repository. The benchmarking suite
@@
 such, there is no splitter in the NFQL implementation. This is described
-in section IV-B: Splitter. This is one of the performance optimizations
-that makes NFQL scale well with real-world traces.
+in section IV-B: Splitter.
 
@@
 
-We _do_ filter flows on the fly.
-This is described in section IV-B: Filter
-
-The previous prototype implementations (Flowy and Flowy 2.0) had this
-issue. There was a separate filter stage and the flows were not filtered
-on the fly. NFQL however _does_ filter flows on the fly. We even
-evaluated the performance improvements of NFQL against Flowy 2.0 and saw
-that NFQL now runs 10 times faster as shown in Fig 3.
+We _do_ filter flows on the fly. This is described in section IV-B:
+Filter Flowy and Flowy 2.0 had this issue. There was a separate filter
+stage and the flows were not filtered on the fly. NFQL however _does_
+filter flows on the fly. We evaluated the performance improvements
+against Flowy 2.0 and saw that NFQL now runs 10 times faster as 
+shown in Fig 3.
 
@@
 
-We have had a python prototype implementation [3] in the past, however
-the prototype was extremely slow. NFQL is not just a reimplementation in
-C, but the pipeline has been reimagined with better data structures to
-allow the implementation to scale to real-world traces. We show that
-NFQL is comparable to contemporary flow-processing tools. This paper
-discusses these implementation aspects.
+NFQL is not just a reimplementation of Flowy in C, but the pipeline has
+been reimagined with better data structures to allow NFQL to scale to
+real-world traces. We show that NFQL is comparable to contemporary
+flow-processing tools. This paper discusses these implementation aspects.
 
@@
 
-[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
-Record Query Language,” DSOM ’09
-[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
-Popular Applications,” IM '11.
-[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
-stream-based IP flow record query language,” AIMS’10
+[1] “Design of a Stream-Based IP Flow Record Query Language,” DSOM ’09
+[2] “Flow signatures of Popular Applications,” IM '11.
+[3] “Implementation of a stream-based IP flow record query language,” AIMS’10

### 2012-10-16 rebuttal.txt
forgot to commit the amend

@@
 
-* an efficient feature-complete implementation of the flow query
-  language, supersedes our python prototype Flowy [3].
-* can now process real-world sized traces in matter of minutes,
-  something which used to take days in Flowy. This expands the scope of
-  flow-processing tools, given the added feature set capability
-  which is not available elsewhere.
-* Performance evaluation reveal that NFQL is comparable to
+* an efficient feature-complete implementation, 
+  supersedes our python prototype Flowy [3].
+* can now process real-world sized traces in minutes,
+  This expands the scope of flow-processing tools, given the added 
+  feature set capability which is not available in other tools.
+* Evaluations reveal that NFQL is comparable to
   contemporary tools. SiLK appears faster in some situations.
-* The evaluation queries developed can incubate into a more general 
+* Evaluation queries can incubate into a more general 
   benchmark suite for flow query tools
@@
 
-The flow query language is described in full detail in [1]. This paper
-discusses the implementation aspects of the language. We briefly
-describe the processing pipeline because that helps create a stand
-ground necessary to understand the tool. Our previous use-case paper [2]
-defines a number of queries that use the language.
+The language is described in detail in [1]. This paper discusses the
+implementation aspects. We briefly describe the pipeline to help create
+a stand ground necessary to understand the tool. Our previous use-case
+paper [2] defines a number of queries to identify application
+flow-signatures.
 
@@
 The metric was used to blackbox each stage of the pipeline. This helped
-benchmarks them separately to identify areas that may need attention.
+benchmark them separately to identify areas that may need attention.
 
@@
 
-The queries are in the git repository referenced as a footnote in the
+The queries are in the git repository referenced in a footnote in the
 paper. The measurements can be repeated by using the cited public trace
-alongwith the queries in the repository. The benchmarking suite
-developed as a part of the research will help automate the task.
+alongwith the queries in the repository. The developed benchmarking suite
+will help automate the task.
 
@@
 implemented in a way to give away the need for a splitter entirely. As
-such, there is no splitter in the NFQL implementation. This is described
-in section IV-B: Splitter.
+such, there is no splitter in the NFQL. This is described in section
+IV-B: Splitter.
 
@@
 
-We _do_ filter flows on the fly. This is described in section IV-B:
-Filter Flowy and Flowy 2.0 had this issue. There was a separate filter
-stage and the flows were not filtered on the fly. NFQL however _does_
-filter flows on the fly. We evaluated the performance improvements
-against Flowy 2.0 and saw that NFQL now runs 10 times faster as 
-shown in Fig 3.
+We _do_ filter flows on the fly. This is described in section
+IV-B:Filter. Flowy and Flowy 2.0 had this issue. There is no longer a
+separate filter stage in NFQL. NFQL now runs 10 times faster against
+Flowy 2.0 as shown in Fig 3.
 
@@
 been reimagined with better data structures to allow NFQL to scale to
-real-world traces. We show that NFQL is comparable to contemporary
-flow-processing tools. This paper discusses these implementation aspects.
+real-world traces. This makes NFQL comparable to current flow-processing
+tools. This paper discusses these implementation aspects.
 
-The abstract briefs out the novelty and is elaborated in further
-sections. For instance, In section II, we reason how NFQL does more:
+The novelty of NFQL is discussed in section II:
 
-flow-tools and nfdump:
-The capabilities of flow-tools and nfdump are just one of the
-five stages of the NFQL's processing pipeline. They do not support
-grouping, aggregation, merging et al.
-
-SiLK: 
-* SiLK can group flows and merge group records. However, the parameters
-over which grouping/merging can be done is faily limited:
-* SiLK only supports sum and or aggregation operations compared to a huge
-set of aggregation options available in NFQL.
-* The NFQL merger can also match group records by putting timing
-characteristics under considerations. No such Allen operations are
-possible in SiLK.
-* Once flows are grouped together in SiLK, there is no way to unfold the
-groups to identify their member flows. As such, there is no ungrouper in SiLK.
+* flow-tools/nfdump features are just one of the 5 stages of the pipeline. 
+* SiLK only supports grouping/merging on an equality operator
+* SiLK only supports sum and or aggregation operations. 
+* SiLK does not support Allen operations. 
+* SiLK cannot unfold groups to identify their member flows.
 

### 2013-01-13 sections/design.tex
reduced the content to 7 pages

@@
 The flow query language consists of a number of independent stages that are
-connected to one another to form a processing pipeline.  Each stage receives
-the content from the previous, performs an operation and pushes the results to
-the next stage as shown in Fig. \ref{fig:nfql-pipeline}. A complete
-description of the semantics of each pipeline stage can be found in
-\cite{vmarinov:2009}.
+connected to one another to form a processing pipeline as shown in Fig.
+\ref{fig:nfql-pipeline}. A complete description of the semantics of each
+pipeline stage can be found in \cite{vmarinov:2009}.
 
@@
 of a flow-record against either a constant value or a value on a different
-field of the \emph{same} flow-record. The filter cannot \emph{relatively}
-compare two different incoming flow-records.
+field of the \emph{same} flow-record.
 
-The grouper performs aggregation of the input flow-records. A flow-record in
-order to be a part of the group should satisfy at-least one clause. A
-flow-record, although, can be a part of multiple clauses, it cannot be part of
-multiple groups. The grouping terms can be either absolute or relative. The
-newly formed groups, which are passed on to the group filter, can also contain
-meta-information about the flow-records contained within the group using the
-aggregate clause defined as part of the grouper query. The possible
-aggregation operations are \emph{static, count, product, sum, logical
-and/or/xor, arithmetic mean, standard deviation, union, median, minimum and
-maximum}.
+The grouper performs aggregation of the input flow-records. The grouping
+terms can be either absolute or relative. The newly formed groups, which are
+passed on to the group filter, can also contain meta-information about the
+flow-records contained within the group using the aggregate clause defined as
+part of the grouper query. The possible aggregation operations are
+\emph{static, count, product, sum, logical and/or/xor, arithmetic mean,
+standard deviation, union, median, minimum and maximum}.
 
@@
 aggregated fields) of a group-record against either a constant value or a value
-on a different field of the \emph{same} group-record. The group-filter cannot
-\emph{relatively} compare two different incoming group-records.
+on a different field of the \emph{same} group-record.
 
@@
 
-The ungrouper is the final processing element of the pipeline. It unwraps the
+The ungrouper is the final processing stage of the pipeline. It unwraps the
 tuples of group-records into individual flow-records, ordered by their
 timestamps. Any duplicate flow-records appearing from several group-records
-are eliminated and are sent as output only once.
+are eliminated.

### 2013-01-13 sections/evaluation.tex
caption consistency with graph legends
reduced the content to 7 pages

@@
-We used the first $20M$ records from a public flow trace to run a number of
-flow queries using our benchmarking suite. We used the publically available
-Trace 7 from the SimpleWeb \cite{simpleweb} repository for all the performance
-evaluations. The input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using
-the \texttt{zlib} suite. It was also converted to \texttt{nfdump} and SiLK
-compatible formats and compressed with \texttt{zlib} keeping the same
-compression level. The suite was run on a machine with $24$ cores of
-$2.5$ GHz clock speed and $18$ GiB of memory.
+We used the first $20M$ records from the publically available Trace 7 from the
+SimpleWeb \cite{simpleweb} repository for our performance evaluations. The
+input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using the \texttt{zlib}
+suite. It was also converted to \texttt{nfdump} and SiLK compatible formats
+keeping the same compression level. The suite was run on a machine with $24$
+cores of $2.5$ GHz clock speed and $18$ GiB of memory.
 
-The first set of queries attempt to stress the filter stage.  We use varying
-values on the \texttt{packet} field offset to determine the amount of flow
-records that are passed by the filter. The resultant filtered records are
-written to \texttt{flow-tools} compatible file format and compressed at
-\texttt{ZLIB\_LEVEL} $5$. The ratio of the number of filtered records in the
-output trace to the number of the flow records in the input trace is plotted
-against time. The evaluation results are shown in Fig.
-\ref{fig:benchmarks-filter}.
+The first set of queries stress the filter stage.  We use varying values on
+the \texttt{packet} field offset to control the amount of flow records that
+are passed by the filter. The resultant filtered records are written as
+\texttt{flow-tools} files and compressed at \texttt{ZLIB\_LEVEL} $5$. The
+ratio of the number of filtered records in the output trace to the number of
+the flow records in the input trace is plotted against time. The results are
+shown in Fig. \ref{fig:benchmarks-filter}.
 
@@
 comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
-lower ratios, but then again SilK and \texttt{nfdump} also use their own file
-format. As a result, the amount of data that needs to be read (or written) may
-be different to what it is for \ac{NFQL} and \texttt{flowtools}.
-\texttt{nfdump} appears to be significantly faster than the rest. This is
-because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
-that are read and written use the \texttt{lzo} compression scheme. It is
-important to note, that all the tools were single-threaded in this evaluation,
-and did not completely utilize the $24$ available cores. It comes as a
-realization, that adding \texttt{lzo} compression will drastically improve
-\ac{NFQL}'s filter performance.
+lower ratios, but then SilK and \texttt{nfdump} also use their own file
+format. \texttt{nfdump} appears to be significantly faster than the rest. This
+is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. It comes
+as a realization, that adding \texttt{lzo} compression will drastically
+improve \ac{NFQL}'s filter performance. It is important to note, that all the
+tools were single-threaded in this evaluation, and did not completely utilize
+the $24$ available cores.
 
-The second set of queries attempt to stress the grouper stage. We reuse the
-filter query that produces a $1.0$ ratio to allow the grouper to receive the
-entire trace as a filtered recordset. The grouper part of the query then
-gradually increases the number of grouping terms in the \ac{DNF} expression to
-increase the output/input ratio. The resultant groups are again written as
+The second set of queries stress the grouper stage. We reuse the filter query
+that produces a $1.0$ ratio to allow the grouper to receive the entire trace
+as a filtered recordset. The grouper part of the query then gradually
+increases the number of grouping terms in the \ac{DNF} expression to increase
+the output/input ratio. The resultant groups are again written as
 \texttt{flowtools} files using the same \texttt{zlib} compression level. The
@@
 
-The evaluation graph reveals that the performance of the \ac{NFQL} grouper
-stage is close to the performance of SiLK. The time taken by the tools are
-comparable on lower ratios, but on higher ratios, \ac{NFQL} starts to drift
-apart. Since most of the time is taken in writing the records to files, it is
-unclear whether SiLK's usage of its own file format, which may reduce
-reads/writes, is responsible for the drift on higher ratios. SiLK's
-\texttt{rwgroup} tool is also supplied a \texttt{-{}-summarize} flag to force it
-to write only the first record of each group, to make both tools write the
-same number of records to files. This gives SiLK the leverage to not store
+The evaluation graph reveals that the time taken by the tools are comparable
+on lower ratios, but on higher ratios, \ac{NFQL} starts to drift apart. Since
+most of the time is taken in writing the records to files, it is unclear
+whether SiLK's usage of its own file format is responsible for the drift.
+SiLK's \texttt{rwgroup} tool is also supplied a \texttt{-{}-summarize} flag to
+force it to write only the first record of each group, to make both tools
+write the same number of records. This gives SiLK the leverage to not store
 information about which members are part of the group. \ac{NFQL} on the other
@@
 to note that both the tools again remained single-threaded throughout the
-evaluation. SiLK took an advantage of an inherent concurency arising from how
-the query is structured as one single \texttt{bash} script using pipes. The
-pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two processes run
-concurrently, the effect of which gets more pronounced on higher ratios and
-can be a drift determining factor.  The profiling results from GNU
-\texttt{gprof} \cite{graham:1982} indicate that $60\%$ of the time is taken in
-\texttt{qsort} comparator calls.  As a result, it comes as no surprise, that
-bifurcating \texttt{qsort} invocation to multiple threads and later merging
-the results back using merge sort will help parallelize the grouper stage and
-maybe reduce the drift on higher ratios. In addition, since all of the
-evaluation queries had grouping terms using an equality comparator, \ac{NFQL}
-can introspect such a grouping rule to dynamically optimize processing
-searches using a hashtable and turn to \texttt{qsort} based grouping only as a
-fallback.
+evaluation. SiLK took advantage of an inherent concurency arising from a pipe
+between \texttt{rwsort} and \texttt{rwgroup}, which makes the two processes
+run concurrently, the effect of which gets more pronounced on higher ratios.
+The profiling results from GNU \texttt{gprof} \cite{graham:1982} indicate that
+$60\%$ of the time is taken in \texttt{qsort} comparator calls.  As a result,
+it comes as no surprise, that bifurcating \texttt{qsort} invocation to
+multiple threads and later merging the results back using merge sort will help
+parallelize the grouper stage and maybe reduce the drift on higher ratios. In
+addition, since all of the evaluation queries had grouping terms using an
+equality comparator, \ac{NFQL} can introspect such a grouping rule to
+dynamically optimize processing searches using a hashtable and turn to
+\texttt{qsort} based grouping only as a fallback.
 
-The third set of queries attempt to stress the group filter stage. We reuse
+The third set of queries stress the group filter stage. We reuse
 the filter and grouper queries that produce a $1.0$ ratio to allow the group
@@
 filter then reuses the same varying values of the \texttt{packet} field offset
-to determine the amount of groups that are filtered ahead. The filtered groups
+to control the amount of groups that are filtered ahead. The filtered groups
 are again written as \texttt{flowtools} files using the same \texttt{zlib}
@@
 
-The fourth set of queries attempt to stress the merger stage. We reuse the
-filter, grouper and group filter queries that produce a $1.0$ ratio. These
-queries are then run in two separate branches to produce identical filtered
-group records.  The merger then applies match rules to produce different
-output to input ratios. The groups that are merged are again written as
-\texttt{flowtools} files using the same \texttt{zlib} compression level. The
-ratio of the number of merged groups to twice\footnote{Each branch pushes the
-entire trace as an input to the merger.} the number of flow records in the
-original trace is plotted against time. A data point for SiLK for the $0.2$
-ratio is not available since the \ac{NFQL} query executed at that data point
-uses \texttt{OR} expressions, which are not supported by SiLK. As a result, an
-equivalent SiLK query is not formulated.
+The fourth set of queries stress the merger stage. We reuse the filter,
+grouper and group filter queries that produce a $1.0$ ratio. These queries are
+then run in two separate branches to produce identical filtered group records.
+The merger then applies match rules to produce different output to input
+ratios. The groups that are merged are again written as \texttt{flowtools}
+files using the same \texttt{zlib} compression level. The ratio of the number
+of merged groups to twice\footnote{Each branch pushes the entire trace as an
+input to the merger.} the number of flow records in the original trace is
+plotted against time. A data point for SiLK for the $0.2$ ratio is not
+available since the \ac{NFQL} query executed at that data point uses
+\texttt{OR} expressions, which are not supported by SiLK.
 
@@
     \includegraphics* [width=0.9\linewidth]{merger}
-    \caption{Merger Stage}
+    \caption{Merger Stage: NFQL vs SiLK}
     \label{fig:benchmarks-merger}
@@
 implementation takes advantage of sorted nature of filtered groups and
-therefore can significantly reduce the number of merger matches. It also writes
-a merged group record to a file only once despite the number of times it has
-matched. Without these optimizations, running such queries on the merger would
-keep the CPU churning for days without results.
+therefore can significantly reduce the number of merger matches. It also
+writes a merged group record to a file only once despite the number of times
+it has matched. Without these optimizations, running such queries on the
+merger kept the CPU churning for days without results.
 
-The last set of queries attempt to stress the ungrouper stage. They reuse the
-entire merger queries as is, but enable the ungrouper as well. This means,
-that the ungrouper now attempts to unfold the merged groups returned by the
-merger to write their member flow records to \texttt{flowtools} files.
-However, since the merger receives each flow record as its own filtered group,
-each merged group has only one member. As a result, the ungrouper ends up
-rewriting the merged groups as \texttt{flowtools} files using the same
-\texttt{zlib} compression level. This means that the execution engine end up
-taking twice the amount of time than the merger. It is important to note, that
-SiLK does not have such an equivalent ungrouping tool and is therefore not
-considered in this final evaluation.
+The last set of queries stress the ungrouper stage. They reuse the entire
+merger queries as is, but enable the ungrouper as well. This means, that the
+ungrouper now attempts to unfold the merged groups returned by the merger to
+write their member flow records to \texttt{flowtools} files.  However, since
+the merger receives each flow record as its own filtered group, each merged
+group has only one member. As a result, the ungrouper ends up rewriting the
+merged groups as \texttt{flowtools} files using the same \texttt{zlib}
+compression level. This means that the execution engine end up taking twice
+the amount of time than the merger. It is important to note, that SiLK does
+not have such an equivalent ungrouping tool and is therefore not considered in
+this final evaluation.

### 2013-01-13 sections/implementation/adaptable-compression-levels.tex
reduced the content to 7 pages

@@
-The flow-records echoed to the standard output can also be written to Netflow
-$v5$ \texttt{flow-tools} files. The \texttt{-{}-dirpath} switch allows one to
-provide a directory path where the results can be stored. Results from each
-stage of the pipeline can also be written to separate files with the increase
-in the verbosity level. In fact, \texttt{-{}-dirpath} and \texttt{-{}-verbose}
-work well together to adjust the level at which the writes are to be made.
-
 \begin{figure}[h!]
@@
 the results written to the \texttt{flow-tools} files. \texttt{zlib} supports
-$9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
-supports \texttt{-{}-zlevel} switch to allow the user to supply its desired
-choice of the of the compression level. A default level of $5$ is used for
-writes if the switch is not supplied during runtime.  Fig.
+$9$ compression levels with $9$ being the highest. \ac{NFQL} allows the user
+to supply its desired choice of the compression level.  A default level of $5$
+is used for writes if a choice is not indicated.  Fig.
 \ref{fig:engine-zlevel} shows the time taken to write a sample of records

### 2013-01-13 sections/implementation/execution-workflow.tex
reduced the content to 7 pages

@@
 
-In order to be able to make comparisons on the field offsets of such a term,
-the comparator needs to know the type of the comparison and the length of the
+In order to be able to make comparisons on the field offsets of a term, the
+comparator needs to know the type of the comparison and the length of the
 field offset. This information is parsed by execution engine once the query is
@@
 in the \texttt{char} array. This eliminated the need to copy all the
-flow-records when moving ahead in the processing pipeline. As a result, there
-is no dedicated splitter stage in the execution engine. Each branch references
-the flow records from a common memory location. This helps keep the memory
-costs at a minimum when multiple branches are involved.
+flow-records when moving ahead in the pipeline. As a result, there is no
+dedicated splitter stage in the execution engine. Each branch references the
+flow records from a common memory location. This helps keep the memory costs
+at a minimum when multiple branches are involved.
 
@@
 reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
-records in the trace. The filtered records are saved in a common location from
-where they are referenced by each branch. This helps keep the memory costs at
-a minimum when multiple branches are involved.  We used the publicly
-available Trace 7 from the SimpleWeb \cite{simpleweb} to compare the
-performance of the optimized \ac{NFQL} implementation against the one defined
-by the flow query language as shown in Fig.  \ref{fig:fv1-fv2-filter}. The
-filter stage implementation with these optimizations runs 10 times faster and
-is more pronounced on higher ratios.
+records in the trace. We used the publicly available Trace 7 from the
+SimpleWeb \cite{simpleweb} to compare the performance of \ac{NFQL} against the
+one defined by the flow query language as shown in Fig.
+\ref{fig:fv1-fv2-filter}. The filter stage implementation with these
+optimizations runs 10 times faster and is more pronounced on higher ratios.
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
+    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
+    \label{fig:fv1-fv2-grouper}
+  \end{center}
+\end{figure}
 
@@
 work on equality comparisons. A better approach, as implemented in \ac{NFQL}
-is to sort the filtered recordset on the field offsets all the requested
+is to sort the filtered recordset on the field offsets of all the requested
 grouping terms of a clause. This helps the execution engine perform a nested
 binary search to reduce the linear pass to a fairly small filtered recordset.
-However, the number of elements upon which the search has to be performed
-needs to be known at each level of the binary lookup. In order to avoid a
-linear run to count the number of elements, the parent level binary search
-invocation returns a boundary with the first and last element to enable such a
-calculation. This helps the grouper perform faster search lookups to find
-records that must group together in $O(n*lg(k))$ time with a preprocessing
-step taking $O(p*n*lg(n))$ in the average case, where $n$ is the number of
-filtered records, $p$ is the number of grouping terms in a clause and $k$ is
-the number of unique filtered records.
-
-\begin{figure}[ht!]
-  \begin{center}
-    \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
-    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
-    \label{fig:fv1-fv2-grouper}
-  \end{center}
-\end{figure}
-
-The grouping approach has further been optimized when the filtered records are
-grouped for equality. In such a scenario, the need to search for unique
-records and a  subsequent binary search goes away.  The groups can now be
-formed in $O(n)$ time with the same preprocessing step taking $O(p*n*lg(n))$.
-The performance evaluation of the grouper handling this special case against
-its behavior when handling generic cases is shown in Fig.
-\ref{fig:fv1-fv2-grouper}. Profiling the performance of the grouper on queries
-that produce higher ratios reveal that a significant amount of time is taken
-in binary search which is eliminated in the special case of equality
-comparisons.
+This helps the grouper perform faster search lookups to find records that must
+group together in $O(n*lg(k))$ time with a preprocessing step taking
+$O(p*n*lg(n))$ in the average case, where $n$ is the number of filtered
+records, $p$ is the number of grouping terms in a clause and $k$ is the number
+of unique filtered records. The grouping approach has further been optimized
+when the filtered records are grouped for equality. In such a scenario, the
+need to search for unique records and a  subsequent binary search goes away.
+The groups can now be formed in $O(n)$ time with the same preprocessing step
+taking $O(p*n*lg(n))$.  The performance evaluation of the grouper handling
+this special case against when handling generic cases is shown in Fig.
+\ref{fig:fv1-fv2-grouper}.
 
@@
 \subsubsection{Group Filter} The groupfilter stage is used to filter the
-groups produced by the grouper based on some absolute rules defined in a
-\ac{DNF} expression. The \texttt{struct} term holds information about the flow
-record offset, the value being compared to and the operator which maps to a
-unique \texttt{enum} value.  This enum value is used to map the operation to a
-specific group-filter function. The group-filter functions are auto-generated
-using a Python script.
+groups based on some absolute rules defined in a \ac{DNF} expression. The
+\texttt{struct} term holds information about the flow record offset, the value
+being compared to and the operator which maps to a unique \texttt{enum} value.
+This enum value is used to map the operation to a specific group-filter
+function.
 
@@
 of $m-$tuple (where $m$ is the number of branches) group record $ID$s was
-needed.  The result of the iterator can then be used to make a match. The
-merger stage begins by initializing this iterator passing it the number of
-branches, and information about each branch. Then, it loops over to get a new
-$m-$tuple of group record $ID$s on each iteration until the iterator returns
-\texttt{false}.
+implemented. The result of the iterator is then be used to make a match.
 
@@
 group records, however, can be reduced by sorting the group records on the
-field offsets used for a match. The \ac{NFQL} implementation optimizes the
-merger to skip over iterator permutations when a state of a current field
-offset value may not allow any further match beyond the index in the current
-branch.  For such an optimization to work, the filtered group records must be
-sorted in the order of field offsets specified in the merger clause.
-Specifying the filtered group records in any other order may lead to undefined
-behavior. This means, that if the same field offsets were used in the grouper
-stage, the terms in the group clause can be rearranged by the query designer
-to align with the order of terms in the merger clause.
+field offsets used for a match. \ac{NFQL} optimizes the merger to skip over
+iterator permutations when a state of a current field offset value may not
+allow any further match beyond the index in the current branch.  For such an
+optimization to work, the filtered group records must be sorted in the order
+of field offsets specified in the merger clause.  Specifying the filtered
+group records in any other order may lead to undefined behavior. This means,
+that if the same field offsets were used in the grouper stage, the terms in
+the group clause can be rearranged by the query designer to align with the
+order of terms in the merger clause.
 
@@
 tuples. This means that a filtered group record can be written to a file
-multiple times if it is part of multiple matched tuples. This situation is
-very common and it worsens when different branches have similar filtered
-groups records. Since, the function of the merger is to find a match of groups
-records across branches based on a predefined condition, all the group records
-across branches that satisfy the condition can be clubbed into one collection
-instead of separate tuples. All the group records within a collection can then
-be written to the file. This eliminates the inherent redundancy and
-significantly improves the merger performance.
-
-The performance comparison of this approach against the one suggested by the
-specification is tricky. The merger implementation of the original
-specification is slow. It is so slow that it keeps churning the CPU for days
-without results. The newer approach takes less than an hour. The performance
-evaluation of this newer approach is discussed in more details in the next
-section
+multiple times if it is part of multiple matched tuples. This situation
+worsens when different branches have similar filtered groups records. Since,
+the function of the merger is to find a match of groups records across
+branches based on a predefined condition, all the group records across
+branches that satisfy the condition can be clubbed into one collection instead
+of separate tuples. All the group records within a collection can then be
+written to the file. This eliminates the inherent redundancy and significantly
+improves the merger performance.
 

### 2013-01-13 sections/implementation/flow-query-intermediate-format.tex
reduced the content to 7 pages

@@
-Each stage of the processing pipeline is expressed in the JSON query as a
-\ac{DNF} expression. The elements of the conjuctive clauses are terms.
-\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
-pipeline stages of the flow query language are also encapsulated in Python
-classes and scripts have been written to facilitate the front-end parser to
-serialize these pipeline objects to a JSON flow query.  The mapping
-of the query to the structs defined in the execution engine is not trivial.
-When reading the JSON query at runtime, the field names of a NetFlow
-$v5$ record are read in as strings.  Utility functions are defined that map
-the field names to internal struct offsets and the field types and the
-operations to internal unique enum members.
+Each stage of the pipeline is expressed in the JSON query as a \ac{DNF}
+expression. The elements of the conjuctive clauses are terms.  \texttt{json-c}
+\cite{jsonc} is used to parse the flow query file. The mapping of the query to
+the structs defined in the execution engine is not trivial.  When reading the
+JSON query at runtime, the field names of a NetFlow $v5$ record are read in as
+strings.  Utility functions are defined that map the field names to internal
+\texttt{struct} offsets and the field types and the operations to internal
+\texttt{enum} members.
 

### 2013-01-13 sections/implementation/performance-optimizations.tex
reduced the content to 7 pages

@@
 pointers are therefore set as late as possible and are called from their
-respective stages just before the comparison is needed.  As a result, we save
-the computation time wasted in setting the function pointer for stage if it is
-is never executed.
-
-\ac{NFQL} pushes the filter stage out of the branches and closer to where the
-trace is originally read. This reduces the memory allocation to only records
-that passed the filter stage, thereby reducing the memory footprint of the
-execution engine.
+respective stages just before the comparison is needed.
 
@@
 merger stage proceeds only when every branch has non-zero filtered groups.
-The iterator initializer \texttt{iter\_init(\ldots)} deallocates and returns
-\texttt{NULL} if any one branch has $0$ filtered groups.  Consequently a check
-is performed in the merger to make sure \texttt{iter} is \emph{not}
-\texttt{NULL}.
+The iterator initializer deallocates and returns \texttt{NULL} if any one
+branch has $0$ filtered groups.  Consequently a check is performed in the
+merger to make sure \texttt{iter} is \emph{not} \texttt{NULL}.
 
-The results from each stage of the pipeline are echoed to the standard output
-just before the execution engine exits. This leads to an additional loop to
-echo the results, however echoes to the standard output are only used for
-debugging purposes. The writes to a file can, however, be legitimately
-requested in a real network analysis task. As a result, it is essential to
-avoid additional loops when performing writes to a file. The execution engine
-therefore writes each result record to a file as soon as it is seen by the
-pipeline stage.
+The flow-records echoed to the standard output can also be written as
+\texttt{flow-tools} files. In fact, results from each stage of the pipeline
+can be written to separate files with the increase in the verbosity level.
+This leads to additional loops over the resultsets if the writes are made at
+the end of the processing pipeline. The execution engine therefore writes each
+result record to a file as soon as it is seen by the pipeline stage.

### 2013-01-13 sections/introduction.tex
reduced the content to 7 pages

@@
 in network and user behavioral patterns of the traffic crossing the internet
-backbone. They want to use this information for the purpose of billing and
-mediation, bandwidth provisioning, detecting malicious attacks, network
-performance evaluation and overall improvement. Traffic measurement techniques
-that have been rapidly evolving in the last decade have today matured enough
-to provide such an insight.
-
-Capturing network flow records has emerged out to be one of the favored
-network measurement techniques. In this technique, packets traversing an
-obvervation point are not captured raw, instead common characteristics of the
-packet header are aggregated into flow records. They are then exported to a
-collector for further analysis.  This reduces the amount of traffic aggregated
-at the collector.
+backbone. Traffic measurement techniques that have been rapidly evolving in
+the last decade have today matured enough to provide insights based on these
+traffic patterns. Capturing network flow records is one of the favored network
+measurement technique used for this task. In this technique, packets
+traversing an obvervation point are not captured raw, instead common
+characteristics of the packet header are aggregated into flow records. They
+are then exported to a collector for further analysis.  This reduces the
+amount of traffic aggregated at the collector.
 
@@
 Systems, which allows routers to generate and export flow records to a
-designated collector. NetFlow v$9$ provides flexibility through user-tailored
-export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
-\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF based
-on NetFlow v$9$. The novelty of the standard lies in its ability to describe
-record formats at runtime using templates based on an extensible and
-well-defined information model. The data transfer mechanism is also simplistic
-and extensible by being unidirectional and transport protocol agnostic.
+designated collector. \ac{IPFIX} \cite{rfc5101} on the other hand is an open
+standard by IETF based on NetFlow v$9$. The novelty of the standard lies in
+its ability to describe record formats at runtime using templates based on an
+extensible and well-defined information model. The data transfer mechanism is
+also simplistic and extensible by being unidirectional and transport protocol
+agnostic.
 

### 2013-01-13 sections/relatedwork.tex
reduced the content to 7 pages

@@
 Stager \cite{oslebo:2006}. All these tools understand the NetFlow format while
-ntop and Stager can also process \ac{IPFIX}. They use a round robin database
-to store flow information, except for Stager which uses a MySQL database.
-Graphical presentation is done via a webinterface or in the case of FlowScan
-via RRDTool.
+ntop and Stager can also process \ac{IPFIX}. \texttt{flow-tools} and
+\texttt{nfdump} are among the most popular tools used for analyzing NetFlow
+data. \texttt{flow-tools} \cite{sromig:2000} is a suite of programs for
+capturing and processing NetFlow v5 flow records. It consists of 24 separate
+tools that work together by connecting them via UNIX pipes. It can capture,
+read, filter, and print flow records internally saved in a fixed-size format.
+\texttt{nfdump} \cite{phaag:2006} is a very similar tool that uses a different
+storage format. The power of filtering rules in both the tools is however
+mostly limited to absolute comparisons of flow attributes. As a result,
+relative comparison amongst different flows or querying a timing relationship
+among them is not possible.
 
-\texttt{flow-tools} and \texttt{nfdump} are among the most popular tools used
-for analyzing NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite
-of programs for capturing and processing NetFlow v5 flow records. It consists
-of 24 separate tools that work together by connecting them via UNIX pipes. It
-can capture, read, filter, and print flow records internally saved in a
-fixed-size format. \texttt{nfdump} \cite{phaag:2006} is a very similar tool
-that uses a different storage format. Flow records are captured using
-\texttt{nfcapd} and then processed by \texttt{nfdump}. The power of filtering
-rules in both the tools is however mostly limited to absolute comparisons of
-flow attributes. As a result, relative comparison amongst different flows or
-querying a timing relationship among them is not possible.
-
-SiLK \cite{SiLK} is a network traffic collection and analysis tool developed
-and maintained by the CERT Network Situational Awareness Team (CERT NetSA) at
-Carnegie Mellon University.  SiLK is the tool that comes quite close to
-providing similar capabilities as \ac{NFQL} and is therefore used as a
-reference point to compare the performance of the \ac{NFQL} execution engine
-in this paper. The design and implementation of SiLK, however, differs a lot
-from that of \ac{NFQL}. For instance, in SiLK there are separate tools to
+SiLK \cite{SiLK} is a network traffic collection and analysis tool that comes
+quite close to providing similar capabilities as \ac{NFQL} and is therefore
+used as a reference point to compare the performance of the \ac{NFQL}
+execution engine. The design and implementation of SiLK, however, differs a
+lot from that of \ac{NFQL}. For instance, in SiLK there are separate tools to
 perform the task of each stage of the \ac{NFQL} processing pipeline. The stage

### 2013-01-14 sections/design.tex
* condensed the abstract * deleted the first paragraph in introduction * made the flow-query language text as one paragraph

@@
 \ref{fig:nfql-pipeline}. A complete description of the semantics of each
-pipeline stage can be found in \cite{vmarinov:2009}.
-
-The pipeline starts with the splitter. It takes the flow-records data as input
-in \texttt{flow-tools} compatible format. The splitter is responsible to
-duplicate the input data out to several branches without any processing
-whatsoever. This allows each of the branches to receive an identical copy of
-the flow data to process it independently.
+pipeline stage can be found in \cite{vmarinov:2009}. The pipeline starts with
+the splitter. It takes the flow-records data as input in \texttt{flow-tools}
+compatible format. The splitter is responsible to duplicate the input data out
+to several branches without any processing whatsoever. This allows each of the
+branches to receive an identical copy of the flow data to process it
+independently.
 
@@
 of a flow-record against either a constant value or a value on a different
-field of the \emph{same} flow-record.
-
-The grouper performs aggregation of the input flow-records. The grouping
-terms can be either absolute or relative. The newly formed groups, which are
-passed on to the group filter, can also contain meta-information about the
-flow-records contained within the group using the aggregate clause defined as
-part of the grouper query. The possible aggregation operations are
-\emph{static, count, product, sum, logical and/or/xor, arithmetic mean,
-standard deviation, union, median, minimum and maximum}.
-
-The group-filter is the last processing stage of a branch. It performs
-\emph{absolute} filtering on group-records. The group-records
-that pass the filtering criterion are forwarded to the merger, the rest of the
+field of the \emph{same} flow-record. The grouper performs aggregation of the
+input flow-records. The grouping terms can be either absolute or relative. The
+newly formed groups, which are passed on to the group filter, can also contain
+meta-information about the flow-records contained within the group using the
+aggregate clause defined as part of the grouper query. The possible
+aggregation operations are \emph{static, count, product, sum, logical
+and/or/xor, arithmetic mean, standard deviation, union, median, minimum and
+maximum}. The group-filter is the last processing stage of a branch. It
+performs \emph{absolute} filtering on group-records. The group-records that
+pass the filtering criterion are forwarded to the merger, the rest of the
 group-records are dropped. The group-filter compares separate fields (or
-aggregated fields) of a group-record against either a constant value or a value
-on a different field of the \emph{same} group-record.
-
-The merger performs relative filtering on the N-tuples of groups formed from
-the N stream of groups passed on from the group-filter as input. The relative
-filtering on the groups are applied to express timing and concurrency
-constraints using Allen interval algebra \cite{fallen:1983}.
-
-The ungrouper is the final processing stage of the pipeline. It unwraps the
-tuples of group-records into individual flow-records, ordered by their
-timestamps. Any duplicate flow-records appearing from several group-records
-are eliminated.
+aggregated fields) of a group-record against either a constant value or a
+value on a different field of the \emph{same} group-record. The merger
+performs relative filtering on the N-tuples of groups formed from the N stream
+of groups passed on from the group-filter as input. The relative filtering on
+the groups are applied to express timing and concurrency constraints using
+Allen interval algebra \cite{fallen:1983}. The ungrouper is the final
+processing stage of the pipeline. It unwraps the tuples of group-records into
+individual flow-records, ordered by their timestamps. Any duplicate
+flow-records appearing from several group-records are eliminated.

### 2013-01-14 sections/introduction.tex
* condensed the abstract * deleted the first paragraph in introduction * made the flow-query language text as one paragraph

@@
-Researchers, service providers and security analysts have long been interested
-in network and user behavioral patterns of the traffic crossing the internet
-backbone. Traffic measurement techniques that have been rapidly evolving in
-the last decade have today matured enough to provide insights based on these
-traffic patterns. Capturing network flow records is one of the favored network
-measurement technique used for this task. In this technique, packets
-traversing an obvervation point are not captured raw, instead common
-characteristics of the packet header are aggregated into flow records. They
-are then exported to a collector for further analysis.  This reduces the
-amount of traffic aggregated at the collector.
-
 NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information

### 2013-01-14 sections/conclusion.tex
updated nfql source code page
back to 6 pages

@@
 We presented
-\ac{NFQL}\footnote{\texttt{\url{https://github.com/vbajpai/nfql}}}, an
-efficient C implementation of the stream-based flow query language. The
-language allows applying absolute or relative filters, aggregating flows
+\texttt{nfql}\footnote{\texttt{\url{http://nfql.vaibhavbajpai.com}}}, an
+efficient C implementation of the stream-based flow query language, \ac{NFQL}.
+The language allows applying absolute or relative filters, aggregating flows
 into groups, evaluating timing relationships among them, and merging them into
-one collection. \ac{NFQL} can execute such complex queries in matter of
+one collection. \texttt{nfql} can execute such complex queries in matter of
 minutes, thereby expanding the scope of current flow record processing tools.
-The conducted performance evaluations reveal that \ac{NFQL} is on par with
+The conducted performance evaluations reveal that \texttt{nfql} is on par with
 tools that support only absolute filters. SiLK, the only openly available
-package that provides tools that are similar to the rest of \ac{NFQL}'s
+package that provides tools that are similar to the rest of \texttt{nfql}'s
 processing pipeline appears faster. This is because it can optimize its

### 2013-01-14 sections/evaluation.tex
back to 6 pages

@@
 format. \texttt{nfdump} appears to be significantly faster than the rest. This
-is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. It comes
-as a realization, that adding \texttt{lzo} compression will drastically
-improve \ac{NFQL}'s filter performance. It is important to note, that all the
-tools were single-threaded in this evaluation, and did not completely utilize
-the $24$ available cores.
+is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. Adding
+\texttt{lzo} compression will likely improve \ac{NFQL}'s filter
+performance. Note that all the tools were single-threaded
+in this evaluation, and did not completely utilize the $24$ available cores.
 

### 2013-01-14 sections/introduction.tex
back to 6 pages

@@
 Systems, which allows routers to generate and export flow records to a
-designated collector. \ac{IPFIX} \cite{rfc5101} on the other hand is an open
-standard by IETF based on NetFlow v$9$. The novelty of the standard lies in
-its ability to describe record formats at runtime using templates based on an
-extensible and well-defined information model. The data transfer mechanism is
-also simplistic and extensible by being unidirectional and transport protocol
-agnostic.
-
-The wide applicability of this approach is easily seen from the pervasive use
-of flow records for a vibrant set of network analysis applications. For
-instance, a survey by Sperotto \textit{et al.} \cite{sperotto:2010} gives an
-overview of how network flow analysis can be used to detect Denial of Service
-attacks, scan probing target systems and the target discovery phase of worms.
-In addition, a survey by Callado \textit{et al.} \cite{callado:2009} also
-lists behaviour analysis of Internet backbone traffic and general anomaly
-detection.
+designated collector. \ac{IPFIX} \cite{rfc5101} is an open standard defined by
+the IETF based on NetFlow v$9$. The wide applicability of this approach is
+easily seen from the pervasive use of flow records for a set of different
+network analysis applications. For instance, a survey by Sperotto \textit{et
+al.} \cite{sperotto:2010} gives an overview of how network flow analysis can
+be used to detect intrusion attacks. In addition, a survey by Callado
+\textit{et al.} \cite{callado:2009} lists behaviour analysis of Internet
+backbone traffic and general anomaly detection.
 
@@
 methods.  We recently proposed a flow query language design
-\cite{vmarinov:2009} that aims to cater to such needs.  It can process
-flow records, aggregate them into groups, apply absolute or relative filters
-and invoke Allen interval algebra rules \cite{fallen:1983}. The expressiveness
-of the language can be seen from \cite{vperelman:2011} where the authors
-formulate flow queries to identify flow signatures of popular applications.
+\cite{vmarinov:2009} that aims to cater to such needs.  It can process flow
+records, aggregate them into groups, apply absolute or relative filters and
+invoke Allen interval algebra rules \cite{fallen:1983} to merge group records.
+The expressiveness of the language can be seen from \cite{vperelman:2011}
+where the authors formulate flow queries to identify flow signatures of
+popular applications.
 
-Flowy \cite{kkanev:2010} was a first feature complete Python prototype of that
-flow query language. Due to performance problems, it was superseeded by a
-complete rewrite in C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this
-paper, we introduce \ac{NFQL} which extends on Flowy 2.0, making it more
-feature complete and optimizing its execution engine with crispier algorithms.
-We show that this iteration of our work is able to scale to real-world sized
-traces and has comparable execution times to contemporary flow analysis tools.
+Flowy \cite{kkanev:2010} was a first feature complete Python prototype of
+\ac{NFQL}. Due to performance problems, it's execution engine was rewritten in
+C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this paper, we introduce
+\texttt{nfql}, which extends Flowy 2.0, making it more feature complete and
+optimizing its execution engine with crispier algorithms.  We show that this
+iteration of our work is able to scale to real-world sized traces and has
+comparable execution times to contemporary flow analysis tools.
 
-The paper is organized as follows. In section \ref{sec:relatedwork} we survey
-the current state-of-the-art flow-processing tools and reason how \ac{NFQL} is
-different from each one of them. In section \ref{sec:design} we describe the
-flow query language by discussing each stage of the processing pipeline. In
-section \ref{sec:implementation} we introduce the inner workings of \ac{NFQL}.
-It begins by providing an overview of the \ac{NFQL} architecture and the
-structure of the intermediate format used to exchange messages. The workflow of
-the execution engine is described next and is supplemented by a number of
-performance optimizations made to make the implementation scale. The section is
-followed by performance evaluations comparing \ac{NFQL} against contemporary
-flow processing tools alongwith with a discussion on its current limitation and
-future outlook in section \ref{sec:evaluation}, with conclusion in section
-\ref{sec:conclusion}.
+The paper is organized as follows. In Section \ref{sec:relatedwork} we survey
+the current state-of-the-art of flow-processing tools and we reason how
+\texttt{nfql} is different from each one of them. In Section \ref{sec:design}
+we describe the flow query language by discussing each stage of the processing
+pipeline. In Section \ref{sec:implementation} we provide an overview of the
+\texttt{nfql} architecture and we describe the workflow of the execution
+engine. A performance evaluations comparing \texttt{nfql} against contemporary
+flow processing tools alongwith with a discussion on its current limitation
+and future outlook is in section \ref{sec:evaluation}. We conclude the paper
+in Section \ref{sec:conclusion}.
+
+%\begin{figure}[ht]
+  %\begin{minipage}[b]{0.5\linewidth}
+    %\centering 
+    %\includegraphics[scale=1]{nfql-pipeline}
+    %\caption{Devebec’s Facade System}
+    %\label{fig:devebec}
+  %\end{minipage}	
+  %\begin{minipage}[b]{0.5\linewidth}
+    %\centering
+    %\includegraphics[scale=1]{nfql-architecture}
+    %\caption{MIT City Scanning Project}
+    %\label{fig:mit}
+  %\end{minipage}	
+%\end{figure}
+
+%\begin{figure}
+        %\centering
+        %\begin{subfigure}[b]{0.3\textwidth}
+                %\centering
+                %\includegraphics[width=\textwidth]{nfql-pipeline}
+                %\caption{NFQL}
+                %\label{fig:nfql-pipeline}
+        %\end{subfigure}%
+        %~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
+          %%(or a blank line to force the subfigure onto a new line)
+        %\begin{subfigure}[b]{0.3\textwidth}
+                %\centering
+                %\includegraphics[width=\textwidth]{nfql-architecture}
+                %\caption{A tiger}
+                %\label{fig:tiger}
+        %\end{subfigure}
+        %\caption{Pictures of animals}\label{fig:animals}
+%\end{figure}
 
 \begin{figure*}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{nfql-pipeline}
-\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
+  \subfloat{\includegraphics[width=0.7\textwidth]{nfql-pipeline}}
+  \label{foo}
+  \hfil
+  \subfloat{\includegraphics[width=0.3\textwidth]{nfql-architecture}}
+  \label{bar}
+  \caption{NFQL Processing Pipeline \cite{vmarinov:2009} (left) and \texttt{nfql}
+Architecture (right)}
   \label{fig:nfql-pipeline}
 \end{figure*}
+
+%\begin{figure*}[!t]
+%\centering
+%\includegraphics[width=0.9\linewidth]{nfql-pipeline}
+%\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
+%\label{fig:nfql-pipeline}
+%\end{figure*}

### 2013-01-14 sections/implementation.tex
* back to 8 pages   * reduced all figure sizes to 0.9/linewidth   * removed tcpdump, wireshark, libpcap paragraph with citations
penultimate, corrections
reduced the content to 7 pages
back to 6 pages
added JSON rfc

@@
 The \ac{NFQL} architecture primarily consists of a front-end parser backed up
-by an execution engine as shown in Fig. \ref{fig:architecture}. The execution
+by an execution engine as shown in Fig. \ref{fig:nfql-pipeline}. The execution
 engine is the brain of \ac{NFQL} where the complete pipeline is processed. It
-receives the flow query at runtime using a JSON intermediate format. The
-execution engine is designed to receive this input from either a machine
-running the front-end parser or a webservice that pushes the JSON from the
-cloud. The execution engine is written in C, however the front-end parser can
+receives the flow query at runtime using a JSON \cite{rfc4627} intermediate
+format. The execution engine is written in C, however the front-end parser can
 be written in any desired language.
 
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics* [width=1.0\linewidth]{nfql-architecture}
-    \caption{NFQL Architecture}
-    \label{fig:architecture}
-  \end{center}
-\end{figure}
+%\begin{figure}[ht!]
+  %\begin{center}
+    %\includegraphics* [width=0.8\linewidth]{nfql-architecture}
+    %\caption{NFQL Architecture}
+    %\label{fig:architecture}
+  %\end{center}
+%\end{figure}
 

### 2013-01-14 sections/evaluation.tex
changed NFQL tool to \texttt{nfql}

@@
 
-It can be seen that the performance of the filter stage in \ac{NFQL} is
+It can be seen that the performance of the filter stage in \texttt{nfql} is
 comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
@@
 is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. Adding
-\texttt{lzo} compression will likely improve \ac{NFQL}'s filter
+\texttt{lzo} compression will likely improve \texttt{nfql}'s filter
 performance. Note that all the tools were single-threaded
@@
 The evaluation graph reveals that the time taken by the tools are comparable
-on lower ratios, but on higher ratios, \ac{NFQL} starts to drift apart. Since
-most of the time is taken in writing the records to files, it is unclear
+on lower ratios, but on higher ratios, \texttt{nfql} starts to drift apart.
+Since most of the time is taken in writing the records to files, it is unclear
 whether SiLK's usage of its own file format is responsible for the drift.
@@
 write the same number of records. This gives SiLK the leverage to not store
-information about which members are part of the group. \ac{NFQL} on the other
-hand needs to allocate resources (which may take time) to keep this
+information about which members are part of the group. \texttt{nfql} on the
+other hand needs to allocate resources (which may take time) to keep this
 information in its data structures, since the ungrouper later may request to
@@
 addition, since all of the evaluation queries had grouping terms using an
-equality comparator, \ac{NFQL} can introspect such a grouping rule to
+equality comparator, \texttt{nfql} can introspect such a grouping rule to
 dynamically optimize processing searches using a hashtable and turn to
@@
 
-It can be seen that the timings of \ac{NFQL} are far apart from that of SiLK.
-It is due to the drift already created by the grouper at the $1.0$ ratio in
-the previous stage. As a result, the group filter comes into play only after
-$300$ seconds, whereas SiLK's group filtering already starts just below $150$
-seconds. Even if we normalize the graph, it can be observed that the \ac{NFQL}
-group filter has a higher slope. This is because it is only executed once the
-grouper returns, and therefore has to reiterate the groups to make a filtering
-decision.
+It can be seen that the timings of \texttt{nfql} are far apart from that of
+SiLK.  It is due to the drift already created by the grouper at the $1.0$
+ratio in the previous stage. As a result, the group filter comes into play
+only after $300$ seconds, whereas SiLK's group filtering already starts just
+below $150$ seconds. Even if we normalize the graph, it can be observed that
+the \texttt{nfql} group filter has a higher slope. This is because it is only
+executed once the grouper returns, and therefore has to reiterate the groups
+to make a filtering decision.
 
@@
 results to the file. These results although look less promising, they are way
-better than the original \ac{NFQL} merger implementation. The optimized merger
-implementation takes advantage of sorted nature of filtered groups and
+better than the original \texttt{nfql} merger implementation. The optimized
+merger implementation takes advantage of sorted nature of filtered groups and
 therefore can significantly reduce the number of merger matches. It also

### 2013-01-14 sections/implementation.tex
changed NFQL tool to \texttt{nfql}

@@
-The \ac{NFQL} architecture primarily consists of a front-end parser backed up
-by an execution engine as shown in Fig. \ref{fig:nfql-pipeline}. The execution
-engine is the brain of \ac{NFQL} where the complete pipeline is processed. It
-receives the flow query at runtime using a JSON \cite{rfc4627} intermediate
-format. The execution engine is written in C, however the front-end parser can
-be written in any desired language.
-
-%\begin{figure}[ht!]
-  %\begin{center}
-    %\includegraphics* [width=0.8\linewidth]{nfql-architecture}
-    %\caption{NFQL Architecture}
-    %\label{fig:architecture}
-  %\end{center}
-%\end{figure}
+The \texttt{nfql} architecture primarily consists of a front-end parser backed
+up by an execution engine as shown in Fig. \ref{fig:nfql-pipeline}. The
+execution engine is the brain of \texttt{nfql} where the complete pipeline is
+processed. It receives the flow query at runtime using a JSON \cite{rfc4627}
+intermediate format. The execution engine is written in C, however the
+front-end parser can be written in any desired language.
 

### 2013-01-14 sections/implementation/execution-workflow.tex
changed NFQL tool to \texttt{nfql}

@@
 
-\subsubsection{Splitter} \ac{NFQL} uses identifiers to reference a flow record
-in the \texttt{char} array. This eliminated the need to copy all the
+\subsubsection{Splitter} \texttt{nfql} uses identifiers to reference a flow
+record in the \texttt{char} array. This eliminated the need to copy all the
 flow-records when moving ahead in the pipeline. As a result, there is no
@@
     \includegraphics* [width=0.9\linewidth]{filter-fv1-fv2}
-    \caption{Filter Stage: Flowy 2.0 vs NFQL}
+    \caption{Filter Stage: Flowy 2.0 vs \texttt{nfql}}
     \label{fig:fv1-fv2-filter}
@@
 therefore slows down when the ratio of number of filtered records to the total
-number of flow-records is high. We optimized this behavior in \ac{NFQL} by
+number of flow-records is high. We optimized this behavior in \texttt{nfql} by
 merging the filter stage with in-memory read of the trace. This means, a
@@
 records in the trace. We used the publicly available Trace 7 from the
-SimpleWeb \cite{simpleweb} to compare the performance of \ac{NFQL} against the
-one defined by the flow query language as shown in Fig.
+SimpleWeb \cite{simpleweb} to compare the performance of \texttt{nfql} against
+the one defined by the flow query language as shown in Fig.
 \ref{fig:fv1-fv2-filter}. The filter stage implementation with these
@@
 
-\begin{figure}[ht!]
-  \begin{center}
+\begin{figure}[ht!] \begin{center}
     \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
-    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
-    \label{fig:fv1-fv2-grouper}
-  \end{center}
-\end{figure}
+    \caption{Grouper Stage: \texttt{nfql} (Generic) vs \texttt{nfql}
+  (Specific)} \label{fig:fv1-fv2-grouper} \end{center} \end{figure}
 
@@
 leading to a complexity of $O(n)$. The hash table approach however, will only
-work on equality comparisons. A better approach, as implemented in \ac{NFQL}
-is to sort the filtered recordset on the field offsets of all the requested
-grouping terms of a clause. This helps the execution engine perform a nested
-binary search to reduce the linear pass to a fairly small filtered recordset.
-This helps the grouper perform faster search lookups to find records that must
-group together in $O(n*lg(k))$ time with a preprocessing step taking
-$O(p*n*lg(n))$ in the average case, where $n$ is the number of filtered
-records, $p$ is the number of grouping terms in a clause and $k$ is the number
-of unique filtered records. The grouping approach has further been optimized
-when the filtered records are grouped for equality. In such a scenario, the
-need to search for unique records and a  subsequent binary search goes away.
-The groups can now be formed in $O(n)$ time with the same preprocessing step
-taking $O(p*n*lg(n))$.  The performance evaluation of the grouper handling
-this special case against when handling generic cases is shown in Fig.
-\ref{fig:fv1-fv2-grouper}.
+work on equality comparisons. A better approach, as implemented in
+\texttt{nfql} is to sort the filtered recordset on the field offsets of all
+the requested grouping terms of a clause. This helps the execution engine
+perform a nested binary search to reduce the linear pass to a fairly small
+filtered recordset.  This helps the grouper perform faster search lookups to
+find records that must group together in $O(n*lg(k))$ time with a
+preprocessing step taking $O(p*n*lg(n))$ in the average case, where $n$ is the
+number of filtered records, $p$ is the number of grouping terms in a clause
+and $k$ is the number of unique filtered records. The grouping approach has
+further been optimized when the filtered records are grouped for equality. In
+such a scenario, the need to search for unique records and a  subsequent
+binary search goes away.  The groups can now be formed in $O(n)$ time with the
+same preprocessing step taking $O(p*n*lg(n))$.  The performance evaluation of
+the grouper handling this special case against when handling generic cases is
+shown in Fig.  \ref{fig:fv1-fv2-grouper}.
 
@@
 group records, however, can be reduced by sorting the group records on the
-field offsets used for a match. \ac{NFQL} optimizes the merger to skip over
-iterator permutations when a state of a current field offset value may not
-allow any further match beyond the index in the current branch.  For such an
-optimization to work, the filtered group records must be sorted in the order
-of field offsets specified in the merger clause.  Specifying the filtered
-group records in any other order may lead to undefined behavior. This means,
-that if the same field offsets were used in the grouper stage, the terms in
-the group clause can be rearranged by the query designer to align with the
-order of terms in the merger clause.
+field offsets used for a match. \texttt{nfql} optimizes the merger to skip
+over iterator permutations when a state of a current field offset value may
+not allow any further match beyond the index in the current branch.  For such
+an optimization to work, the filtered group records must be sorted in the
+order of field offsets specified in the merger clause.  Specifying the
+filtered group records in any other order may lead to undefined behavior. This
+means, that if the same field offsets were used in the grouper stage, the
+terms in the group clause can be rearranged by the query designer to align
+with the order of terms in the merger clause.
 

### 2013-01-14 sections/implementation/performance-optimizations.tex
changed NFQL tool to \texttt{nfql}

@@
 field becomes less useful, since the members of the grouped record will always
-have the same value for that field. \ac{NFQL} realizes this redundant request
-and ignores such aggregations.
+have the same value for that field. \texttt{nfql} realizes this redundant
+request and ignores such aggregations.
 
-\ac{NFQL} has dedicated comparator functions for each type of operation and
-the type of the field offset it operates upon. It is not guaranteed that given
-the type of the query and trace, the program will eventually go through each
-stage of the pipeline. It is also possible that the program exits before,
+\texttt{nfql} has dedicated comparator functions for each type of operation
+and the type of the field offset it operates upon. It is not guaranteed that
+given the type of the query and trace, the program will eventually go through
+each stage of the pipeline. It is also possible that the program exits before,
 because there is nothing more for the next stage to compute. The function

### 2013-01-14 sections/relatedwork.tex
back to 6 pages
changed NFQL tool to \texttt{nfql}

@@
 quite close to providing similar capabilities as \ac{NFQL} and is therefore
-used as a reference point to compare the performance of the \ac{NFQL}
+used as a reference point to compare the performance of the \texttt{nfql}
 execution engine. The design and implementation of SiLK, however, differs a
-lot from that of \ac{NFQL}. For instance, in SiLK there are separate tools to
+lot from that of \texttt{nfql}. For instance, in SiLK there are separate tools to
 perform the task of each stage of the \ac{NFQL} processing pipeline. The stage
@@
 operations, such as \emph{equal, not equal, greater than, less than, greater
-or equal, less or equal}. There are also stringent requirements to how
+or equal, less or equal}. There are also stringent requirements in SiLK how
 flow-data needs to be organized before it can be piped into a tool. The

### 2013-01-14 sections/evaluation.tex
changed flowtools -> flow-tools

@@
     \includegraphics* [width=0.9\linewidth]{filter}
-    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
+    \caption{Filter Stage: \texttt{nfql} vs SiLK, \texttt{flow-tools},
+                           \texttt{nfdump}}
     \label{fig:benchmarks-filter}
@@
 It can be seen that the performance of the filter stage in \texttt{nfql} is
-comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
+comparable to that of \texttt{flow-tools} and SiLK. SiLK takes less time on
 lower ratios, but then SilK and \texttt{nfdump} also use their own file
@@
 the output/input ratio. The resultant groups are again written as
-\texttt{flowtools} files using the same \texttt{zlib} compression level. The
+\texttt{flow-tools} files using the same \texttt{zlib} compression level. The
 ratio of the number of groups formed to the number of the input filtered
-records is plotted against time. \texttt{nfdump} and \texttt{flowtools} do not
+records is plotted against time. \texttt{nfdump} and \texttt{flow-tools} do not
 support grouping, and therefore are not considered in this evaluation. The
@@
     \includegraphics* [width=0.9\linewidth]{grouper}
-    \caption{Grouper Stage: NFQL vs SiLK}
+    \caption{Grouper Stage: \texttt{nfql} vs SiLK}
     \label{fig:benchmarks-grouper}
@@
 to control the amount of groups that are filtered ahead. The filtered groups
-are again written as \texttt{flowtools} files using the same \texttt{zlib}
+are again written as \texttt{flow-tools} files using the same \texttt{zlib}
 compression level. The ratio of the number of output filtered groups to the
@@
     \includegraphics* [width=0.9\linewidth]{groupfilter}
-    \caption{Group Filter Stage: NFQL vs SiLK}
+    \caption{Group Filter Stage: \texttt{nfql} vs SiLK}
     \label{fig:benchmarks-groupfilter}
@@
 The merger then applies match rules to produce different output to input
-ratios. The groups that are merged are again written as \texttt{flowtools}
+ratios. The groups that are merged are again written as \texttt{flow-tools}
 files using the same \texttt{zlib} compression level. The ratio of the number
@@
     \includegraphics* [width=0.9\linewidth]{merger}
-    \caption{Merger Stage: NFQL vs SiLK}
+    \caption{Merger Stage: \texttt{nfql} vs SiLK}
     \label{fig:benchmarks-merger}
@@
 addition, each branch is writing the results of the filter, grouper and group
-filter stage to \texttt{flowtools} files. As a result, the amount of disk I/O
+filter stage to \texttt{flow-tools} files. As a result, the amount of disk I/O
 involved is twice as much as well. Even though each branch is delegated to a
@@
 ungrouper now attempts to unfold the merged groups returned by the merger to
-write their member flow records to \texttt{flowtools} files.  However, since
+write their member flow records to \texttt{flow-tools} files.  However, since
 the merger receives each flow record as its own filtered group, each merged
 group has only one member. As a result, the ungrouper ends up rewriting the
-merged groups as \texttt{flowtools} files using the same \texttt{zlib}
+merged groups as \texttt{flow-tools} files using the same \texttt{zlib}
 compression level. This means that the execution engine end up taking twice

### 2013-01-15 sections/evaluation.tex
re-read the paper; few grammatical corrections

@@
 format. \texttt{nfdump} appears to be significantly faster than the rest. This
-is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. Adding
-\texttt{lzo} compression will likely improve \texttt{nfql}'s filter
-performance. Note that all the tools were single-threaded
-in this evaluation, and did not completely utilize the $24$ available cores.
+is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. It goes
+to show that adding \texttt{lzo} compression will likely improve
+\texttt{nfql}'s filter performance. Note that all the tools were
+single-threaded in this evaluation, and did not completely utilize the $24$
+available cores.
 
@@
 compression level. The ratio of the number of output filtered groups to the
-number of the input group is plotted against time. The results are shown in
+number of the input groups is plotted against time. The results are shown in
 Fig. \ref{fig:benchmarks-groupfilter}.
@@
 input to the merger.} the number of flow records in the original trace is
-plotted against time. A data point for SiLK for the $0.2$ ratio is not
+plotted against time. The results are shown in Fig.
+\ref{fig:benchmarks-merger}. A data point for SiLK for the $0.2$ ratio is not
 available since the \ac{NFQL} query executed at that data point uses
@@
 results to the file. These results although look less promising, they are way
-better than the original \texttt{nfql} merger implementation. The optimized
+better than the original \ac{NFQL} merger implementation. The optimized
 merger implementation takes advantage of sorted nature of filtered groups and
@@
 merged groups as \texttt{flow-tools} files using the same \texttt{zlib}
-compression level. This means that the execution engine end up taking twice
+compression level. This means that the execution engine ends up taking twice
 the amount of time than the merger. It is important to note, that SiLK does

### 2013-01-15 sections/implementation/execution-workflow.tex
re-read the paper; few grammatical corrections

@@
 flow-record is stored in a \texttt{char} array and the offsets to each field
-are stored in separate \texttt{structs} as shown in listing
+are stored in separate \texttt{structs} as shown in Listing
 \ref{lst:c-lib-struct}. The array of such records are indexed allowing fast
@@
 generates C source code for these comparators at compile time conforming to
-the structure shown in listing \ref{lst:filter-struct}. This allows the term
+the structure shown in Listing \ref{lst:filter-struct}. This allows the term
 definitions to make runtime calls using a function name derived from the
@@
 is read from the trace. In addition, if a request to write the filter stage
-results to a flow-tools file has been made, the writes are also made as soon
+results to a \texttt{flow-tools} file has been made, the writes are also made as soon
 the filter stage decision is available, thereby allowing
@@
 means, that if the same field offsets were used in the grouper stage, the
-terms in the group clause can be rearranged by the query designer to align
+terms in the grouper clause can be rearranged by the query designer to align
 with the order of terms in the merger clause.

### 2013-01-15 sections/introduction.tex
changed NFQL tool to \texttt{nfql}
re-read the paper; few grammatical corrections

@@
 Flowy \cite{kkanev:2010} was a first feature complete Python prototype of
-\ac{NFQL}. Due to performance problems, it's execution engine was rewritten in
+\ac{NFQL}. Due to performance problems, its execution engine was rewritten in
 C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this paper, we introduce
@@
 
-%\begin{figure}[ht]
-  %\begin{minipage}[b]{0.5\linewidth}
-    %\centering 
-    %\includegraphics[scale=1]{nfql-pipeline}
-    %\caption{Devebec’s Facade System}
-    %\label{fig:devebec}
-  %\end{minipage}	
-  %\begin{minipage}[b]{0.5\linewidth}
-    %\centering
-    %\includegraphics[scale=1]{nfql-architecture}
-    %\caption{MIT City Scanning Project}
-    %\label{fig:mit}
-  %\end{minipage}	
-%\end{figure}
-
-%\begin{figure}
-        %\centering
-        %\begin{subfigure}[b]{0.3\textwidth}
-                %\centering
-                %\includegraphics[width=\textwidth]{nfql-pipeline}
-                %\caption{NFQL}
-                %\label{fig:nfql-pipeline}
-        %\end{subfigure}%
-        %~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
-          %%(or a blank line to force the subfigure onto a new line)
-        %\begin{subfigure}[b]{0.3\textwidth}
-                %\centering
-                %\includegraphics[width=\textwidth]{nfql-architecture}
-                %\caption{A tiger}
-                %\label{fig:tiger}
-        %\end{subfigure}
-        %\caption{Pictures of animals}\label{fig:animals}
-%\end{figure}
-
 \begin{figure*}[!t]
-  \subfloat{\includegraphics[width=0.7\textwidth]{nfql-pipeline}}
-  \label{foo}
-  \hfil
-  \subfloat{\includegraphics[width=0.3\textwidth]{nfql-architecture}}
-  \label{bar}
-  \caption{NFQL Processing Pipeline \cite{vmarinov:2009} (left) and \texttt{nfql}
-Architecture (right)}
-  \label{fig:nfql-pipeline}
+  \subfloat{\includegraphics[width=0.7\textwidth]{nfql-pipeline}} \label{foo}
+  \hfil \subfloat{\includegraphics[width=0.3\textwidth]{nfql-architecture}}
+  \label{bar} \caption{NFQL Processing Pipeline \cite{vmarinov:2009} (left)
+  and \texttt{nfql} Architecture (right)} \label{fig:nfql-pipeline}
 \end{figure*}
-
-%\begin{figure*}[!t]
-%\centering
-%\includegraphics[width=0.9\linewidth]{nfql-pipeline}
-%\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
-%\label{fig:nfql-pipeline}
-%\end{figure*}
